### 2020-02-29 Paper/main.tex
re-added paper

\listfiles
\documentclass[preprint,5p]{elsarticle}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}%
\usepackage{todonotes}
\usepackage{booktabs}
\usepackage{amssymb}
%\usepackage{lineno,hyperref}
%\modulolinenumbers[5]

\journal{None}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%number of pca components
\newcommand{\pca}{25}
%length of templates before pcs
\newcommand{\prepca}{92}
%number of neurons in models
\newcommand{\nneurons}{200}
%number of connections
\newcommand{\nweights}{37 thousand}

\newcommand{\noise}{\mathcal{N}}
\newcommand{\noisefloor}{\mathcal{N}_f}
\newcommand{\stochasticnoise}{\mathcal{N}_s}
\newcommand{\db}{dB}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frontmatter}

\title{The low dimensional world of echolocating bats}

%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
%\author{Elsevier\fnref{myfootnote}}
%\address{Radarweg 29, Amsterdam}
%\fntext[myfootnote]{Since 1880.}

%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
%\ead[url]{www.elsevier.com}

%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
%\cortext[mycorrespondingauthor]{Corresponding author}
%\ead{support@elsevier.com}

%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
This template helps you to create a properly formatted \LaTeX\ manuscript.
\end{abstract}

%\begin{keyword}
%\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
%\MSC[2010] 00-01\sep  99-00
%\end{keyword}

\end{frontmatter}

%\linenumbers

\section{introduction}

Echolocating bats are long-lived, mobile animals. They use the landscape intensively for foraging, roosting, and commuting. They have excellent spatial memory \citep{Barchi2013,VonHelversen2005} and can rely on sonar to navigate to notable locations like foraging grounds, drinking places \citep[see][for references]{Vanderelst2016,Vanderelst2017}. Furthermore, bats have been reported to use sonar-based recognition of landmarks, both under experimental \citep{Jensen2005,Yu2019}, and natural \citep{Verboom1999} conditions.

Research on animal navigation has revealed a range of potential strategies that can be used to reach a distant goal \citep[Reviewed by][]{Franz2000}. Irrespective of the strategy an animal uses, a mechanism for recognizing previously visited places is required \citep{Vanderelst2016,Vanderelst2017}. This implies that echolocating bats can identify previously visited sites (or landmarks defining those places) using sonar. Currently, it is still an open question of how bats use sonar to recognize locations in the environment.

One mechanism for place recognition would be for bats to parse the stream of echoes returning from the environment \cite[e.g.,][]{Yovel2009,Vanderelst2016} into localized and identified objects. However, as we argued before, it is highly questionable whether bats can interpret echoes in terms of a 3D environment model or acoustic image \citep[e.g.,][]{Vanderelst2015,Vanderelst2016,Steckel2013}. Sonar maps the three-dimensional structure of the environment on two one-dimensional waveforms arriving at the ears. While the time of arrival of echoes encodes object distance, azimuth and elevation directions are not directly available from the acoustic signal. In principle, azimuth and elevation can be estimated using spectral cues imposed by the head-related transfer function. However, complex environments return many overlapping and interfering echoes, making reconstructing the direction of origin of echoes much harder, and perhaps impossible. Extracting spatial information is further hampered by the limited temporal resolution of the sonar system \citep{Simmons1989,Wiegrebe1996,Surlykke1996}.

These limitations of the bat's sonar system suggest that sonar provides the bat with a sparse, low dimensional description of the outside 3D world. Recently, \citet{Geberl2019} confirmed that bio-sonar has a very low spatial resolution. In addition, neurophysiological data suggest that closely spaced objects are not perceived individually by echolocating bats. In contrast, neural responses to echo cascades from densely spaced objects represent a single extended stimulus \citep{Warnecke2018}.

Nevertheless, low dimensional descriptions potentially contain a lot of information and can be sufficient for recognizing locations \citep{Kuc1997b,Kuc1997}. We previously demonstrated that bats might use sonar template-based scene recognition \citep{Vanderelst2016,Vanderelst2017}. In this approach, bats are assumed to compare the acoustic signature of the environment to a set of previously stored signatures. We operationalized these acoustic signatures as time-intensity profiles, which could be derived directly from the cochlear output.

The finding that templates provide sufficient information to recognize locations prompted us to directly evaluate the hypothesis that sonar provides bats inherently with a low dimensional, yet sufficiently rich, description of the world. Testing this, we show that complex synthetic echo sequences can be mapped into a low dimensional space, after filtering them through a model of the bat's auditory periphery. Next, we show that templates derived from real bat habitats can also be projected into this space. Finally, we demonstrate that low dimensional feature space allows recognizing locations in real habitats.

Together these results suggest that echolocating bats live in a low dimensional world. The sonar system and the auditory periphery map the complexity of the 3D environment onto a small number of dimensions. Nevertheless, these low dimensional descriptions are sufficiently rich for bats to recognize locations.

\section{Methods}

\subsection{Synthetic echoes and templates}

In this section of the paper, we derive a low dimensional space for describing templates. We do this by presenting a simple generative model for echo sequences received by bats. Next, we assess whether the templates derived from these synthetic echo sequences can be projected into a low dimensional space.

We created artificial echo sequences by randomly spacing ideal point reflectors. The distances of the point reflectors were selected using a two-step process. First we selected $n_{s}$ seed points $s$, uniformly distributed over a range from 0.5 to 7 meters. Next, for every seed point $s$, we selected $n_{c}$ cloud points normally distributed around $s$, with standard deviation 0.25 m. This two step process to modeling natural reflectors was inspired by the results of \citet{Yovel2009}, who found that echoes returning from vegetation could be better modeled using a similar two-step process than simply using uniformly spaced points. We generated xx sets of points for each combination of $n_s \in [xx,xx,xx, xx]$ and $n_c \in [xx, xx, xx, xx]$.

Each set of points was converted to an impulse response with a sample frequency $f_s$ of 219 kHz. The impulse response was limited to 7499 samples. These parameters correspond to the recording length and the sampling frequency used in the acoustic data collection (see below). In converting the point to an impulse response, we modeled attenuation of the echoes due to spherical spreading (both ways) and atmospheric attenuation. The atmospheric attenuation was xx dB/m, corresponding to the attenuation at xx kHz. We did not model any variation in target strength.

Finally, the impulse responses were converted to echo sequences by convolving them with a 1 millisecond long linear frequency-swept cosine starting at 100 kHz and ending at 40 Khz. Again, these parameters correspond to those used in the acoustic data collection described below. Example of the resulting echo sequences are shown in figure \ref{fig:synthetictemplates}.

\begin{figure*}
	\centering
	\includegraphics[width=1\linewidth]{../results/synthetic_templates}
	\caption{Low dimensional mapping of synthetic echoes. Panels xx-xx, examples of artificial echo sequences. Panels xx-xx, templates derived from each of the echo sequences shown in xx-xx. The templates constructed from the artificial echoes could be projected in a lower dimensional space. A principal component analysis showed that \pca\ components captured almost all variance. Panels xx-xx, also depict (in blue) the result of projecting the templates into the \pca D space and transforming back to the orginal space. The good fit between the orginal and the reconstructed templates attests to the amount of variance captured by the \pca\ principal components.}
	\label{fig:synthetictemplates}
\end{figure*}

Artificial echo sequences were filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}. This model simulates the temporal and spectral resolution of the bat's cochlea. The output of the cochlear model was averaged across frequencies to obtain a time-intensity profile. The templates were downsampled to a sampling rate of 2.8 kHz. This sampling rate corresponds to a temporal integration of 350 $\mu$s, corresponding to what has been reported for bats' auditory systems \citep{Wiegrebe1996,Simmons1989}. This operation reduced the number of samples in a template from 7499 to \prepca\ data points. Finally, the first 6 ms of data were removed. This was done to ensure the data corresponded with the acoustic data. As indicated below, we removed the first 6 ms of those data to avoid including the pickup of the emitted signal. Selected examples of templates are depicted in figure \ref{fig:synthetictemplates}.

We used principal component analysis to assess the dimensionality of the artificial templates. We found that almost all of the variance in the templates could be captured by \pca\ components (see figure \ref{fig:synthetictemplates}xx). Some examples of templates reconstructed from the principal components are shown in \ref{fig:synthetictemplates}xx). Panel xx of figure \ref{fig:synthetictemplates}xx) shows that most of the variance not captured by these components results in some additional noise for very low values of the templates. The components are visualized in figure \ref{fig:components}.

\begin{figure}
	\centering
	\includegraphics[width=1\linewidth]{../results/components}
	\caption{Plots showing the \pca\ principal components derived from the templates constructed using synthetic echoes.}
	\label{fig:components}
\end{figure}

In this section, we proposed a simple generative model for naturalistic echoes. We assumed that echoes sequences received by bats can be considered as generated by anisotropically distributed point reflectors. Using this model, we have derived a low dimensional space ($D = \pca$) which can be used to describe the templates resulting from the artificial echo sequences. This implies that, after filtering through (a model of) the bats auditory system, these artificial echoes can be converted to low dimensional descriptions. The original echo sequences, generated by xx to xx point reflectors effectively map to \pca\ dimensional vector. In the remainder of the paper, we will assess whether this low dimensional space allows describing the echoes from real bat habitats as collected using an ensonification device.

\subsection{Natural Templates}

\subsubsection{Acoustic data}

In this work, we reuse the large library of templates constructed by \citet{Vanderelst2016}. Below, we briefly discuss the construction of the templates, and we refer to our previous work for more details. \citet{Vanderelst2016} constructed an ensonification device featuring 31 Knowles FG series microphones connected to a custom-built sonar data acquisition board \citep{Steckel2013a}. The emitter was a Sensecomp 7000 Series ultrasonic broadband speaker. The ensonification device was mounted on a pan-tilt system allowing to rotate it in azimuth and elevation. The device and pan-tilt system were mounted on a tripod.

\begin{figure}[tb]
	\centering
	\includegraphics[width=1\linewidth]{figures/datacollection}
	\caption{Pictures of the sonar data acquisition process. Top: the ensonification device used at the Israeli site at Sde Boker. Bottom: The device used at the Bristol University Royal Fort Gardens. The tape measure indicated the transect along which the device was positioned in 40 steps.}
	\label{fig:datacollection}
\end{figure}

In our previous work, the ensonification device was used to ensonify bat habitats at three different sites. However, here we only use data collected along two linear pathways through natural habitats. At the first site, a public park in Sde Boker, Israel (Fig. \ref{fig:datacollection}). The environment consisted of a walkway lined with rocks and vegetation. Unpublished bat monitoring data at the site confirmed that bat used this corridor to commute. The ensonification device was placed at 50 positions along a straight line spanning a total of 10 meters Locations were spaced 20 cm apart. At each location, echoes were collected for 31 azimuth directions (-150 to +150 degrees) and 7 elevation directions (-40 to +25 degrees). Therefore, at each location, echoes were collected from 31 x 7 = 217 directions. At each azimuth and elevation direction, three repeat measurements were recorded. In this way, a total of 32,550 echo trains (50 positions x 217 directions x 3 repeats) for each of the 31 microphones in the device were collected.

The second site, at the University of Bristol Royal Fort Gardens (UK), data were collected at 40 different positions along a line (Fig. \ref{fig:datacollection}). Locations were spaced 25 cm apart. As for the Israel site, we collected data for 217 directions, and measurements were repeated three times. This resulted in 26,040 echo trains (40 locations x 217 directions x 3 repeats) for each of 31 microphones. The recording of the echoes started with the emission of a hyperbolic frequency modulated sweep and was recorded for 34 ms ($\sim$5.7 m) at a sampling rate of 219 kHz.

\subsubsection{Template construction}

For every azimuth and elevation direction at each of the  40 or 50 locations, a biologically plausible acoustic signature of the environment was derived from the echoes, i.e., a template. Below, we briefly outline the steps used to derive templates from the acoustic data. Again, we refer to our previous work for details on this procedure \citep{Vanderelst2016}.

As we did for the synthetic echoes sequences, the echoes trains collected in bat habitats were filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}. Again, the output of the cochlear model was averaged across frequencies to obtain a time-intensity profile. The beam of the emitter used was more narrow than the typically combined hearing and emission directionality of bats \citep{Vanderelst2010a,Reijniers2010,Jakobsen2012}. Therefore, templates were averaged across three neighboring directions. The templates were downsampled to a sampling rate of 2.8 kHz. Finally, the first 6 ms of data were removed to avoid using emitted signals for analysis. Accordingly, each template consisted of \pca\ points. The processing steps outlined above resulted in 3 $\times$ 217 templates, for each location at each site, corresponding to 217 directions and three repeats.

Among the sets of templates, some had very little energy in them, notably those collected for directions without reflectors. For example, aiming the ensonification device upwards often resulted in no echoes returning. These templates could not be classified \citep{Vanderelst2016}. Therefore, we removed them from the current analyses by omitting all templates with very low energy. The resulting number of templates included in the current analyses are given in table \ref{tab:sizes}.

\subsubsection{Noise floor}

Internal (and, potentially, external) noise results in template values greater than zero, even in the absences of reflectors returning echoes. To avoid training the neural network on this noise, we estimated the noise floor in our data by collecting templates while pointing the device to empty space. The maximum template value obtained from these template measurements was used as the noise floor $\noisefloor$. The value of the noise floor $\noisefloor$ used had a numerical value of 0.15. Any template values below this value were set to this value. As we argued before \citep{Vanderelst2016}, using the empirical measurement noise to estimate $\noisefloor$ is biologically warranted as bats have a better signal to noise ratio in their auditory systems compared to our measuring device. Indeed, their dynamic range is at least 30 \db\ larger than our measurement device.

\subsubsection{Mapping templates into the \pca-dimensional space}

The templates from both sites were mapped into the \pca-dimensional space derived from the synthetic echoes (i.e., using the components depicted in figure \ref{fig:components}), described above. We assessed how well the templates collected in real habitats can be represented by the \pca\ components by mapping the templates into this space and transforming the results back to the original space. Next, we calculated the correlation (and explained variance) between the original templates and the reconstructed templates.

\subsection{Artificial Neural Networks}

We wished to assess whether the templates projected into the \pca-dimensional space still contained sufficient information to be recognized. Previously, we assessed whether templates can be used to recognize locations using, what \citet{Baddeley2012} called, a perfect memory approach. Templates were stored in a lookup table, and discriminability (or recognizability) was assessed based on perfect representations of the templates. The Euclidean distance between two echo signatures was the metric used to assess how well bats should be able to distinguish between templates. This table lookup and distance calculation require a comprehensive search of memory. Hence, in our previous work, we established that templates could be classified, but this required a significant amount of memory and computational resources.

Here, we used a more biologically plausible memory system: an artificial neural network. These are neurally inspired way of distributed information storage in artificial neural networks \citep{Mcleod1998}. However, we acknowledge that current training methods most lickely have no biological counterpart.

Simple artificial feedforward neural networks were implemented using Tensorflow 2 for Python. The first layer of the neural network added Gaussian noise to each input sample ($\sigma(\stochasticnoise) = 0.01$). The following layers were completely connected and used the rectified linear activation function. The output layer used a softmax activation function (or the normalized exponential function), normalizing the output into a probability distribution consisting of probabilities proportional to the exponentials of the output layer's activation. The loss was calculated as the cross-entropy between the target output and the actual output of the neural network. Weights were updated after each epoch using the Adam algorithm \citep{Kingma2014} .

For each site, we trained three neural networks, each taking the 25-sample templates as input. The first neural network was trained to output the azimuth direction associated with the template (31 targets). The second network was trained to return the elevation (7 different targets). The final network returned the location (50 or 40 different targets). The targets for the network were encoded using the one-hot encoding scheme.

Training a different neural network for each of these features ensured that the training algorithm paid equal attention to each feature. When training a single network for all features, the optimizer could preferably reduce errors in the location predictions as this feature has the largest variance. The full architecture can be considered as a single network with different streams for each feature (fig. \ref{fig:networks}).

We should point that we did not take any steps to avoid so-called overfitting by the neural networks. Overfitting occurs when a neural network \citep{Ghotra2017}, or other mathematical models \citep{Hawkins2004}, extract  complicated relationships between inputs and outputs that do not hold beyond the training examples. Overfitting typically reduces the ability of the model to generalize its predictive power to new data. However, in this case, the model is not required to generalize. Instead, we are only interested whether the neural networks are able to learn and classify the set of training examples.

\begin{figure}
	\centering
	\includegraphics[width=1\linewidth]{figures/networks}
	\caption{Depiction of the neural network architecture trained for each site.}
	\label{fig:networks}
\end{figure}

\subsubsection{Training examples}

We wished to ensure that the neural networks encoded the templates and were able to recognize them in the presence of realistic noise. Indeed, it can be assumed that neural noise interferes with encoding or memorizing templates. Therefore, we trained the neural networks using templates to which noise was added.

In our previous work, we describe a procedure, based on \citet{Dau1996}, to derive a realistic model for stochastic noise. In brief, we empirically determined a noise level that resulted in a 75\% correct discrimination of two artificial bat calls differing by 2 \db\ in amplitude. This criterion is a conservative estimate of the just noticeable difference in loudness in bats. Based on this procedure we assumed internal (independent) Gaussian stochastic noise $\stochasticnoise$ for each sample with $\sigma(\stochasticnoise) = 0.01$ and $\mu=0$.

\subsection{Perfect Memory}

To be able to directly compare the results of the neural networks with the perfect memory approach \citep{Baddeley2012}, we implemented a version of the classification scheme employed in our earlier work \citep{Vanderelst2016}. We generated noisy templates by adding Gaussian noise to them ($\sigma(\stochasticnoise) = 0.01$). Next, we performed the classification of these noisy templates by finding the closest template (according to the Euclidean distance) in the original set of templates. The location, azimuth, and elevation of the closest template were considered as the noisy template's classification. This was repeated ten times, and we report on the average performance across these repeats.

\section{Results}

We found that the templates derived from acoustic data collected in real habitats could be well represented by the \pca\ components derived from synthetic data. This \pca-dimensional space could represent xx\% of the variance in the templates. Figure \ref{fig:templates} shows examples of templates and their reconstructed counterparts. Therefore, although the model used to generate echo sequences was very simple, the resulting underlying \pca D feature space could adequately represent templates resulting from real acoustic data.

\begin{figure}
	\centering
	\includegraphics[width=1\linewidth]{../results/templates}
	\caption{Examples of three templates selected from the Israel and Royal Fort data (red). Blue lines depict the result of projecting the templates in the \pca-dimensions space and reconstructing by transforming the results back to the original space.}
	\label{fig:templates}
\end{figure}

In our previous work \cite{Vanderelst2016}, we showed that the templates could be used to recognize locations in bat habitats. As we have just shown that the templates can be projected into a lower \pca D feature space without loss, it can be expected that these \pca\ features can be used to locations as well. We found this to be indeed the case.

Figure \ref{fig:history} depicts the training history of all networks. This figure shows that, in all cases, training almost reached an asymptote around or before 250 epochs. However, some additional training can be seen up to xxx epochs.

\begin{figure}
	\centering
	\includegraphics[width=1\linewidth]{../results/history}
	\caption{Training history of the neural networks. The plots give the value of the loss function as a function of training epoch. Notice the logarithmic x-axis.}
	\label{fig:history}
\end{figure}

Figure \ref{fig:israelperformance} summarizes the performance of the three neural networks trained on the Israel templates. The performance was obtained by assuming that the most active output neuron represented the classification of the network. Inspecting the cumulative distribution of error reveals that the neural network encoding outperformed the perfect memory approach. The number of small errors was larger for the neural network than for the perfect memory. The neural network was able to classify about 70\% of templates with an error of less than 30 degrees in azimuth. In elevation, almost all templates were classified with an error below 15 degrees. Most (70\%) errors in location attribution were smaller than 0.5 meters. The confusion matrices confirm that most classification errors were small.

Interestingly, for azimuth and elevation, classification errors were not uniform across directions. This can be seen by the classification entropy plotted in figure \ref{fig:israelperformance}. In particular, these curves show interesting peaks around zero azimuth and medium elevations. The geometry of the environment can explain these peaks. The corridor (fig. \ref{fig:datacollection}) does not return a lot of echoes from azimuths around zero. Indeed, this direction is mostly open space. The same is true for elevations above 0. Elevations around 25 return very few echoes and this serves as a reliable indicator for classifying these templates, explaining the drop in entropy for elevation = 25 degrees.

\begin{figure*}
	\centering
	\includegraphics[width=1\linewidth]{figures/israel_performance}
	\caption{Classification results for the Israel site. Top row: cumulative error distributions for the azimuth, elevation and location classification. Middle row: Confusion matrices for the classifications. Bottom row: remaining classification entropy for each target (derived from the confusion matrices).}
	\label{fig:israelperformance}
\end{figure*}

Figure \ref{fig:royalperformance} shows the results for the Royal Fort Park data. These results mirror those in figure \ref{fig:israelperformance}. Again the neural network outperformed the perfect memory approach by having more smaller errors. Overall, the performance for this data set was somewhat better than for the Israel data.

\begin{figure*}
	\centering
	\includegraphics[width=1\linewidth]{figures/royal_performance}
	\caption{Simular as fig \ref{fig:israelperformance}, but for the Royal Fort Park site.}
	\label{fig:royalperformance}
\end{figure*}

\section{Discussion}

\subsection{The low dimensional world of bats}

We assumed echoes received by bats can be modeled as originating from a large collection of point reflectors. Passing these synthetic echoes through a model of the auditory periphery, we obtained a time-intensity profile (template) for each echo sequence. These templates could be well described by a low dimensional space of about \pca\ principal components (figure \ref{fig:synthetictemplates}). We found that the same \pca D space could be used to describe templates originating from echoes collected in real bat habitats (figure \ref{fig:templates}). Therefore, both templates derived from synthetic echoes and templates constructed from real echoes map into the same lower dimensional space. This suggest that, in general, the filtering of the auditory periphery greatly reduces the dimensionality of the description of the bat's environment.

Despite the low dimensional character of the templates, they are sufficient to recognize environments. Indeed, this was demonstrated in our previous work \citep{Vanderelst2016}. Here, we directly demonstrated that the  \pca\ principal components are sufficient to discriminate their associated locations using neural networks.

Here, we show that a low dimensional signal is sufficient for recognizing environments. This is a critical component of navigation. However, bats do more than navigating. They also use sonar to control flight, for example, avoid obstacles or following guiding structures. The description the bat has access to for flight control is has the same low level character than that used for place recognition. Therefore, our work suggests that this low level description is sufficient for flight control as well. We have shown before that a low level description could be sufficient for obstacle avoidance and corridor following \citep{Vanderelst2015a,Mansour2019}. However, in that work, we converted both the echoes at the left and at the right ear to template-like representation. Subsequent work will have to demonstrate that the parameters of the \pca D space presented here can be used directly for controlling flight. In other words, we conjecture it should be possible to extract \pca\ features from an echo signal and convert this vector into a motor command. One can presume a neural network that converts the \pca\ features to flight control parameters.

\subsection{Template encoding}

We were unable to find numbers on the number of neurons in bat brains. However, the mouse cortex has been estimated to house about $9 \cdot 10^4$ neurons or $7 \cdot 10^8$ synapses per cubic millimeter \citep{Braitenberg2013}. The neural networks presented in this paper consisted of about \nneurons\ neurons and \nweights\ connections. Therefore, in as much as these neural networks can be considered as biologically plausible models of memory, they are orders of magnitude smaller than a single cubic millimeter of mouse cortex. Let alone the complete cortex, which has been estimated to contain 75 million neurons \citep{Williams2000}.

Despite their small size, the networks in this paper were able to encode (memorize) the templates as least as well as the the perfect memory approach. The neural networks presented here are tiny compared to the density of cortical matter. However, the networks are also small compared to the data they encode. Table \ref{tab:sizes} lists the number of weights for the artificial neural networks and compares this number with the size of the \pca-dimensional templates. The number of floats making up the templates is larger than the number of weights, by a factor of 2-3. Therefore, the networks have a smaller memory footprint than the data they encode. It should be noted that we have not made an attempt to try and to optimize (minimize) the size of the networks.

\begin{table}
	\centering
	\begin{tabular}{lrrrrl}
		\toprule
		Site &       T &     n &      W &  W (nb) &      R \\
		\midrule
		Israel &  179730 &  5991 &  73050 &   72000 &  40.6\% \\
		Royal &  214890 &  7163 &  50745 &   49875 &  23.6\% \\
		\bottomrule
	\end{tabular}
	\label{tab:sizes}
	\caption{Table indicating the size of both template data sets and number of weights in the networks. The template size $T$ is obtained as $\pca \times n$, with $n$ the number of templates in the set. The number of weights $W$ includes all connections between neurons as well as the bias weights for each of the three networks trained. $W (nb)$ omits the bias weights. The ratio $R$ gives the ratio $W/T$.}
\end{table}

\subsection{Conclusion}

From our current work, in combination with previous results \citep{Vanderelst2016,Vanderelst2015a,Mansour2019}, we conclude that the auditory periphery of bats maps complex echoes to a limited number of features. These features can be efficiently stored to recognize scenes. Moreover, we conjecture that a (binaraural) version of the features should allow for sonar-based flight control. As such, a view of echolocating bats living in a low dimensional sensory space emerges.

%Together these two findings lend additional credibility to our hypothesis that bats could recognize locations using their acoustic signature, operationalized as templates. Before, we demonstrated that templates are highly specific to locations and directions. Here, we show that storing such templates does not require extensive neural hardware. Small neural networks can be used to store many (6000-7000) templates. Moreover, this is feasible because templates can effectively be compressed, which we demonstrated here by employing PCA and neural networks of a limited size.

%It is well know that even small, shallow, neural networks are remarkably good at approximating potentially very complex function, even to the extent that it becomes valid to ask why they so powerful. \citet{Lin2017a} suggested that most datasets we are interested in originate from some simple physical law or some generative model with relatively
%few free parameters. In the case of templates, the physical laws that generate them are quite straightforward. In addition, it can be assumed that the structure of natural environments can be described by generative models. Therefore, the ideas put forward by \citet{Lin2017a} might explain, on an abstract level why templates should not require extensive neural hardware to encode.

\bibliography{references}
\end{document}

### 2020-02-29 Paper/old.tex
re-added paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Document type, global settings, and packages
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt]{report}

\usepackage{graphicx}  %for images and plots
\usepackage{setspace}
\usepackage{hyperref}
\usepackage[numbers]{natbib}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amscd}
\usepackage{upgreek}
\usepackage[letterpaper, left=1in, right=1in, top=0.75in, bottom=0.75in]{geometry}
\usepackage{subcaption}
\usepackage[section]{placeins}

%Dieter: I have switched this of for a second
%\usepackage{arev}
%\usepackage[T1]{fontenc}

%%%%%%%%%%%%%%%%%%%%%%
% Start of Document
%%%%%%%%%%%%%%%%%%%%%%

\title{Master's Thesis ACA} %Write title here
\author{darsha4590 }
\date{August 2019}
\begin{document}
\maketitle
\tableofcontents

\begin{abstract}
    Echo locating bats use sonar for navigating to salient locations. Navigation requires the recognition of locations. Previously, we proposed template-based place recognition to explain how bats recognize locations using sonar. We hypothesized that bats recognize places based on their echo signature and do not compute the 3D layout of a scene. Collecting a large corpus of echoes from different habitats and processing these using a model of the bat's auditory system, we showed that biologically plausible templates contain sufficient information to recognize previously visited locations. In this previous work, we assumed that the bat has perfect memory. We assumed that the bat stores all previously encountered templates and recognizes locations by comparing a new template with all stored templates. As the bat would need to store many templates, this kind of perfect memory is computationally expensive and biologically implausible. In this work, we explore how bats could store and use templates more efficiently using neural networks, a biologically plausible substrate for the memory of bats. We trained a feed-forward neural architecture to return the location and viewing direction associated with a given template as a model of a bat recognizing its current location and flight direction from the echoes. The performance of this imperfect memory approach was comparable to the performance of the previously implemented perfect memory, where performance was fundamentally limited by the noise inherent in sonar data and the signal to noise ratio of the templates. Hence, our results indicate that a large corpus of sonar data can be stored in a neural network (which is small compared to the neural capacity available to bats). This work indicates that our hypothesized mechanism for scene recognition is biologically plausible. In addition, our work opens up avenues for more efficient implementations of sonar-based navigation for robots using neuromorphic hardware.
\end{abstract}

\chapter{Introduction}

Echolocating bats have excellent spatial memory \cite{barchi2013spatial} and can use sonar to navigate to salient locations like foraging grounds, drinking places (look up references from place recognition paper). \textit{Myotis spp} and \textit{Phyllostomus hastatus} deprived of sight were able to return to their roost \cite{stones1969use}, showing that successful navigation is possible relying solely on echolocation. A fundamental prerequisite for navigation is place recognition \cite{schnitzler2003spatial}. Currently, it is largely unknown how bats recognize places based on the echoes they return.

One potential mechanism for place recognition would be for bats to parse the stream of echoes returning from the environment \cite[e.g.,][]{Yovel2009,Vanderelst2016} into localized and identified objects (see figure \ref{fig:3Dlayout}). Subsequently, the bat could match this representation to previously-stored representations to recognize the present scene. However, as \citet{Vanderelst2016} and \citet{Vanderelst2017} argued, several limitations of bio-sonar make it unlikely that bats can localize and identify the objects making up complex scenes. Most notable, among these limitations, is the limited temporal and spatial resolution of bio-sonar. Echoes arriving (almost) simultaneously interfere and are integrated by the auditory system. This probably makes it impossible for the bat to reconstruct the 3D layout of a scene returning overlapping echoes from vegetation and other complex reflectors.

\begin{figure}
    \includegraphics[width=\linewidth]{images/3Dlayout.JPG}
    \caption{This is sample 3D reconstruction}
    \label{fig:3Dlayout}
\end{figure}

As an alternative hypothesis, \citet{Vanderelst2016} have proposed that bats might use sonar template-based scene recognition. In this approach, bats are assumed to compare a cochlear output to a previously set of stored templates. This hypothesis is analogous to view-based navigation believed to be used by insects \citep{Zeil2003} \citet{Vanderelst2016} collected echoes from a range of bat habitats and converted these, using a biologically plausible model of the peripheral auditory processing in bats, into templates. The authors verified whether the resulting templates had properties that would allow them to be used for navigation. They found that the templates allowed to uniquely identify locations (provided the echoes contain sufficient energy). They also found the templates to vary smoothly as a function of translation and rotation through the environment. These findings supported the hypothesis that bats could use acoustic templates as signatures for recognizing previously visited places.

The templates used by \citet{Vanderelst2016} were constructed using a biologically plausible model of the bat's auditory system. However, the memory model they used was not bio-inspired. These authors employed a perfect memory approach: all templates were stored in a look-up table, and discriminability was assessed based on perfect representations of the templates. The Euclidean distance between two echo signatures was the metric used to identify templates. This requires a comprehensive search of the memory. Templates could be classified, but this required a significant amount of memory and computational resources.

In biology, perfect memory is highly implausible. Therefore, we set out to assess whether templates can be stored and recognized using a more biologically plausible substrate. In particular, we test whether the echo templates can be stored and recognized by training a neural network.

Sonar is quite common in robotics. Currently, applications of artificial in-air sonar are almost exclusively limited to simple ranging and obstacle detection. However, as bats demonstrate, sonar can also be used for navigation and localization. Indeed, \citet{steckel2013batslam} has demonstrated... Understanding how templates could be used effectively and efficiently could help develop a methodology for navigation and localization in robots. Especially, small energy constrained robots (micro uavs) to exhibit autonomy we can use sonar as one sensory method for effective navigation. A neural network based approach suggest in our work will help design efficient and simple computational model suitable for small processors associated with small robots. A neural approach can be a great prospect for implementation on a neuromorphic hardware which will make it an exceptionally efficient system

\chapter{Data Collection and Pre-Processing}

\section{Echo data collection}

In this work, we reused the templates constructed by \citet{Vanderelst2016}. Below we briefly discuss the constructing of the templates, and we refer to our previous work for more details. \citet{Vanderelst2016} constructed an ensonification device featuring 31 Knowles FG series microphones connected to a custom-built sonar data acquisition board \citet{steckel2013batslam}. The emitter was a sensecomp 7000 Series ultrasonic speaker. The ensonification device was mounted on a pan-tilt system allowing to rotate it from -150 to 150 degrees in azimuth and from -25 to 40 degrees in elevation.

The ensonification device was used to ensonify bat habitats at three different sites (described below). At each location at the three sites, echoes were collected for 31 azimuth directions and 7 elevation directions (i.e. $31 \times 7 = 217$ directions). At each azimuth and elevation direction, three repeat measurements were recorded. The recording of the echoes started with the emission and was recorded for 34 ms ($\sim$5.7 m) at a sampling rate of 219 kHz. Data were collected at three different sites: St Andrews Park, a public park in Bristol (UK), a public park in xxx (Israel), and at the Royal Fort Gardens, a University of Bristol park (UK). The structure of the data is depicted in figure \ref{sites}.

At St Andrews Park, data were collected at 12 locations, consisting of four open sites, four semi cluttered sites, four cluttered sites which represent different densities of environments for bats. As such, data consisted of 7812 (12 positions x 217 directions x 3 repeats), echo trains from each of the 31 microphones. At the second site, in Israel, the ensonification device was placed at 50 positions along a straight line spanning a total of 10 meters. Locations were spaced 20 cm apart. In this way, a total of 32,550 echo trains (50 positions x 217 directions x 3 repeats) for each of the 31 microphones were collected. At the Royal Fort Gardens, data were collected at 40 different positions along a line. Locations were spaced 25 cm apart, resulting in a total of 26,040 echo trains (40 locations x 217 directions x 3 repeats) for each of 31 microphones.

\begin{figure}[ht]
	\includegraphics[width=\linewidth,,height=11cm]{images/sites.JPG}
	\caption{Illustration of the data collected}
	\label{sites}
\end{figure}

\section{Template construction}

For every azimuth and elevation direction at each of the 12, 40 or 50 locations, the following processing steps were used to construct a template:
\begin{itemize}
    \item Application of a cochlear model and dechirping
    \item Averaging across frequencies and microphones
    \item Averaging across 3x3 directions
    \item Reducing the sampling frequency to 2.8KHz
    \item Removing the emission from the template
\end{itemize}

First, echo trains were individually filtered using a model of the bat's auditory periphery as proposed by \citet{wiegrebe2008autocorrelation}. This model simulates the temporal and spectral resolution of the bat's cochlea. The model consists of a Gammatone filterbank with central frequencies ranging from 30KHz to 100KHz in steps of 5KHz. Each channel is exponentially compressed (exponent value of 0.4), and low pass filtered. The cochleogram returned by the model is dechirped by shifting each frequency channel in time to maximize activation, corresponding to the pick up of the emitted signal, is aligned across all frequency channels. Next, the first 5.8ms is removed across each frequency channel as this contains the pick up of the emitted signal.

The beam of the emitter used was more narrow than the typically combined hearing and emission directionality of bats \cite{reijniers2010morphology},\cite{jakobsen2013intensity}. Therefore, templates were averaged across three neighboring directions. Finally, the templates were downsampled to a sampling rate of 2.8KHz. This sampling rate corresponds to a temporal integration of 350 $\mu$s. This operation reduced the number of samples in a template from 7499 to 96 data points. Finally, the first 14 points were removed to avoid pick up of emitted signal for analysis. Accordingly, each template consisted of 82 points. The processing steps outlined above resulted in 3 $\times$ 217 templates, for each location at each site, corresponding to 217 directions and three repeats.

Internal (and, potentially, external) noise results in template values greater than zero, even in the absences of reflectors returning echoes. To avoid training the neural network on noise, we estimated the noise floor in our data by collecting templates while pointing the device to empty space. The maximum template value obtained from these template measurements was used as the noise floor \textit{nf}. The value of the noise floor used had numerical value of \textit{nf} = 0.15. Any template values below the noise floor were set to this value. Some sample templates can be viewed in Figure \ref{templates}.

\begin{figure}[ht]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{images/templats/First201.jpg} % first figure itself
        %\caption{first figure}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{images/templats/Second282.jpg} % second figure itself
        %\caption{second figure}
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{images/templats/Third2141.jpg} % second figure itself
        %\caption{second figure}
    \end{minipage}
    \caption{Examples of templates used at the St.Andrews location. The templates are from semi-cluttered 3 environment for 3 different azimuth's and elevations}
    \label{templates}
\end{figure}

%The original raw data of the template data is the template constructed on a original sampling rate of 219 KHz and the number of data points on the template are 7,499 points but the bat auditory model (ref Weigrabe) has a temporal integration interval of 350 $\mu$s so a corresponding sampling rate of 2.8 KHz was used

%\subsection{\textbf{Noise Floor}}
%In case there were templates with no return echo which still contains non-zero signal values for the templates which were part of the environment noise and internal noise of the ensonification device. In order to avoid these noise values to contribute to the classification performance. (Place recognition paper) has evaluated the noise floor using the templates when the ensonification device was pointed to open spaces. T

\section{Principal Component Analysis}

Templates are time-energy time series with some level of interdependence between nearby samples. The correlation between template values is problematic because...  We used principal component analysis (PCA) to convert the templates in several uncorrelated template features. PCA reduces the dimensions by creating new uncorrelated features while maintaining the variance in the data. We found that reducing the 82-point templates to 25 principle components allowed capturing 99.5\% variance across the three sites.
The variance captured per principal component and cumulative variance captured is illustrated in Figure \ref{PCAs}

\begin{figure}[ht]
    \centering
    \begin{minipage}{0.5\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{images/AndrewsPCsVariance.jpg} % first figure itself
        %\caption{first figure}
    \end{minipage}\hfill
    \begin{minipage}{0.5\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{images/AndrewsCumilativePCsVariance.jpg} % second figure itself
        %\caption{second figure}
    \end{minipage}
    \begin{minipage}{0.49\textwidth}
        \centering
        \includegraphics[width=1.0\textwidth]{images/PCsVariance.jpg} % second figure itself
        %\caption{second figure}
    \end{minipage}
    \begin{minipage}{0.49\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{images/CumilativePCsVariance.jpg} % second figure itself
        %\caption{second figure}
    \end{minipage}
    \caption{Given are the variance retained per principal component for the three sites}
    \label{PCAs}
\end{figure}

\section{Synthetic Data Generation}

We wished to ensure that the neural network memorized the templates and was able to classify them when faced with noise. Indeed, both in the biological system modeled and the envisioned application, templates will always be corrupted by some noise. Therefore, we added noise to the templates to generate the training data.

We estimated the measurement noise level based on the data. At every location and orientation, echo measurement was conducted 3 times, and the templates generated were also across these three measurements. To create synthetic data, we assumed that the environment remains the same, and the data variability would arise just from the measurement noise.

Using the empirical measurement noise to create training data is biologically warranted as bats have a better signal to noise ratio in their auditory systems compared to our measuring device. Indeed, their dynamic range is at least 30 dB larger than our measurement device. For robotic applications, the measurement noise in our current study is a representative example of the noise that can be expected in other ultrasonic devices.

%To make sure the neural network is less data sensitive it was important to add noise related data to the training process. As a new recorded template would always have some variability training the neural network with just the original data would most likely give margins for generalization errors.
%Biologically The additional data would give us some better generalization accounting for the instrument inherent measurement noise.

\begin{figure}[htbp]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{images/201.jpg} % first figure itself
        %\caption{first figure}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{images/282.jpg} % second figure itself
        %\caption{second figure}
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{images/2141.jpg} % second figure itself
        %\caption{second figure}
    \end{minipage}
    \caption{Three noise generated templates across three different templates}
    \label{NoiseData}
\end{figure}

Post the synthetic data generation added to the original data. The data set was randomly shuffled and setup the set for implementing a neural network for training and testing.

Setup the table for data amount for each site and the dimensions of the feature table for various sites \ref{process}
\begin{figure}[ht]
	\includegraphics[width=\linewidth,,height=11cm]{images/ProcessFlow.JPG}
	\caption{Process flow of the templates for neural network were generated across all three sites}
	\label{process}
\end{figure}

\chapter{Template Classification}
There exists various approaches to classify the template data. The original approach used in (cite place recognition) is based on a perfect memory approach where every template is stored in memory, the new measured template is compared to every template stored in memory and the minimum distance value is used to classify the corresponding best match. This approach is very memory extensive and needs a fair bit of computation and this mechanism is not possibly how bats localize in real time. The methodology used in this study has been inspired from the fact that bat brain is made of millions of complex neural nets to make complex decisions.

\section{Artificial Neural Network (ANN):}
Bats brain is a complex neural structure we are trying to replicate the a process used by bats to localize themselves using sonar templates. Machine Learning algorithms are resource intensive training highly complex data needs a complex structures which are computationally expensive. There has been significant development in the field of graphics processing units (GPUs) and Tensor processing units (TPUs) which allow us to handle complex large computations but result in high energy consumption so new research has been evolving in the neuromorphic hardware where neural circuits are made on a chip which does the same function of a complex neural network on a small chip and can be exceptionally fast. Inspired by this application an Artificial neural network (ANN) has been chosen as the classification methodology. A Feed Forward neural network architecture is explored to find a possible efficient neural network methodology to efficiently and effectively classify the template data. Although a neural network will take computational resources to train the best model. The model once implemented will be less memory intensive and computationally less expensive when we increase the amount of data to be processed during real time classification. The figure \ref{ANN} represents the basic structure of a ANN. It consists of the input feature vector which is passed through a number of hidden layers of neurons and output of each layer is acted upon by an activation function, the final layer represents the one hot representation of the template labels
\begin{figure}[ht]
	\includegraphics[width=15cm,,height=6cm]{images/ANNfigure.png}
	\caption{A feed Forward Neural Network}
	\label{ANN}
\end{figure}

%\currentpdfbookmark{Title Page}{titlePage}  %add PDF bookmark for this page

\bibliographystyle{apalike}
\bibliography{references}
\end{document}

### 2020-03-01 fit.py
intermediate

@@ -51,3 +51,13 @@ print(table)
 
 pyplot.imshow(table)
 pyplot.show()
+
+#%%
+selected = 1
+t = templates_synthetic[selected,:]
+r = reconstructed_synthetic[selected,:]
+
+pyplot.plot(t)
+pyplot.plot(r)
+
+pyplot.show()

### 2020-03-01 library/generate_functions.py

@@ -24,17 +24,23 @@ def create_emission():
     return emission
 
 def generative_model(parameters):
-    minimum_distance = 0.25
-    standard_deviation = 0.05
+    minimum_distance = 1
+    maximum_distance = 7
+    standard_deviation = 0.1
     n_seed_points = parameters.pop('n_seed_points')
     n_cloud_points = parameters.pop('n_cloud_points')
     keys = parameters.keys()
     if len(keys) > 0: print('***********Warning', keys)
     distances = numpy.array([])
-    seed_points = random.uniform(minimum_distance, 7, n_seed_points)
+    seed_points = random.uniform(minimum_distance, maximum_distance, n_seed_points)
     for loc in seed_points:
         g = random.normal(loc=loc, scale=standard_deviation, size=n_cloud_points)
         distances = numpy.concatenate((distances, g))
+
+    # add close echo to ensure that even seqs with low n have a strong echo
+    close = random.uniform(minimum_distance, minimum_distance+1, 1)
+    g = numpy.array([minimum_distance])
+    distances = numpy.concatenate((distances, close))
     return distances
 
 def distances2echo_sequence(distances, caller, emission):

### 2020-03-01 script_plot_template_examples.py

@@ -25,8 +25,8 @@ mx_templates = 0.35
 mn_templates = 0.15
 
 label0 = '5_5_0'
-label1 = '40_40_0'
-label2 = '160_160_0'
+label1 = '10_10_0'
+label2 = '250_250_0'
 
 echo_0 = synthetic_echoes[label0]
 echo_1 = synthetic_echoes[label1]

### 2020-03-01 script_synthetic_echoes_to_PCA.py

@@ -14,9 +14,9 @@ pyplot.style.use(settings.style)
 # n_seed_levels = [5, 10, 20, 40, 80, 160, 320]
 # n_cloud_levels = [5, 10, 20, 40, 80, 160, 320]
 
-repeats = 10
-n_seed_levels = [1, 2, 5, 10, 250]
-n_cloud_levels = [1, 2, 5, 10, 250]
+repeats = 25
+n_seed_levels = [1, 2, 5, 10, 20, 250]
+n_cloud_levels = [1, 2, 5, 10, 20,  250]
 
 print('---> CREATE SYNTHETIC TEMPLATES')
 
@@ -53,6 +53,7 @@ for seed_level_i in range(len(n_seed_levels)):
 
             # Make echo sequence
             echo_sequence = generate_functions.distances2echo_sequence(distances, caller, emission)
+            #numpy.random.shuffle(echo_sequence)
 
             # Store echoes
             a = n_seed_levels[seed_level_i]

### 2020-03-01 Paper/main.tex

@@
 
-These limitations of the bat's sonar system suggest that sonar provides the bat with a sparse, low dimensional description of the outside 3D world.
-Recently, \citet{Geberl2019} confirmed that bio-sonar has a very low spatial resolution.
+We have suggested before that limitations of the bat's sonar system result in bats having access to a sparse, low dimensional description of the outside 3D world only \citep{Vanderelst2015a,Vanderelst2016}.
+Some recent evidence seems to confirm this.
+First, \citet{Geberl2019} confirmed that bio-sonar has a very low spatial resolution, both for distance and angular dimensions.
 In addition, neurophysiological data suggest that closely spaced objects are not perceived individually by echolocating bats.
-In contrast, neural responses to echo cascades from densely spaced objects represent a single extended stimulus \citep{Warnecke2018}.
+In contrast, neural responses to echo cascades from densely spaced objects seem to represent a single extended stimulus \citep{Warnecke2018}.
+\todo[inline]{Look at recent Simmons paper where bats refuse to fly through hoop}
 
-Nevertheless, low dimensional descriptions potentially contain a lot of information and can be sufficient for recognizing locations \citep{Kuc1997b,Kuc1997}.
+Low dimensional descriptions potentially contain a lot of information \citep{Kuc1997b,Kuc1997}.
 We previously demonstrated that bats might use sonar template-based scene recognition \citep{Vanderelst2016,Vanderelst2017}.
@@
 
-The finding that templates provide sufficient information to recognize locations prompted us to directly evaluate the hypothesis that sonar provides bats inherently with a low dimensional, yet sufficiently rich, description of the world.
-Testing this, we show that complex synthetic echo sequences can be mapped into a low dimensional space, after filtering them through a model of the bat's auditory periphery.
-Next, we show that templates derived from real bat habitats can also be projected into this space.
-Finally, we demonstrate that low dimensional feature space allows recognizing locations in real habitats.
+In this paper, we directly evaluate the hypothesis that sonar provides bats inherently with a low dimensional, yet sufficiently rich, description of the world.
+We test this hypothesis in two steps.
+In a first step, we generate a large collection of synthetic echo sequences.
+We filter these sequences using a model of the bat's auditory periphery \citep{Wiegrebe2008}.
+The synthetic echo sequences differ greatly in complexity.
+Nevertheless, we show that their resulting cochlear output can be mapped into a lower dimensional space.
+In a second step, we show that echoes collected in real bat habitats can be mapped into the same lower dimensional space, without loss of information.
+We do this by demonstrating that the low dimensional feature space allows recognizing locations in real habitats.
 
-Together these results suggest that echolocating bats live in a low dimensional world.
-The sonar system and the auditory periphery map the complexity of the 3D environment onto a small number of dimensions.
-Nevertheless, these low dimensional descriptions are sufficiently rich for bats to recognize locations.
+Together these results suggest that echolocating bats live in a low dimensional perceptual world.
+The physics of the sonar system and the auditory periphery map the complexity of the 3D environment onto a small number of features, irrespective of whether we use artificial or real echo sequences.
+Nevertheless, these features are sufficiently rich for bats to recognize locations, and as we discuss, support flight control.
 
@@
 We generated xx sets of points for each combination of $n_s \in [xx,xx,xx, xx]$ and $n_c \in [xx, xx, xx, xx]$.
+Synthetic echoes consisting of few reflectors are likely to consist of only weak echoes.
+To ensure that all sequences included at least one strong echo, we added a single reflector at a distance between xx and xx to each collection of points.
 
@@
 The components are visualized in figure \ref{fig:components}.
+
+\todo[inline]{Should we talk about the fact that the PCA model fits well across templates generated with different number of reflectors?
+This seems to depend on the echo having at least one strong (aka nearby) reflector.}
 

### 2020-03-01 library/process_functions.py
Removed subsampling from preprocessing

@@ -17,14 +17,14 @@ def preprocess(file_name, verbose=True):
 
     initial_omission_samples = math.ceil(initial_zero_time * fs)
     integration_samples = math.ceil(fs * integration_time)
-    final_samples = math.ceil(raw_collected_samples / integration_samples)
+    #final_samples = math.ceil(raw_collected_samples / integration_samples)
 
     if verbose:
         print('+-' * 10 + '+')
         print('PREPROCESSING', file_name)
-        print('integration samples', integration_samples)
+        #print('integration samples', integration_samples)
         print('initial zero samples', initial_omission_samples)
-        print('final samples', final_samples)
+        #print('final samples', final_samples)
 
     data = io.loadmat(file_name)
     data = data['Templates']
@@ -32,7 +32,8 @@ def preprocess(file_name, verbose=True):
     azimuth = data[0, 0][1]
     elevation = data[0, 0][2]
 
-    processed = numpy.zeros((7, 31, final_samples, 3))
+    #processed = numpy.zeros((7, 31, final_samples, 3))
+    processed = numpy.zeros((7, 31, raw_collected_samples, 3))
 
     for i in range(3):
         repetition = templates[:, :, i]
@@ -54,12 +55,21 @@ def preprocess(file_name, verbose=True):
         el_box = el_box[numpy.ix_(row_indices, col_indices)]
         mn_box = mn_box[numpy.ix_(row_indices, col_indices)]
 
-        # Average across directions and time
-        mask = numpy.ones((3, 3, integration_samples))
+        # Average across directions
+        mask = numpy.ones((3, 3, 1))
         mask = mask / numpy.sum(mask)
         mn_box = convolve(mn_box, mask, mode='same')
-        # Subsample
-        mn_box = mn_box[:, :, ::integration_samples]
+
+        # # Average across directions and time
+        # mask = numpy.ones((3, 3, integration_samples))
+        # mask = mask / numpy.sum(mask)
+        # mn_box = convolve(mn_box, mask, mode='same')
+        # # Subsample
+        # mn_box = mn_box[:, :, ::integration_samples]
+
+        if verbose: print('final samples', mn_box.shape[2])
+
+
         processed[:, :, :, i] = mn_box
         processed[processed < noise_floor] = noise_floor
     return processed, az_box, el_box
@@ -94,7 +104,6 @@ def process_data_set(data_set, filter_threshold = 0.1):
[…]

### 2020-03-01 script_run_training.py

@@ -9,84 +9,81 @@ from library import misc, process_functions, settings
 data_set = 'royal'
 
 
-generate_data = False
-do_training = True
 layers = [50, 75, 75, 50]
 nr_epochs = 5000
 
 for selected_dimension in ['lcs', 'azs','els']:
     file_names = misc.folder_names(data_set, selected_dimension)
-    if generate_data: process_functions.process_data_set(data_set)
-    if do_training:
-        # Read prepared data
-        data = numpy.load(file_names['npz_file'])
-        pca_model = misc.pickle_load(settings.pca_file)
 
-        # Select inputs
-        if selected_dimension == 'lcs': unencoded_data = data['long_lcs']
-        if selected_dimension == 'azs': unencoded_data = data['long_azs']
-        if selected_dimension == 'els': unencoded_data = data['long_els']
-        if selected_dimension == 'lcs' and data_set in ['israel', 'royal']: unencoded_data = misc.map_lcs_to_distances(data)
+    # Read prepared data
+    data = numpy.load(file_names['npz_file'])
+    pca_model = misc.pickle_load(settings.pca_file)
 
-        # Encode targets
-        encoder, targets = process_functions.get_encoding(unencoded_data)
-        target_n = targets.shape[1]
+    # Select inputs
+    if selected_dimension == 'lcs': unencoded_data = data['long_lcs']
+    if selected_dimension == 'azs': unencoded_data = data['long_azs']
+    if selected_dimension == 'els': unencoded_data = data['long_els']
+    if selected_dimension == 'lcs' and data_set in ['israel', 'royal']: unencoded_data = misc.map_lcs_to_distances(data)
 
-        # Get PCA-ed template inputs
-        templates = data['long_data']
-        pca_templates = pca_model.transform(templates)
-        n_components = settings.n_components
-        inputs = pca_templates[:, :n_components]
+    # Encode targets
+    encoder, targets = process_functions.get_encoding(unencoded_data)
+    target_n = targets.shape[1]
 
-        # Scale inputs to a minimum of zero
-        inputs = inputs - numpy.min(inputs)
+    # Get PCA-ed template inputs
+    templates = data['long_data']
+    pca_templates = pca_model.transform(templates)
+    n_components = settings.n_components
+    inputs = pca_templates[:, :n_components]
 
-        # Make model
-        model = keras.Sequential()
[…]

### 2020-03-01 script_generate_echoes.py
removed subsampling from synthetic template generation.

@@ -1,76 +1,0 @@
-import math
-from matplotlib import pyplot
-import numpy
-
-from library import generate_functions
-from library import misc
-from library import settings
-from pyBat import Call
-from pyBat import Wiegrebe
-
-pyplot.style.use(settings.style)
-
-repeats = 10
-
-sample_frequency = settings.sample_frequency
-number_of_samples = settings.raw_collected_samples
-initial_zero_time = settings.initial_zero_time
-emission_frequency_mean = settings.emission_frequency_mean
-
-integration_samples = math.ceil(sample_frequency * settings.integration_time)
-
-caller = Call.Call(None, [emission_frequency_mean])
-wiegrebe = Wiegrebe.ModelWiegrebe(sample_frequency, emission_frequency_mean, 4)
-pca_model = misc.pickle_load(settings.pca_file)
-emission = generate_functions.create_emission()
-
-all_fits = numpy.zeros((3, 3))
-all_inverse_fits = numpy.zeros((3, 3))
-
-n_seed_levels = [50, 100, 150]
-n_cloud_levels = [50, 100, 150]
-
-for seed_level_i in range(len(n_seed_levels)):
-    for cloud_level_i in range(len(n_cloud_levels)):
-        fits = []
-        inverse_fits = []
-        print(seed_level_i, cloud_level_i)
-        for i in range(repeats):
-            # Make distances
-            generative_model_parameters = {}
-            generative_model_parameters['n_seed_points'] = n_seed_levels[seed_level_i]
-            generative_model_parameters['n_cloud_points'] = n_cloud_levels[cloud_level_i]
-            distances = generate_functions.generative_model(generative_model_parameters)
-
-            # Make echo sequence
-            echo_sequence = generate_functions.distances2echo_sequence(distances, caller, emission)
-
-            # Create template
-            template = generate_functions.echo_sequence2template(echo_sequence, wiegrebe)
-            inverse_template = numpy.flip(template)
-
-            # Evaluate
-            template, reconstructed, explained_variance = generate_functions.evaluate_fit(template, pca_model)
-            fits.append(explained_variance)
-
-            inverse_template, inverse_reconstructed, explained_variance = generate_functions.evaluate_fit(inverse_template, pca_model)
-            inverse_fits.append(explained_variance)
-
[…]

### 2020-03-01 script_synthetic_echoes_to_PCA.py

@@ -15,8 +15,8 @@ pyplot.style.use(settings.style)
 # n_cloud_levels = [5, 10, 20, 40, 80, 160, 320]
 
 repeats = 25
-n_seed_levels = [1, 2, 5, 10, 20, 250]
-n_cloud_levels = [1, 2, 5, 10, 20,  250]
+n_seed_levels = [1, 2]
+n_cloud_levels = [1, 2]
 
 print('---> CREATE SYNTHETIC TEMPLATES')
 

### 2020-03-01 script_run_PCAs.py
deleted script_run_pcas
older script, not needed anymore.

@@ -1,81 +1,0 @@
-import numpy
-import os
-from matplotlib import pyplot
-from sklearn.decomposition import PCA
-
-from library import misc, process_functions, settings
-
-generate_data_israel = False
-generate_data_royal = False
-generate_data_andrews = False
-
-pyplot.style.use(settings.style)
-pca_plot_file = os.path.join(settings.figure_folder, 'pca.pdf')
-
-if generate_data_israel: process_functions.process_data_set('israel')
-if generate_data_royal: process_functions.process_data_set('royal')
-if generate_data_andrews: process_functions.process_data_set('andrews')
-
-data_set = 'israel'
-file_names = misc.folder_names(data_set, None)
-data = numpy.load(file_names['npz_file'])
-templates_isreael = data['long_data']
-
-data_set = 'royal'
-file_names = misc.folder_names(data_set, None)
-data = numpy.load(file_names['npz_file'])
-templates_royal = data['long_data']
-
-data_set = 'andrews'
-file_names = misc.folder_names(data_set, None)
-data = numpy.load(file_names['npz_file'])
-templates_andrews = data['long_data']
-
-all_templates = numpy.row_stack((templates_royal, templates_isreael))
-print('---> RUN AND SAVE PCA MODEL')
-pca_model = PCA(n_components=settings.n_components)
-pca_model.fit(all_templates)
-misc.pickle_save(settings.pca_file, pca_model)
-
-
-transformed = pca_model.transform(all_templates)
-reconstructed = pca_model.inverse_transform(transformed)
-ratio = numpy.cumsum(pca_model.explained_variance_ratio_)
-
-x = reconstructed.flatten()
-y = all_templates.flatten()
-r0 = numpy.corrcoef(x, y)[0, 1]
-
-# map andrews data
-
-transformed_andrews = pca_model.transform(templates_andrews)
-reconstructed_andrews = pca_model.inverse_transform(transformed_andrews)
-
-x = reconstructed_andrews.flatten()
-y = templates_andrews.flatten()
-r1 = numpy.corrcoef(x, y)[0, 1]
-
-print(r0 ** 2, r1 ** 2)
-
-# %% plots
-# Plot Explained variance
-pyplot.plot(ratio)
-pyplot.xlabel('Nr of PCs')
-pyplot.ylabel('Proportion of explained variance')
-pyplot.tight_layout()
-pyplot.savefig(pca_plot_file)
[…]

### 2020-03-02 library/settings.py
First version of this approach

@@ -18,7 +18,7 @@ noise_floor = 0.18 # should be determined on open space
 stochaistic_noise = 0.02 # stdv, See paper for determination, before pca mapping
 integration_time = 350 / 1000000  # integration constant
 initial_zero_time = 6 / 1000  # initial time to omit
-n_components = 20
+n_components = 100
 ####################################################
 ##Generative modelling settings
 ####################################################
@@ -27,13 +27,17 @@ emission_frequency_low = 40000
 emission_frequency_high = 100000
 emission_frequency_mean = round((emission_frequency_low + emission_frequency_high) / 2)
 
-synthetic_echoes_file = os.path.join(result_folder, 'synthetic_echoes.pck')
-synthetic_echoes_indices = os.path.join(result_folder, 'synthetic_echoes_indices.pck')
-synthetic_templates_file = os.path.join(result_folder, 'synthetic_templates.npy')
-reconstructed_synthetic_templates_file = os.path.join(result_folder, 'reconstructed_synthetic_templates.npy')
-scaled_synthetic_templates_file = os.path.join(result_folder, 'scaled_synthetic_templates.npy')
+#synthetic_echoes_file = os.path.join(result_folder, 'synthetic_echoes.pck')
+#synthetic_echoes_indices = os.path.join(result_folder, 'synthetic_echoes_indices.pck')
 
-reconstructed_templates_file =  os.path.join(result_folder, 'reconstructed_templates.npy')
+synthetic_echoes_file = os.path.join(result_folder, 'synthetic_echoes.npz')
+synthetic_templates_file = os.path.join(result_folder, 'synthetic_templates.npz')
+scaled_synthetic_templates_file = os.path.join(result_folder, 'scaled_synthetic_templates.npz')
+reconstructed_synthetic_templates_file = os.path.join(result_folder, 'reconstructed_synthetic_templates.npz')
+
+#Dont save this, as it results in a big file that takes ages to sync.
+#reconstructed_templates_file =  os.path.join(result_folder, 'reconstructed_templates.npz')
+
 
 pca_file = os.path.join(result_folder, 'PCA.pca')
 pca_components_plot = os.path.join(result_folder, 'components.pdf')

### 2020-03-02 script_plot_components.py

@@ -13,8 +13,6 @@ n = settings.n_components
 pca_model = misc.pickle_load(settings.pca_file)
 components = pca_model.components_
 
-
-
 mx = numpy.max(numpy.abs(components))
 n = components.shape[1]
 x = numpy.array(range(n))

### 2020-03-02 script_synthetic_echoes_to_PCA.py

@@ -1,7 +1,4 @@
-import math
 import numpy
-
-from library import misc
 from matplotlib import pyplot
 from library import generate_functions
 from library import settings
@@ -10,13 +7,7 @@ from pyBat import Wiegrebe
 
 pyplot.style.use(settings.style)
 
-# repeats = 30
-# n_seed_levels = [5, 10, 20, 40, 80, 160, 320]
-# n_cloud_levels = [5, 10, 20, 40, 80, 160, 320]
-
-repeats = 25
-n_seed_levels = [1, 2]
-n_cloud_levels = [1, 2]
+reflector_distances = numpy.linspace(1,6, 500)
 
 print('---> CREATE SYNTHETIC TEMPLATES')
 
@@ -24,8 +15,6 @@ sample_frequency = settings.sample_frequency
 number_of_samples = settings.raw_collected_samples
 initial_zero_time = settings.initial_zero_time
 emission_frequency_mean = settings.emission_frequency_mean
-
-integration_samples = math.ceil(sample_frequency * settings.integration_time)
 
 caller = Call.Call(None, [emission_frequency_mean])
 emission = generate_functions.create_emission()
@@ -36,50 +25,35 @@ padded_emission = numpy.pad(emission, (0, emission_padding), 'constant')
 wiegrebe = Wiegrebe.ModelWiegrebe(sample_frequency, emission_frequency_mean, 4, emission=padded_emission)
 
 templates_synthetic = []
-echoes_synthetic = {}
-echoes_synthetic_indices = {}
-index = 0
-for seed_level_i in range(len(n_seed_levels)):
-    for cloud_level_i in range(len(n_cloud_levels)):
-        fits = []
-        inverse_fits = []
-        print(seed_level_i, cloud_level_i)
-        for i in range(repeats):
-            # Make distances
-            generative_model_parameters = {}
-            generative_model_parameters['n_seed_points'] = n_seed_levels[seed_level_i]
-            generative_model_parameters['n_cloud_points'] = n_cloud_levels[cloud_level_i]
-            distances = generate_functions.generative_model(generative_model_parameters)
+echoes_synthetic = []
 
-            # Make echo sequence
-            echo_sequence = generate_functions.distances2echo_sequence(distances, caller, emission)
-            #numpy.random.shuffle(echo_sequence)
+for reflector_distance in reflector_distances:
+    print('Current distance:', reflector_distance)
+    # Make echo sequence
[…]

### 2020-03-02 PCA_syn_echoes_TTTEST.py
before clearing some files

@@ -1,0 +1,163 @@
+from scipy.signal import hilbert
+import math
+import numpy
+import os
+from matplotlib import pyplot
+from library import misc, settings
+from sklearn.decomposition import PCA
+from library import generate_functions
+from pyBat import Call
+from pyBat import Wiegrebe
+
+repeats = 10
+
+n_seed_levels = [5, 10, 20, 50, 100, 200]
+n_cloud_levels = [5, 10, 20, 50, 100, 200]
+
+sample_frequency = settings.sample_frequency
+number_of_samples = settings.raw_collected_samples
+initial_zero_time = settings.initial_zero_time
+emission_frequency_mean = settings.emission_frequency_mean
+initial_zero_samples = math.ceil(initial_zero_time * sample_frequency)
+
+caller = Call.Call(None, [emission_frequency_mean])
+emission = generate_functions.create_emission()
+
+emission_padding = number_of_samples - emission.shape[0]
+padded_emission = numpy.pad(emission, (0, emission_padding), 'constant')
+
+wiegrebe = Wiegrebe.ModelWiegrebe(sample_frequency, emission_frequency_mean, 4, emission=padded_emission)
+
+synthetic_echoes = []
+synthetic_templates = []
+
+for seed_level_i in range(len(n_seed_levels)):
+    for cloud_level_i in range(len(n_cloud_levels)):
+        print(seed_level_i, cloud_level_i)
+        for i in range(repeats):
+            generative_model_parameters = {}
+            generative_model_parameters['n_seed_points'] = n_seed_levels[seed_level_i]
+            generative_model_parameters['n_cloud_points'] = n_cloud_levels[cloud_level_i]
+            distances = generate_functions.generative_model(generative_model_parameters)
+            echo_sequence = generate_functions.distances2echo_sequence(distances, caller, emission)
+            echo_sequence[0:initial_zero_samples] = 0
+            template = generate_functions.echo_sequence2template(echo_sequence, wiegrebe)
+            synthetic_echoes.append(echo_sequence)
+            synthetic_templates.append(template)
+
+print('---> SAVE ECHOES AND TEMPLATES')
+
+synthetic_echoes = numpy.array(synthetic_echoes)
[…]

### 2020-03-02 library/generate_functions.py

@@ -39,9 +39,9 @@ def generative_model(parameters):
         distances = numpy.concatenate((distances, g))
 
     # add close echo to ensure that even seqs with low n have a strong echo
-    close = random.uniform(minimum_distance, minimum_distance+1, 1)
-    g = numpy.array([minimum_distance])
-    distances = numpy.concatenate((distances, close))
+    #close = random.uniform(minimum_distance, minimum_distance+1, 1)
+    #g = numpy.array([minimum_distance])
+    #distances = numpy.concatenate((distances, close))
     return distances
 
 def distances2echo_sequence(distances, caller, emission):
@@ -74,7 +74,7 @@ def distances2echo_sequence(distances, caller, emission):
 def echo_sequence2template(echo_sequence, wiegrebe):
     sample_frequency = settings.sample_frequency
     integration_time = settings.integration_time
-    integration_samples = math.ceil(sample_frequency * integration_time)
+    #integration_samples = math.ceil(sample_frequency * integration_time)
 
     # Run Wiegrebe model
     wiegrebe_result = wiegrebe.run_model(echo_sequence, dechirp=True)

### 2020-03-02 library/misc.py

@@ -7,6 +7,20 @@ import scipy.interpolate as interpolate
 from matplotlib import pyplot
 from sklearn.preprocessing import Normalizer
 
+def load_all_templates():
+    data_set = 'israel'
+    file_names = folder_names(data_set, None)
+    data = numpy.load(file_names['npz_file'])
+    template_israel = data['long_data']
+
+    data_set = 'royal'
+    file_names = folder_names(data_set, None)
+    data = numpy.load(file_names['npz_file'])
+    templates_royal = data['long_data']
+
+    all_templates = numpy.row_stack((templates_royal, template_israel))
+    return all_templates
+
 
 def map_lcs_to_distances(data):
     distances = extract_distances_from_filenames(data['files'])

### 2020-03-02 library/settings.py

@@ -18,13 +18,13 @@ noise_floor = 0.18 # should be determined on open space
 stochaistic_noise = 0.02 # stdv, See paper for determination, before pca mapping
 integration_time = 350 / 1000000  # integration constant
 initial_zero_time = 6 / 1000  # initial time to omit
-n_components = 100
+n_components = 75
 ####################################################
 ##Generative modelling settings
 ####################################################
 emission_duration = 0.001
-emission_frequency_low = 40000
-emission_frequency_high = 100000
+emission_frequency_low = 50000
+emission_frequency_high = 50001
 emission_frequency_mean = round((emission_frequency_low + emission_frequency_high) / 2)
 
 #synthetic_echoes_file = os.path.join(result_folder, 'synthetic_echoes.pck')

### 2020-03-02 script_plot_template_examples.py

@@ -7,12 +7,25 @@ from library import misc, settings
 # Synthetic echoes, templates and reconstruction
 #
 
+mx_echoes = 0.05
+mx_templates = 3.5
+mn_templates = 0.15
+
+index0 = 10
+index1 = 50
+index2 = 400
+
 pyplot.style.use(settings.style)
 
-synthetic_echoes = misc.pickle_load(settings.synthetic_echoes_file)
-synthetic_indices = misc.pickle_load(settings.synthetic_echoes_indices)
-synthetic_templates = numpy.load(settings.scaled_synthetic_templates_file)
-reconstructed_synthetic_templates = numpy.load(settings.reconstructed_synthetic_templates_file)
+print('-----> LOADING DATA')
+loaded = numpy.load(settings.synthetic_echoes_file)
+synthetic_echoes = loaded['echoes_synthetic']
+
+loaded = numpy.load(settings.scaled_synthetic_templates_file)
+synthetic_templates = loaded['templates_synthetic']
+
+loaded = numpy.load(settings.reconstructed_synthetic_templates_file)
+reconstructed_synthetic_templates = loaded['reconstructed_synthetic']
 
 pca_model = misc.pickle_load(settings.pca_file)
 cummulative_variance = numpy.cumsum(pca_model.explained_variance_ratio_)
@@ -20,21 +33,11 @@ cummulative_variance = numpy.cumsum(pca_model.explained_variance_ratio_)
 x = synthetic_templates.flatten()
 y = reconstructed_synthetic_templates.flatten()
 
-mx_echoes = 0.25
-mx_templates = 0.35
-mn_templates = 0.15
+print('-----> GETTING DATA FROM ARRAYS')
 
-label0 = '5_5_0'
-label1 = '10_10_0'
-label2 = '250_250_0'
-
-echo_0 = synthetic_echoes[label0]
-echo_1 = synthetic_echoes[label1]
-echo_2 = synthetic_echoes[label2]
-
-index0 = synthetic_indices[label0]
-index1 = synthetic_indices[label1]
-index2 = synthetic_indices[label2]
+echo_0 = synthetic_echoes[index0]
+echo_1 = synthetic_echoes[index1]
+echo_2 = synthetic_echoes[index2]
 
 template_0 = synthetic_templates[index0, :]
[…]

### 2020-03-03 PCA_syn_echoes_TTTEST.py
before wrapping some stuff in functions

@@ -1,163 +1,0 @@
-from scipy.signal import hilbert
-import math
-import numpy
-import os
-from matplotlib import pyplot
-from library import misc, settings
-from sklearn.decomposition import PCA
-from library import generate_functions
-from pyBat import Call
-from pyBat import Wiegrebe
-
-repeats = 10
-
-n_seed_levels = [5, 10, 20, 50, 100, 200]
-n_cloud_levels = [5, 10, 20, 50, 100, 200]
-
-sample_frequency = settings.sample_frequency
-number_of_samples = settings.raw_collected_samples
-initial_zero_time = settings.initial_zero_time
-emission_frequency_mean = settings.emission_frequency_mean
-initial_zero_samples = math.ceil(initial_zero_time * sample_frequency)
-
-caller = Call.Call(None, [emission_frequency_mean])
-emission = generate_functions.create_emission()
-
-emission_padding = number_of_samples - emission.shape[0]
-padded_emission = numpy.pad(emission, (0, emission_padding), 'constant')
-
-wiegrebe = Wiegrebe.ModelWiegrebe(sample_frequency, emission_frequency_mean, 4, emission=padded_emission)
-
-synthetic_echoes = []
-synthetic_templates = []
-
-for seed_level_i in range(len(n_seed_levels)):
-    for cloud_level_i in range(len(n_cloud_levels)):
-        print(seed_level_i, cloud_level_i)
-        for i in range(repeats):
-            generative_model_parameters = {}
-            generative_model_parameters['n_seed_points'] = n_seed_levels[seed_level_i]
-            generative_model_parameters['n_cloud_points'] = n_cloud_levels[cloud_level_i]
-            distances = generate_functions.generative_model(generative_model_parameters)
-            echo_sequence = generate_functions.distances2echo_sequence(distances, caller, emission)
-            echo_sequence[0:initial_zero_samples] = 0
-            template = generate_functions.echo_sequence2template(echo_sequence, wiegrebe)
-            synthetic_echoes.append(echo_sequence)
-            synthetic_templates.append(template)
-
-print('---> SAVE ECHOES AND TEMPLATES')
-
-synthetic_echoes = numpy.array(synthetic_echoes)
[…]

### 2020-03-03 fit.py

@@ -9,7 +9,7 @@ echoes = misc.pickle_load(settings.synthetic_echoes_file)
 indices = misc.pickle_load(settings.synthetic_echoes_indices)
 templates_synthetic = numpy.load(settings.synthetic_templates_file)
 
-pca_model = misc.pickle_load(settings.pca_file)
+pca_model = misc.pickle_load(settings.pca_templates_file)
 
 
 transformed_synthetic = pca_model.fit_transform(templates_synthetic)

### 2020-03-03 library/generate_functions.py

@@ -68,7 +68,7 @@ def distances2echo_sequence(distances, caller, emission):
     # Generate echo sequency
     echo_sequence = numpy.convolve(emission, impulse_response, mode='same')
     echo_sequence[0:number_of_zero_samples] = 0
-    return echo_sequence
+    return echo_sequence, impulse_response
 
 
 def echo_sequence2template(echo_sequence, wiegrebe):

### 2020-03-03 library/misc.py

@@ -7,6 +7,35 @@ import scipy.interpolate as interpolate
 from matplotlib import pyplot
 from sklearn.preprocessing import Normalizer
 
+from sklearn.decomposition import PCA
+
+
+def corr2(a, b):
+    n = a.shape()[0]
+    correlations = numpy.zerosn()
+    for i in range(n):
+        x = a[i, :]
+        y = b[i, :]
+        r = numpy.corrcoef(x, y)[0, 1]
+        correlations[i] = r
+    correlation = numpy.mean(correlations)
+    return correlation
+
+
+def run_pca(data, criterion=0.99):
+    # Run all components
+    pca_model = PCA()
+    pca_model.fit(data)
+    cummmulative_echoes = numpy.cumsum(pca_model.explained_variance_ratio_)
+    suggestion = numpy.min(numpy.where(cummmulative_echoes > criterion)[0])
+
+    # Run with n components for criterion
+    pca_model = PCA(n_components=suggestion)
+    transformed = pca_model.fit_transform(data)
+    reconstructed = pca_model.inverse_transform(transformed)
+    correlations = corr2(data, reconstructed)
+
+
 def load_all_templates():
     data_set = 'israel'
     file_names = folder_names(data_set, None)
@@ -74,7 +103,7 @@ def folder_names(data_set, dimension):
     result['result_folder'] = result_folder
     result['log_folder'] = log_folder
     result['npz_file'] = npz_file
-    #call_result['pca_file'] = pca_file
+    # call_result['pca_file'] = pca_file
     result['model_file'] = model_file
     result['results_file'] = res_file
     result['history_file'] = hist_file
@@ -96,7 +125,6 @@ def pickle_load(file):
     return object
 
 
-
 def make_confusion_matrix(results, normalize=True):
     N = Normalizer()
     values = results.target.unique()
@@ -109,7 +137,7 @@ def make_confusion_matrix(results, normalize=True):
             pv = values[pi]
             selected = results.query("target==@tv and prediction==@pv")
             selected_n = selected.shape[0]
-            #print(selected_n)
+            # print(selected_n)
             table[ti, pi] = selected_n
     labels = values
     table[numpy.isnan(table)] = 0
@@ -125,16 +153,15 @@ def label_confusion_matrix(labels):
[…]

### 2020-03-03 library/settings.py

@@ -18,28 +18,33 @@ noise_floor = 0.18 # should be determined on open space
 stochaistic_noise = 0.02 # stdv, See paper for determination, before pca mapping
 integration_time = 350 / 1000000  # integration constant
 initial_zero_time = 6 / 1000  # initial time to omit
-n_components = 75
+n_components_templates = 210
+n_components_echoes = 210
 ####################################################
 ##Generative modelling settings
 ####################################################
 emission_duration = 0.001
-emission_frequency_low = 50000
-emission_frequency_high = 50001
+emission_frequency_low = 100000
+emission_frequency_high = 40000
 emission_frequency_mean = round((emission_frequency_low + emission_frequency_high) / 2)
 
 #synthetic_echoes_file = os.path.join(result_folder, 'synthetic_echoes.pck')
 #synthetic_echoes_indices = os.path.join(result_folder, 'synthetic_echoes_indices.pck')
 
+synthetic_impulses_file = os.path.join(result_folder, 'synthetic_impulses.npz')
 synthetic_echoes_file = os.path.join(result_folder, 'synthetic_echoes.npz')
 synthetic_templates_file = os.path.join(result_folder, 'synthetic_templates.npz')
 scaled_synthetic_templates_file = os.path.join(result_folder, 'scaled_synthetic_templates.npz')
+
 reconstructed_synthetic_templates_file = os.path.join(result_folder, 'reconstructed_synthetic_templates.npz')
+reconstructed_synthetic_echoes_file = os.path.join(result_folder, 'reconstructed_synthetic_echoes.npz')
 
 #Dont save this, as it results in a big file that takes ages to sync.
 #reconstructed_templates_file =  os.path.join(result_folder, 'reconstructed_templates.npz')
 
+pca_templates_file = os.path.join(result_folder, 'PCA_templates.pca')
+pca_echoes_file = os.path.join(result_folder, 'PCA_echoes.pca')
 
-pca_file = os.path.join(result_folder, 'PCA.pca')
 pca_components_plot = os.path.join(result_folder, 'components.pdf')
 
 synthetic_templates_plot = os.path.join(result_folder, 'synthetic_templates.pdf')

### 2020-03-03 script_plot_components.py

@@ -8,9 +8,9 @@ from library import misc, settings
 from PyAstronomy.pyaC import zerocross1d
 
 pyplot.style.use(settings.style)
-n = settings.n_components
+n = settings.n_components_templates
 
-pca_model = misc.pickle_load(settings.pca_file)
+pca_model = misc.pickle_load(settings.pca_templates_file)
 components = pca_model.components_
 
 mx = numpy.max(numpy.abs(components))

### 2020-03-03 script_plot_template_examples.py

@@ -7,31 +7,35 @@ from library import misc, settings
 # Synthetic echoes, templates and reconstruction
 #
 
-mx_echoes = 0.05
-mx_templates = 3.5
-mn_templates = 0.15
+mx_echoes = 0.1
+mx_templates = 0.25
+mn_templates = 0.17
 
 index0 = 10
-index1 = 50
-index2 = 400
+index1 = 100
+index2 = 250
 
 pyplot.style.use(settings.style)
 
 print('-----> LOADING DATA')
 loaded = numpy.load(settings.synthetic_echoes_file)
-synthetic_echoes = loaded['echoes_synthetic']
+synthetic_echoes = loaded['synthetic_echoes']
 
 loaded = numpy.load(settings.scaled_synthetic_templates_file)
-synthetic_templates = loaded['templates_synthetic']
+synthetic_templates = loaded['synthetic_templates']
 
 loaded = numpy.load(settings.reconstructed_synthetic_templates_file)
-reconstructed_synthetic_templates = loaded['reconstructed_synthetic']
+reconstructed_synthetic_templates = loaded['reconstructed_synthetic_templates']
 
-pca_model = misc.pickle_load(settings.pca_file)
+loaded = numpy.load(settings.reconstructed_synthetic_echoes_file)
+reconstructed_synthetic_echoes = loaded['reconstructed_synthetic_echoes']
+
+
+pca_model = misc.pickle_load(settings.pca_templates_file)
 cummulative_variance = numpy.cumsum(pca_model.explained_variance_ratio_)
 
-x = synthetic_templates.flatten()
-y = reconstructed_synthetic_templates.flatten()
+#x = synthetic_templates.flatten()
+#y = reconstructed_synthetic_templates.flatten()
 
 print('-----> GETTING DATA FROM ARRAYS')
 
@@ -42,6 +46,10 @@ echo_2 = synthetic_echoes[index2]
 template_0 = synthetic_templates[index0, :]
 template_1 = synthetic_templates[index1, :]
 template_2 = synthetic_templates[index2, :]
+
+reconstructed_echo_0 = reconstructed_synthetic_echoes[index0, :]
+reconstructed_echo_1 = reconstructed_synthetic_echoes[index1, :]
[…]

### 2020-03-03 script_run_perfect_memory.py

@@ -25,7 +25,7 @@ correct_ids[:, 0] = corridor_distances
 
 templates = data['long_data']
 pca_templates = pca.transform(templates)
-n_components = settings.n_components
+n_components = settings.n_components_templates
 inputs = pca_templates[:, :n_components]
 
 # Scale inputs to a minimum of zero

### 2020-03-03 script_run_training.py

@@ -17,7 +17,7 @@ for selected_dimension in ['lcs', 'azs','els']:
 
     # Read prepared data
     data = numpy.load(file_names['npz_file'])
-    pca_model = misc.pickle_load(settings.pca_file)
+    pca_model = misc.pickle_load(settings.pca_templates_file)
 
     # Select inputs
     if selected_dimension == 'lcs': unencoded_data = data['long_lcs']
@@ -32,7 +32,7 @@ for selected_dimension in ['lcs', 'azs','els']:
     # Get PCA-ed template inputs
     templates = data['long_data']
     pca_templates = pca_model.transform(templates)
-    n_components = settings.n_components
+    n_components = settings.n_components_templates
     inputs = pca_templates[:, :n_components]
 
     # Scale inputs to a minimum of zero

### 2020-03-03 script_synthetic_echoes_to_PCA.py

@@ -1,81 +1,83 @@
+import math
 import numpy
-from matplotlib import pyplot
-from library import generate_functions
 from library import settings
+
+from library import generate_functions
 from pyBat import Call
 from pyBat import Wiegrebe
 
-pyplot.style.use(settings.style)
+repeats = 10
 
-reflector_distances = numpy.linspace(1,6, 500)
-
-print('---> CREATE SYNTHETIC TEMPLATES')
+n_seed_levels = [5, 10, 20, 50, 100, 200]
+n_cloud_levels = [5, 10, 20, 50, 100, 200]
 
 sample_frequency = settings.sample_frequency
 number_of_samples = settings.raw_collected_samples
 initial_zero_time = settings.initial_zero_time
 emission_frequency_mean = settings.emission_frequency_mean
+initial_zero_samples = math.ceil(initial_zero_time * sample_frequency)
 
 caller = Call.Call(None, [emission_frequency_mean])
 emission = generate_functions.create_emission()
 
-emission_padding = settings.raw_collected_samples - emission.shape[0]
+emission_padding = number_of_samples - emission.shape[0]
 padded_emission = numpy.pad(emission, (0, emission_padding), 'constant')
 
 wiegrebe = Wiegrebe.ModelWiegrebe(sample_frequency, emission_frequency_mean, 4, emission=padded_emission)
 
-templates_synthetic = []
-echoes_synthetic = []
+synthetic_impulses = []
+synthetic_echoes = []
+synthetic_templates = []
 
-for reflector_distance in reflector_distances:
-    print('Current distance:', reflector_distance)
-    # Make echo sequence
-    distances = numpy.array([reflector_distance])
-    echo_sequence = generate_functions.distances2echo_sequence(distances, caller, emission)
-
-    # Create template
-    template = generate_functions.echo_sequence2template(echo_sequence, wiegrebe)
-    templates_synthetic.append(template)
-    echoes_synthetic.append(echo_sequence)
+for seed_level_i in range(len(n_seed_levels)):
+    for cloud_level_i in range(len(n_cloud_levels)):
+        print(seed_level_i, cloud_level_i)
+        for i in range(repeats):
+            generative_model_parameters = {}
+            generative_model_parameters['n_seed_points'] = n_seed_levels[seed_level_i]
[…]

### 2020-03-03 script_tabulate_memory_use.py

@@ -16,7 +16,7 @@ for data_set in ['israel', 'royal']:
     data = numpy.load(file_names['npz_file'])
     templates = data['long_data']
 
-    total_n_numbers = templates.shape[0] * settings.n_components
+    total_n_numbers = templates.shape[0] * settings.n_components_templates
 
     new_model = keras.models.load_model(file_names['model_file'])
     weights = new_model.get_weights()

### 2020-03-03 library/misc.py
done wrapping

@@ -10,9 +10,39 @@ from sklearn.preprocessing import Normalizer
 from sklearn.decomposition import PCA
 
 
+def scale_synthetic_templates(empirical_templates, synthetic_templates):
+    # Plot means
+    mean_empirical = numpy.mean(empirical_templates, axis=0)
+    mean_synthetic = numpy.mean(synthetic_templates, axis=0)
+
+    pyplot.plot(mean_empirical)
+    pyplot.plot(mean_synthetic)
+    pyplot.legend(['Empirical', 'Synthetic'])
+    pyplot.show()
+
+    # Plot correct mean and offset
+    r_empirical = numpy.max(mean_empirical) - numpy.min(mean_empirical)
+    r_synthetic = numpy.max(mean_synthetic) - numpy.min(mean_synthetic)
+    scale = r_empirical / r_synthetic
+    offset = numpy.min(mean_empirical)
+    synthetic_templates = synthetic_templates * scale + offset
+
+    # Plot again
+    mean_empirical = numpy.mean(empirical_templates, axis=0)
+    mean_synthetic = numpy.mean(synthetic_templates, axis=0)
+
+    pyplot.plot(mean_empirical)
+    pyplot.plot(mean_synthetic)
+    pyplot.legend(['Empirical', 'Synthetic'])
+    pyplot.show()
+
+
+    return synthetic_templates
+
+
 def corr2(a, b):
-    n = a.shape()[0]
-    correlations = numpy.zerosn()
+    n = a.shape[0]
+    correlations = numpy.zeros(n)
     for i in range(n):
         x = a[i, :]
         y = b[i, :]
@@ -21,8 +51,15 @@ def corr2(a, b):
     correlation = numpy.mean(correlations)
     return correlation
 
+def project_and_reconstruct(pca_model, data):
+    transformed = pca_model.fit_transform(data)
+    reconstructed = pca_model.inverse_transform(transformed)
+    correlation = corr2(data, reconstructed)
+    return transformed, reconstructed, correlation
 
-def run_pca(data, criterion=0.99):
+
+
+def run_pca(data, criterion=0.99, save_file = None):
     # Run all components
     pca_model = PCA()
     pca_model.fit(data)
@@ -31,9 +68,22 @@ def run_pca(data, criterion=0.99):
 
     # Run with n components for criterion
     pca_model = PCA(n_components=suggestion)
-    transformed = pca_model.fit_transform(data)
-    reconstructed = pca_model.inverse_transform(transformed)
-    correlations = corr2(data, reconstructed)
[…]

### 2020-03-03 library/settings.py

@@ -28,27 +28,21 @@ emission_frequency_low = 100000
 emission_frequency_high = 40000
 emission_frequency_mean = round((emission_frequency_low + emission_frequency_high) / 2)
 
-#synthetic_echoes_file = os.path.join(result_folder, 'synthetic_echoes.pck')
-#synthetic_echoes_indices = os.path.join(result_folder, 'synthetic_echoes_indices.pck')
-
 synthetic_impulses_file = os.path.join(result_folder, 'synthetic_impulses.npz')
 synthetic_echoes_file = os.path.join(result_folder, 'synthetic_echoes.npz')
 synthetic_templates_file = os.path.join(result_folder, 'synthetic_templates.npz')
-scaled_synthetic_templates_file = os.path.join(result_folder, 'scaled_synthetic_templates.npz')
 
-reconstructed_synthetic_templates_file = os.path.join(result_folder, 'reconstructed_synthetic_templates.npz')
-reconstructed_synthetic_echoes_file = os.path.join(result_folder, 'reconstructed_synthetic_echoes.npz')
+synthetic_impulses_pca_results = os.path.join(result_folder, 'synthetic_impulses_pca.npz')
+synthetic_echoes_pca_results = os.path.join(result_folder, 'synthetic_echoes_pca.npz')
+synthetic_templates_pca_results = os.path.join(result_folder, 'synthetic_templates_pca.npz')
 
-#Dont save this, as it results in a big file that takes ages to sync.
-#reconstructed_templates_file =  os.path.join(result_folder, 'reconstructed_templates.npz')
+pca_impulses_model_file = os.path.join(result_folder, 'PCA_model_impulses.pca')
+pca_echoes_model_file = os.path.join(result_folder, 'PCA_model_echoes.pca')
+pca_templates_model_file = os.path.join(result_folder, 'PCA_model_templates.pca')
 
-pca_templates_file = os.path.join(result_folder, 'PCA_templates.pca')
-pca_echoes_file = os.path.join(result_folder, 'PCA_echoes.pca')
-
-pca_components_plot = os.path.join(result_folder, 'components.pdf')
-
-synthetic_templates_plot = os.path.join(result_folder, 'synthetic_templates.pdf')
-templates_plot = os.path.join(result_folder, 'templates.pdf')
+#pca_components_plot = os.path.join(result_folder, 'components.pdf')
+#synthetic_templates_plot = os.path.join(result_folder, 'synthetic_templates.pdf')
[…]

### 2020-03-03 script_synthetic_echoes_to_PCA.py

@@ -1,15 +1,19 @@
 import math
 import numpy
 from library import settings
-
+from library import misc
 from library import generate_functions
 from pyBat import Call
 from pyBat import Wiegrebe
 
 repeats = 10
-
 n_seed_levels = [5, 10, 20, 50, 100, 200]
 n_cloud_levels = [5, 10, 20, 50, 100, 200]
+
+#repeats = 3
+#n_seed_levels = [5, 10]
+#n_cloud_levels = [5, 10]
+
 
 sample_frequency = settings.sample_frequency
 number_of_samples = settings.raw_collected_samples
@@ -40,14 +44,20 @@ for seed_level_i in range(len(n_seed_levels)):
             echo_sequence, impulse_response = generate_functions.distances2echo_sequence(distances, caller, emission)
             echo_sequence[0:initial_zero_samples] = 0
             template = generate_functions.echo_sequence2template(echo_sequence, wiegrebe)
+
+            synthetic_impulses.append(impulse_response)
             synthetic_echoes.append(echo_sequence)
             synthetic_templates.append(template)
-
-print('---> SAVE ECHOES AND TEMPLATES')
 
 synthetic_impulses = numpy.array(synthetic_impulses)
 synthetic_echoes = numpy.array(synthetic_echoes)
 synthetic_templates = numpy.array(synthetic_templates)
+
+print('---> SCALE SYNTHETIC TEMPLATES')
+empirical_templates = misc.load_all_templates()
+synthetic_templates = misc.scale_synthetic_templates(empirical_templates, synthetic_templates)
+
+print('---> SAVE ECHOES AND TEMPLATES')
 
 numpy.savez_compressed(settings.synthetic_impulses_file, synthetic_impulses=synthetic_impulses)
 numpy.savez_compressed(settings.synthetic_echoes_file, synthetic_echoes=synthetic_echoes)
@@ -58,103 +68,36 @@ import importlib
 import numpy
 from library import settings
 from library import misc
-from matplotlib import pyplot
-
 
 importlib.reload(settings)
+importlib.reload(misc)
 
-print('-----> LOADING EMPIRICAL DATA')
-empirical_templates = misc.load_all_templates()
+print('---> LOADING SYNTHETIC DATA')
+loaded = numpy.load(settings.synthetic_impulses_file)
[…]

### 2020-03-11 Paper/main.tex
updates to paper

@@
 \usepackage{amssymb}
+\usepackage{acronym}
+\usepackage{glossaries}
 %\usepackage{lineno,hyperref}
@@
 \newcommand{\db}{dB}
+
+\newacronym{PCA}{PCA}{Principal Component Analysis}
+\newcommand{\PCA}{\gls{PCA}}
+
+\newacronym{pc}{PC}{principal component}
+\newcommand{\pc}{\gls{pc}} 
+\newcommand{\pcs}{\glspl{pc}}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 In this paper, we directly evaluate the hypothesis that sonar provides bats inherently with a low dimensional, yet sufficiently rich, description of the world.
+
 We test this hypothesis in two steps.
 In a first step, we generate a large collection of synthetic echo sequences.
-We filter these sequences using a model of the bat's auditory periphery \citep{Wiegrebe2008}.
+We assess the compressibility of these echo signals using \PCA.
+Next, we filter the echo sequences using a model of the bat's auditory periphery \citep{Wiegrebe2008}.
 The synthetic echo sequences differ greatly in complexity.
-Nevertheless, we show that their resulting cochlear output can be mapped into a lower dimensional space.
+Nevertheless, we show that they can be compressed considerably.
+Moreover, the output of the cochlear model can be mapped onto an even lower dimensional space.
 In a second step, we show that echoes collected in real bat habitats can be mapped into the same lower dimensional space, without loss of information.
@@
 
-\subsection{Synthetic echoes and templates}
+\subsection{Synthetic echoes}
 
-In this section of the paper, we derive a low dimensional space for describing templates.
+In this section of the paper, we asses how the filtering in the auditory periphery reduces the complexity of input signals.
 We do this by presenting a simple generative model for echo sequences received by bats.
-Next, we assess whether the templates derived from these synthetic echo sequences can be projected into a low dimensional space.
+Next, we assess whether the cochlear representation derived from these synthetic echo sequences can be projected into a low dimensional space.
 
@@
 
+\todo[inline]{Implement variations in echo strength. And include frequency specific absorption.}
+
 Finally, the impulse responses were converted to echo sequences by convolving them with a 1 millisecond long linear frequency-swept cosine starting at 100 kHz and ending at 40 Khz.
@@
 
-\begin{figure*}
+Artificial echo sequences were filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}.
+This model simulates the temporal and spectral resolution of the bat's cochlea.
+The model consists of a gammatone filterbank followed by half-wave rectification and non-linear compression.
+Finally, the model assumes a 1 kHz low-pass filter modeling the response characteristics of the inner hair cells.
+In the current implementation, we used a gammatone filterbank with xx filters, centered around xx kHz.
+The filters were spaced to correspond equivalent rectangular bandwidths \citep{Glasberg1990}.
+The first 6 ms of data were removed.
+This was done to ensure the data corresponded with the acoustic data.
+As indicated below, we removed the first 6 ms of those data to avoid including the pickup of the emitted signal.
+
+Figure \ref{fig:processing} depicts a single example of a synthetic echo sequence.
+It also depicts the resulting response for each of the xx channels of the \citet{Wiegrebe2008} model.
+From these responses, it can be inferred that these are highly correlated.
+This is confirmed when analyzing the average cross-channel correlation for all xx synthetic echo sequences (Fig.
+\ref{fig:processing}xx).
+From these high cross-channel correlations we conclude that the output of the cochlear model can be summarized by averaging across channels without extensive loss of information.
+An example of such a cross-channel sum is plotted in figure \ref{fig:processing}xx.
+In correspondence to our our previous work \citep{Vanderelst2016}, we refer to this average across frequency channels as a template.
+
+\begin{figure*}[tb]
+	\centering
+	\includegraphics[width=1\linewidth]{../results/processing}
+	\caption{Example of synthetic echo sequence and its processing using the \citet{Wiegrebe2008} model.
+(xx) Example of a single synthetic echo sequence.
+(xx-xx) Response of the xx frequency channels in the \citet{Wiegrebe2008} model.
+From these plots, it can be seen that channels highly correlate.
+This is confirmed in panel xx, depicting the average correlation across channels for all xx synthetic echo sequences analyzed.
+(xx) The output of the \citet{Wiegrebe2008} model summed across the frequency channels, is referred to as a template in this paper.
+The template resulting from the echo sequence in panel xx is shown in panel xx.
+Panel xx shows the cumulative proportion of explained variance as a function of number of \pc\ across all templates for all synthetic echo sequences.}
+	\label{fig:processing}
+\end{figure*}
+
+We used \PCA\ to assess the dimensionality of the  templates generated from the synthetic echo sequences.
+We found that almost all of the variance in the templates could be captured by \pca\ components (see figure \ref{fig:processing}xx).
+Some examples of templates reconstructed from the principal components are shown in \ref{fig:synthetictemplates}xx).
+
+\begin{figure*}[tb]
 	\centering
 	\includegraphics[width=1\linewidth]{../results/synthetic_templates}
-	\caption{Low dimensional mapping of synthetic echoes.
-Panels xx-xx, examples of artificial echo sequences.
-Panels xx-xx, templates derived from each of the echo sequences shown in xx-xx.
-The templates constructed from the artificial echoes could be projected in a lower dimensional space.
-A principal component analysis showed that \pca\ components captured almost all variance.
-Panels xx-xx, also depict (in blue) the result of projecting the templates into the \pca D space and transforming back to the orginal space.
-The good fit between the orginal and the reconstructed templates attests to the amount of variance captured by the \pca\ principal components.}
+	\caption{Example of synthetic echo sequences, their corresponding templates and the templates reconstructed from the \pca\ \pcs.
+(xx-xx) Three examples of synthetic echo sequence waveforms.
+Panels xx-xx depict the resulting templates (i.e., summation across the output channels of the \citet{Wiegrebe2008} model).
+These panels also show the result of projecting the templates into a \pca D space and transforming the result back to the original coordinates.}
 	\label{fig:synthetictemplates}
@@
 
-Artificial echo sequences were filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}.
-This model simulates the temporal and spectral resolution of the bat's cochlea.
-The output of the cochlear model was averaged across frequencies to obtain a time-intensity profile.
+%Selected examples of templates are depicted in figure \ref{fig:synthetictemplates}.
+
+%The output of the cochlear model was averaged across frequencies to obtain a time-intensity profile.
 The templates were downsampled to a sampling rate of 2.8 kHz.
@@
 This operation reduced the number of samples in a template from 7499 to \prepca\ data points.
-Finally, the first 6 ms of data were removed.
-This was done to ensure the data corresponded with the acoustic data.
-As indicated below, we removed the first 6 ms of those data to avoid including the pickup of the emitted signal.
-Selected examples of templates are depicted in figure \ref{fig:synthetictemplates}.
+Finally,
 
-We used principal component analysis to assess the dimensionality of the artificial templates.
+%We used principal component analysis to assess the dimensionality of the artificial templates.
 We found that almost all of the variance in the templates could be captured by \pca\ components (see figure \ref{fig:synthetictemplates}xx).
@@
 The components are visualized in figure \ref{fig:components}.
-
-\todo[inline]{Should we talk about the fact that the PCA model fits well across templates generated with different number of reflectors?
-This seems to depend on the echo having at least one strong (aka nearby) reflector.}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=1\linewidth]{../results/components}
-	\caption{Plots showing the \pca\ principal components derived from the templates constructed using synthetic echoes.}
-	\label{fig:components}
-\end{figure}
 
@@
 As we did for the synthetic echoes sequences, the echoes trains collected in bat habitats were filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}.
-Again, the output of the cochlear model was averaged across frequencies to obtain a time-intensity profile.
+As we concluded from the analyzing the synthetic echoes that the response in different frequency channels is highly correlated, the output of the cochlear model was averaged across frequencies to obtain a time-intensity profile.
 The beam of the emitter used was more narrow than the typically combined hearing and emission directionality of bats \citep{Vanderelst2010a,Reijniers2010,Jakobsen2012}.
 Therefore, templates were averaged across three neighboring directions.
-The templates were downsampled to a sampling rate of 2.8 kHz.
-Finally, the first 6 ms of data were removed to avoid using emitted signals for analysis.
+The first 6 ms of data were removed to avoid using emitted signals for analysis.
 Accordingly, each template consisted of \pca\ points.
@@
 
-The templates from both sites were mapped into the \pca-dimensional space derived from the synthetic echoes (i.e., using the components depicted in figure \ref{fig:components}), described above.
+The templates from both sites were mapped into the \pca-dimensional space derived from the synthetic echoes, described above.
 We assessed how well the templates collected in real habitats can be represented by the \pca\ components by mapping the templates into this space and transforming the results back to the original space.
@@
 
-For each site, we trained three neural networks, each taking the 25-sample templates as input.
+For each site, we trained three neural networks, each taking the \pca-sample templates as input.
 The first neural network was trained to output the azimuth direction associated with the template (31 targets).
@@
 This \pca-dimensional space could represent xx\% of the variance in the templates.
-Figure \ref{fig:templates} shows examples of templates and their reconstructed counterparts.
+Figure \ref{fig:empiricaltemplates} shows examples of templates and their reconstructed counterparts.
 Therefore, although the model used to generate echo sequences was very simple, the resulting underlying \pca D feature space could adequately represent templates resulting from real acoustic data.
 
-\begin{figure}
+\begin{figure*}[tb]
 	\centering
-	\includegraphics[width=1\linewidth]{../results/templates}
-	\caption{Examples of three templates selected from the Israel and Royal Fort data (red).
-Blue lines depict the result of projecting the templates in the \pca-dimensions space and reconstructing by transforming the results back to the original space.}
-	\label{fig:templates}
-\end{figure}
+	\includegraphics[width=1\linewidth]{../results/empirical_templates}
+	\caption{Examples of three templates generated from the echoes collected at the Israel and Royal Fort Gardens site. The panels also depict the result of projecting the templates in the 
+	\pca-dimensional space and transforming the result back to the original coordinates.}
+	\label{fig:empiricaltemplates}
+\end{figure*}
 
 In our previous work \cite{Vanderelst2016}, we showed that the templates could be used to recognize locations in bat habitats.
-As we have just shown that the templates can be projected into a lower \pca D feature space without loss, it can be expected that these \pca\ features can be used to locations as well.
+As we have shown that the templates can be projected into a lower \pca D feature space without loss, it can be expected that these \pca\ features can be used to locations as well.
 We found this to be indeed the case.
@@
 This figure shows that, in all cases, training almost reached an asymptote around or before 250 epochs.
-However, some additional training can be seen up to xxx epochs.
+However, some additional training can be seen up to xx epochs.
 
@@
 	\centering
-	\includegraphics[width=1\linewidth]{figures/israel_performance}
+	\includegraphics[width=1\linewidth]{../results/israel_results.pdf}
 	\caption{Classification results for the Israel site.
@@
 	\centering
-	\includegraphics[width=1\linewidth]{figures/royal_performance}
+	\includegraphics[width=1\linewidth]{../results/royal_results.pdf}
 	\caption{Simular as fig \ref{fig:israelperformance}, but for the Royal Fort Park site.}
@@
 \bibliography{references}
+
+\onecolumn
+\appendix
+\section{Supplementary material}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1\linewidth]{../results/components}
+	\caption{}
+	\label{fig:components}
+\end{figure}
+
 \end{document}

### 2020-03-13 Paper/main.tex
before running paper through grammarly

@@
 
-One mechanism for place recognition would be for bats to parse the stream of echoes returning from the environment \cite[e.g.,][]{Yovel2009,Vanderelst2016} into localized and identified objects.
+One mechanism for place recognition would be for bats to parse the stream of echoes returning from the environment into localized and identified objects [REF].
 However, as we argued before, it is highly questionable whether bats can interpret echoes in terms of a 3D environment model or acoustic image \citep[e.g.,][]{Vanderelst2015,Vanderelst2016,Steckel2013}.
@@
 While the time of arrival of echoes encodes object distance, azimuth and elevation directions are not directly available from the acoustic signal.
-In principle, azimuth and elevation can be estimated using spectral cues imposed by the head-related transfer function.
+In principle, azimuth and elevation can be estimated using spectral cues imposed by the head-related transfer function [REF].
 However, complex environments return many overlapping and interfering echoes, making reconstructing the direction of origin of echoes much harder, and perhaps impossible.
@@
 Some recent evidence seems to confirm this.
-First, \citet{Geberl2019} confirmed that bio-sonar has a very low spatial resolution, both for distance and angular dimensions.
-In addition, neurophysiological data suggest that closely spaced objects are not perceived individually by echolocating bats.
+First, \citet{Geberl2019} confirmed that bio-sonar has a low spatial resolution, both for distance and angular dimensions.
+Also, neurophysiological data suggest that closely spaced objects are not perceived seperatly by echolocating bats.
 In contrast, neural responses to echo cascades from densely spaced objects seem to represent a single extended stimulus \citep{Warnecke2018}.
@@
 
-In this paper, we directly evaluate the hypothesis that sonar provides bats inherently with a low dimensional, yet sufficiently rich, description of the world.
-
+In this paper, we directly evaluate the hypothesis that sonar provides bats inherently with a low dimensional, yet sufficiently detailed, description of the world.
 We test this hypothesis in two steps.
-In a first step, we generate a large collection of synthetic echo sequences.
+In a first step, we generate an extensive collection of synthetic echo sequences.
 We assess the compressibility of these echo signals using \PCA.
@@
 Nevertheless, we show that they can be compressed considerably.
-Moreover, the output of the cochlear model can be mapped onto an even lower dimensional space.
-In a second step, we show that echoes collected in real bat habitats can be mapped into the same lower dimensional space, without loss of information.
+Moreover, the output of the cochlear model can be mapped onto an even lower-dimensional space.
+In a second step, we show that echoes collected in real bat habitats can be mapped into the same low-dimensional space, without loss of information.
 We do this by demonstrating that the low dimensional feature space allows recognizing locations in real habitats.
@@
 Together these results suggest that echolocating bats live in a low dimensional perceptual world.
-The physics of the sonar system and the auditory periphery map the complexity of the 3D environment onto a small number of features, irrespective of whether we use artificial or real echo sequences.
+The physics of the sonar system and the auditory periphery map the complexity of the 3D environment onto a small number of independent features, irrespective of whether we use artificial or real echo sequences.
 Nevertheless, these features are sufficiently rich for bats to recognize locations, and as we discuss, support flight control.

### 2020-03-15 Paper/main.tex
intermediate

@@
 
-One mechanism for place recognition would be for bats to parse the stream of echoes returning from the environment into localized and identified objects [REF].
+One mechanism for place recognition would be for bats to parse the stream of echoes returning from the environment into localized and identified objects \citep{Lee2017,Barchi2013,Moss2001,Schnitzler2003,Simmons2012,Ulanovsky2008,Clare2015,Surlykke2016}.
 However, as we argued before, it is highly questionable whether bats can interpret echoes in terms of a 3D environment model or acoustic image \citep[e.g.,][]{Vanderelst2015,Vanderelst2016,Steckel2013}.
@@
 In principle, azimuth and elevation can be estimated using spectral cues imposed by the head-related transfer function [REF].
-However, complex environments return many overlapping and interfering echoes, making reconstructing the direction of origin of echoes much harder, and perhaps impossible.
+However, complex environments, and even single plants [REF], return many overlapping and interfering echoes, making reconstructing the direction of origin of echoes much harder, and perhaps impossible.
 Extracting spatial information is further hampered by the limited temporal resolution of the sonar system \citep{Simmons1989,Wiegrebe1996,Surlykke1996}.
@@
 The synthetic echo sequences differ greatly in complexity.
-Nevertheless, we show that they can be compressed considerably.
-Moreover, the output of the cochlear model can be mapped onto an even lower-dimensional space.
+Nevertheless, the output of the cochlear model can be mapped onto a low-dimensional space.
 In a second step, we show that echoes collected in real bat habitats can be mapped into the same low-dimensional space, without loss of information.
@@
 The physics of the sonar system and the auditory periphery map the complexity of the 3D environment onto a small number of independent features, irrespective of whether we use artificial or real echo sequences.
-Nevertheless, these features are sufficiently rich for bats to recognize locations, and as we discuss, support flight control.
+Nevertheless, this limited set of features is sufficiently rich for bats to recognize locations, and as we discuss, support flight control.
 
@@
 
-Each set of points was converted to an impulse response with a sample frequency $f_s$ of 219 kHz.
-The impulse response was limited to 7499 samples.
+Each set of points was converted to an impulse response with a sampling frequency $f_s$ of 219 kHz.
+The impulse response length was set to 7499 samples.
 These parameters correspond to the recording length and the sampling frequency used in the acoustic data collection (see below).
-In converting the point to an impulse response, we modeled attenuation of the echoes due to spherical spreading (both ways) and atmospheric attenuation.
-The atmospheric attenuation was xx dB/m, corresponding to the attenuation at xx kHz.
-We did not model any variation in target strength.
+In converting the point to an impulse response, we modeled attenuation of the echoes due to spherical spreading (both ways).
+We varied the amplitude of the impulses according to a normal distribution with a standard deviation of 10 dB.
 
-\todo[inline]{Implement variations in echo strength. And include frequency specific absorption.}
+The impulse responses were converted to echo sequences by convolving them with a 1 millisecond long linear frequency-swept cosine starting at 100 kHz and ending at 40 Khz.
+Again, these parameters correspond to those used in the acoustic data collection described below.
+Examples of the resulting echo sequences are shown in figures \ref{fig:processing} and \ref{fig:synthetictemplates}.
 
-Finally, the impulse responses were converted to echo sequences by convolving them with a 1 millisecond long linear frequency-swept cosine starting at 100 kHz and ending at 40 Khz.
-Again, these parameters correspond to those used in the acoustic data collection described below.
-Example of the resulting echo sequences are shown in figure \ref{fig:synthetictemplates}.
-
-Artificial echo sequences were filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}.
+Artificial echo sequences were filtered using a model of the bat's auditory periphery, as proposed by \citet{Wiegrebe2008}.
 This model simulates the temporal and spectral resolution of the bat's cochlea.
@@
 The filters were spaced to correspond equivalent rectangular bandwidths \citep{Glasberg1990}.
+The ouput of the channels were filtered with a time (distance) weighted factor corresponding to the atmospheric attenuation at the channel's center frequency.
+This filtering approximated frequency-dependent atmospheric attenuation.
 The first 6 ms of data were removed.
@@
 Figure \ref{fig:processing} depicts a single example of a synthetic echo sequence.
-It also depicts the resulting response for each of the xx channels of the \citet{Wiegrebe2008} model.
+It also illustrates the resulting response for each of the xx channels of the \citet{Wiegrebe2008} model.
 From these responses, it can be inferred that these are highly correlated.
@@
 \ref{fig:processing}xx).
-From these high cross-channel correlations we conclude that the output of the cochlear model can be summarized by averaging across channels without extensive loss of information.
+From these high cross-channel correlations, we conclude that the output of the cochlear model can be summarized by averaging across channels without extensive loss of information.
 An example of such a cross-channel sum is plotted in figure \ref{fig:processing}xx.
-In correspondence to our our previous work \citep{Vanderelst2016}, we refer to this average across frequency channels as a template.
+In correspondence to our previous work \citep{Vanderelst2016}, we refer to this average across frequency channels as a template.
 
@@
 
-%Selected examples of templates are depicted in figure \ref{fig:synthetictemplates}.
-
-%The output of the cochlear model was averaged across frequencies to obtain a time-intensity profile.
-The templates were downsampled to a sampling rate of 2.8 kHz.
-This sampling rate corresponds to a temporal integration of 350 $\mu$s, corresponding to what has been reported for bats' auditory systems \citep{Wiegrebe1996,Simmons1989}.
-This operation reduced the number of samples in a template from 7499 to \prepca\ data points.
-Finally,
-
-%We used principal component analysis to assess the dimensionality of the artificial templates.
-We found that almost all of the variance in the templates could be captured by \pca\ components (see figure \ref{fig:synthetictemplates}xx).
-Some examples of templates reconstructed from the principal components are shown in \ref{fig:synthetictemplates}xx).
-Panel xx of figure \ref{fig:synthetictemplates}xx) shows that most of the variance not captured by these components results in some additional noise for very low values of the templates.
-The components are visualized in figure \ref{fig:components}.
-
-In this section, we proposed a simple generative model for naturalistic echoes.
+In this section, we  used a generative model for naturalistic echoes.
 We assumed that echoes sequences received by bats can be considered as generated by anisotropically distributed point reflectors.
-Using this model, we have derived a low dimensional space ($D = \pca$) which can be used to describe the templates resulting from the artificial echo sequences.
-This implies that, after filtering through (a model of) the bats auditory system, these artificial echoes can be converted to low dimensional descriptions.
+Using this model, we have derived a low dimensional space ($D = \pca$), which can be used to describe the templates resulting from the artificial echo sequences.
+This implies that, after filtering through (a model of) the bat's auditory system, these artificial echoes can be converted to low dimensional descriptions.
 The original echo sequences, generated by xx to xx point reflectors effectively map to \pca\ dimensional vector.
@@
 
-In this work, we reuse the large library of templates constructed by \citet{Vanderelst2016}.
-Below, we briefly discuss the construction of the templates, and we refer to our previous work for more details.
+We reused the large library of echoes gathered in real bat habitats by \citet{Vanderelst2016}.
+Below, we briefly discuss the construction of templates based on these echoes, and we refer to our previous work for more details.
 \citet{Vanderelst2016} constructed an ensonification device featuring 31 Knowles FG series microphones connected to a custom-built sonar data acquisition board \citep{Steckel2013a}.
@@
 The ensonification device was mounted on a pan-tilt system allowing to rotate it in azimuth and elevation.
-The device and pan-tilt system were mounted on a tripod.
+The device and pan-tilt system was mounted on a tripod.
 
@@
 Below, we briefly outline the steps used to derive templates from the acoustic data.
-Again, we refer to our previous work for details on this procedure \citep{Vanderelst2016}.
 
 As we did for the synthetic echoes sequences, the echoes trains collected in bat habitats were filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}.
-As we concluded from the analyzing the synthetic echoes that the response in different frequency channels is highly correlated, the output of the cochlear model was averaged across frequencies to obtain a time-intensity profile.
+We concluded from analyzing the synthetic echoes that the response in different frequency channels is highly correlated.
+Therefore, the output of the cochlear model was averaged across frequencies to obtain a time-intensity profile, or template.
 The beam of the emitter used was more narrow than the typically combined hearing and emission directionality of bats \citep{Vanderelst2010a,Reijniers2010,Jakobsen2012}.
@@
 The first 6 ms of data were removed to avoid using emitted signals for analysis.
-Accordingly, each template consisted of \pca\ points.
 The processing steps outlined above resulted in 3 $\times$ 217 templates, for each location at each site, corresponding to 217 directions and three repeats.
@@
 Therefore, we removed them from the current analyses by omitting all templates with very low energy.
-The resulting number of templates included in the current analyses are given in table \ref{tab:sizes}.
+The resulting number of templates included in the present analyses are given in table \ref{tab:sizes}.
 
@@
 Previously, we assessed whether templates can be used to recognize locations using, what \citet{Baddeley2012} called, a perfect memory approach.
-Templates were stored in a lookup table, and discriminability (or recognizability) was assessed based on perfect representations of the templates.
-The Euclidean distance between two echo signatures was the metric used to assess how well bats should be able to distinguish between templates.
+Templates were stored in a lookup table.
+The Euclidean distance between two echo templates was the metric used to assess how well bats should be able to distinguish between templates.
 This table lookup and distance calculation require a comprehensive search of memory.
@@
 Here, we used a more biologically plausible memory system: an artificial neural network.
-These are neurally inspired way of distributed information storage in artificial neural networks \citep{Mcleod1998}.
-However, we acknowledge that current training methods most lickely have no biological counterpart.
+These are a neurally inspired way of distributed information storage in artificial neural networks \citep{Mcleod1998}.
+However, we acknowledge that current training methods most likely have no biological counterpart.
 
@@
 
-We should point that we did not take any steps to avoid so-called overfitting by the neural networks.
+We should point out that we did not take any steps to avoid so-called overfitting by the neural networks.
 Overfitting occurs when a neural network \citep{Ghotra2017}, or other mathematical models \citep{Hawkins2004}, extract complicated relationships between inputs and outputs that do not hold beyond the training examples.
@@
 However, in this case, the model is not required to generalize.
-Instead, we are only interested whether the neural networks are able to learn and classify the set of training examples.
+Instead, we are only interested in whether the neural networks can learn and classify the set of training examples.
 
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=1\linewidth]{../results/history}
-	\caption{Training history of the neural networks. The plots give the value of the loss function as a function of training epoch. Notice the logarithmic x-axis.}
-	\label{fig:history}
-\end{figure}
-
 Figure \ref{fig:israelperformance} summarizes the performance of the three neural networks trained on the Israel templates.
@@
 The number of small errors was larger for the neural network than for the perfect memory.
-The neural network was able to classify about 70\% of templates with an error of less than 30 degrees in azimuth.
+The neural network was able to classify about 70\% of templates with an error of fewer than 30 degrees in azimuth.
 In elevation, almost all templates were classified with an error below 15 degrees.
@@
 
+\citet{Reijniers2010a} quantified the transmission of auditory information in the cochlea of bats.
+In that study, the authors adapted the Meddis inner hair cell model \citep{Meddis2006} to assess the amount of information in the acoustic signal preserved by the cochlea.
+These authors concluded that spectral notches introduced by overlapping echoes are faithfully encoded by auditory nerve
+spike patterns.
+However, quantifying the information transfer rate also found that about half of the information (bits) are lost in the transduction from acoustic energy to spike trains.
+In the current paper, we expand on this work and quantify the dimension reduction due to cochlear encoding.
+Moreover, we show that this reduction in complexity still supports an ecologically valid task: sonar-based scene recognition.
+
 We assumed echoes received by bats can be modeled as originating from a large collection of point reflectors.
 Passing these synthetic echoes through a model of the auditory periphery, we obtained a time-intensity profile (template) for each echo sequence.
-These templates could be well described by a low dimensional space of about \pca\ principal components (figure \ref{fig:synthetictemplates}).
-We found that the same \pca D space could be used to describe templates originating from echoes collected in real bat habitats (figure \ref{fig:templates}).
-Therefore, both templates derived from synthetic echoes and templates constructed from real echoes map into the same lower dimensional space.
-This suggest that, in general, the filtering of the auditory periphery greatly reduces the dimensionality of the description of the bat's environment.
+These templates could be well described by a low dimensional space of \pca\ principal components (figure \ref{fig:synthetictemplates}).
+We found that the same \pca D space could be used to describe templates originating from echoes collected in real bat habitats (figure \ref{fig:empiricaltemplates}).
+This suggests the filtering of the auditory periphery provides the bat with a greatly reduced acoustic description of its environment.
+To identify the locus of the compression, we performed a number of additional analyses.
+In particular, we evaluated whether the low-dimensional description is a property of the echo sequences, the gammatone filters, or the 1 kHz low-pass filter in the \citet{Wiegrebe2008} model.
+
+We generated templates from the synthetic echo sequences using two versions of the \citet{Wiegrebe2008} model; with and without the 1 Khz low-pass filter.
+Per default, the model includes this filter to model the response characteristics of the inner hair cells \citep{Meddis2006,Reijniers2010a}.
+By removing this lowpass filter, we can assess to what extent it is responsible for the observed compressibility of the templates.
+We also generated templates using echo sequences of which the samples were randomly shuffled.
+This allows inferring whether the low-dimensionality of the templates is a property of the echo signals (before being filtered through the cochlear model).
+
+Figure \ref{fig:mechanism} shows the number of \pcs\ needed to capture 99\% of the templates variance.
+From this plot, it can be seen that the number of \pcs\ increases when shuffling the echo sequences increases from xx to xx.
+Omitting the 1 kHz filter from the  \citet{Wiegrebe2008} model increases the number of \pcs\ to xx.
+Inspecting the average spectrum of the templates (fig.
+\ref{fig:mechanism}, inset) shows that most energy in the templates is contained in frequencies below 1 kHz.
+Therefore, it ought to be possible to completely represent templates by sampling them at 2 kHz,  or about 70 samples per template, as templates consist of 7499 samples at 219 kHz sample rate.
+This corresponds to the number of \pcs\ required to represent templates derived from echo sequences that were shuffled  (fig.
+\ref{fig:mechanism}).
+The number of \pcs\ required to represent unshuffled templates is about half of this, i.c., \pca.
+Therefore, we conclude that the 1 kHz filter is the main, but not exclusive, locus of the dimensionality reduction.
+The structure of the echo signals also increases the dimensionality of the templates.
+
+\begin{figure}[tb]
+	\centering
+	\includegraphics[width=1\linewidth]{../results/mechanism}
+	\caption{}
+	\label{fig:mechanism}
+\end{figure}
 
@@
 They also use sonar to control flight, for example, avoid obstacles or following guiding structures.
-The description the bat has access to for flight control is has the same low level character than that used for place recognition.
-Therefore, our work suggests that this low level description is sufficient for flight control as well.
-We have shown before that a low level description could be sufficient for obstacle avoidance and corridor following \citep{Vanderelst2015a,Mansour2019}.
+The description the bat has access to for flight control has the same low-level character than that used for place recognition.
+Therefore, our work suggests that this low-level description is sufficient for flight control as well.
+We have shown before that a low-level description could be sufficient for obstacle avoidance and corridor following \citep{Vanderelst2015a,Mansour2019}.
 However, in that work, we converted both the echoes at the left and at the right ear to template-like representation.
@@
 
-We were unable to find numbers on the number of neurons in bat brains.
-However, the mouse cortex has been estimated to house about $9 \cdot 10^4$ neurons or $7 \cdot 10^8$ synapses per cubic millimeter \citep{Braitenberg2013}.
+We were unable to find numbers on the number of neurons in the cortex of bat species.
+However, the mouse cortex has been estimated to contain about $9 \cdot 10^4$ neurons or $7 \cdot 10^8$ synapses per cubic millimeter \citep{Braitenberg2013}.
 The neural networks presented in this paper consisted of about \nneurons\ neurons and \nweights\ connections.
@@
 
-Despite their small size, the networks in this paper were able to encode (memorize) the templates as least as well as the the perfect memory approach.
+Despite their small size, the networks in this paper were able to encode (memorize) the templates as least as well as the perfect memory approach.
 The neural networks presented here are tiny compared to the density of cortical matter.
@@
 Table \ref{tab:sizes} lists the number of weights for the artificial neural networks and compares this number with the size of the \pca-dimensional templates.
-The number of floats making up the templates is larger than the number of weights, by a factor of 2-3.
+The number of floats making up the templates is larger than the number of weights by a factor of 2-3.
 Therefore, the networks have a smaller memory footprint than the data they encode.
@@
 	\toprule
-		Site &       T &     n &      W &  W (nb) &      R \\
+	Site &       T &     n &      W &  W (nb\} &      R \\
 	\midrule
-		Israel &  179730 &  5991 &  73050 &   72000 &  40.6\% \\
-		Royal &  214890 &  7163 &  50745 &   49875 &  23.6\% \\
+	Israel &  206661 &  5299 &  53322 &   52425 &  25.8\% \\
+	Royal &  302991 &  7769 &  51945 &   51075 &  17.1\% \\
 	\bottomrule
@@
 
-%Together these two findings lend additional credibility to our hypothesis that bats could recognize locations using their acoustic signature, operationalized as templates.
-Before, we demonstrated that templates are highly specific to locations and directions.
-Here, we show that storing such templates does not require extensive neural hardware.
-Small neural networks can be used to store many (6000-7000) templates.
-Moreover, this is feasible because templates can effectively be compressed, which we demonstrated here by employing PCA and neural networks of a limited size.
-
-%It is well know that even small, shallow, neural networks are remarkably good at approximating potentially very complex function, even to the extent that it becomes valid to ask why they so powerful.
-\citet{Lin2017a} suggested that most datasets we are interested in originate from some simple physical law or some generative model with relatively
-%few free parameters.
-In the case of templates, the physical laws that generate them are quite straightforward.
-In addition, it can be assumed that the structure of natural environments can be described by generative models.
-Therefore, the ideas put forward by \citet{Lin2017a} might explain, on an abstract level why templates should not require extensive neural hardware to encode.
-
 \bibliography{references}
@@
 \section{Supplementary material}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1\linewidth]{../results/history}
+	\caption{Training history of the neural networks. The plots give the value of the loss function as a function of training epoch. Notice the logarithmic x-axis.}
+	\label{fig:history}
+\end{figure}
+
 \begin{figure}[h]

### 2020-03-15 RUN_ALL.py

@@ -1,0 +1,11 @@
+import script_preprocess_empirical_data
+
+import script_synthetic_echoes_to_PCA
+import script_plot_processing_synthetic
+
+import script_plot_template_examples
+import script_plot_components
+
+import script_run_training
+
+import script_run_perfect_memory

### 2020-03-15 dechirp_test.py

@@ -1,0 +1,32 @@
+from numpy.fft import fft, ifft
+from numpy import conj, pad, abs
+from library import generate_functions
+from matplotlib import pyplot
+
+pyplot.figure(figsize=(10,10))
+
+emission = generate_functions.create_emission()
+signal = generate_functions.create_emission()
+
+emission = pad(emission, (0, 100), 'constant')
+signal = pad(signal, (50, 50), 'constant')
+
+a = conj(fft(emission))
+b = fft(signal)
+c = ifft(a * b)
+dechriped = abs(c)
+
+
+pyplot.subplot(2,2,1)
+pyplot.plot(emission)
+pyplot.title('Emission')
+
+pyplot.subplot(2,2,2)
+pyplot.plot(signal)
+pyplot.title('Signal == Emission')
+
+pyplot.subplot(2,2,3)
+pyplot.plot(dechriped)
+pyplot.title('Dechirped signal (magnitude)')
+
+pyplot.show()

### 2020-03-15 fit.py

@@ -1,63 +1,0 @@
-from library import misc
-from library import settings
-from matplotlib import pyplot
-import pandas
-import numpy
-
-
-echoes = misc.pickle_load(settings.synthetic_echoes_file)
-indices = misc.pickle_load(settings.synthetic_echoes_indices)
-templates_synthetic = numpy.load(settings.synthetic_templates_file)
-
-pca_model = misc.pickle_load(settings.pca_templates_file)
-
-
-transformed_synthetic = pca_model.fit_transform(templates_synthetic)
-reconstructed_synthetic = pca_model.inverse_transform(transformed_synthetic)
-#%%
-n = reconstructed_synthetic.shape[0]
-keys = list(echoes.keys())
-
-correlations = []
-n_seeds = []
-n_clouds = []
-n_repeats = []
-
-for i in range(n):
-    k = keys[i]
-    k = k.split('_')
-    n_seed = int(k[0])
-    n_cloud = int(k[1])
-    n_repeat = int(k[2])
-
-    x = reconstructed_synthetic[i, :]
-    y = templates_synthetic[i, :]
-    r = numpy.corrcoef(x, y)[0, 1]
-
-    correlations.append(r)
-    n_seeds.append(n_seed)
-    n_clouds.append(n_cloud)
-    n_repeats.append(n_repeat)
-
-results = {'n_seed': n_seeds, 'n_cloud': n_clouds, 'n_repeat': n_repeats, 'corr': correlations}
-results = pandas.DataFrame(results)
-#%%
-grp = results.groupby(['n_seed', 'n_cloud'])
-mns = grp.mean()
-mns = mns.reset_index()
-
-table = mns.pivot_table(index='n_seed', columns='n_cloud', values='corr')
-print(table)
-
-pyplot.imshow(table)
-pyplot.show()
-
-#%%
-selected = 1
-t = templates_synthetic[selected,:]
-r = reconstructed_synthetic[selected,:]
-
-pyplot.plot(t)
-pyplot.plot(r)
-
-pyplot.show()

### 2020-03-15 library/generate_functions.py

@@ -10,6 +10,7 @@ from pyBat import Signal
 
 from scipy.signal import convolve
 
+
 def create_emission():
     emission_duration = settings.emission_duration
     sample_frequency = settings.sample_frequency
@@ -39,26 +40,35 @@ def generative_model(parameters):
         distances = numpy.concatenate((distances, g))
 
     # add close echo to ensure that even seqs with low n have a strong echo
-    #close = random.uniform(minimum_distance, minimum_distance+1, 1)
-    #g = numpy.array([minimum_distance])
-    #distances = numpy.concatenate((distances, close))
+    close = random.uniform(minimum_distance, minimum_distance + 1, 1)
+    distances = numpy.concatenate((distances, close))
+    # add "emission" for dechirping to work properly
+    # distances = numpy.concatenate((distances, [0]))
     return distances
 
-def distances2echo_sequence(distances, caller, emission):
+
+def distances2echo_sequence(distances, emission, noise_sd=10):
+    spreading_parameter = settings.spreading_parameter
+    call_level = settings.call_level
+    reflection_parameter = settings.reflection_parameter
+
+
     emission_duration = settings.emission_duration
     sample_frequency = settings.sample_frequency
     number_of_samples = settings.raw_collected_samples
     number_of_zero_samples = math.ceil(settings.initial_zero_time * sample_frequency)
-    # Get intensities
-    azimuths = numpy.zeros(distances.shape)
-    elevations = numpy.zeros(distances.shape)
-    call_result = caller.call(azimuths, elevations, distances)
-    left_db = call_result['echoes_left']
-    delays = call_result['delays']
+    # Get intensities - the atmospheric attenuation is taken care of when running the wiegrebe model
+    spreading = spreading_parameter * numpy.log10(distances)
+    echo_db = call_level + spreading + reflection_parameter
+
+    noise = random.normal(0, noise_sd, echo_db.shape)
+    echo_db = echo_db + noise
+    echo_db[echo_db<settings.detection_threshold] = 0
+    echo_db[echo_db>call_level] = call_level
+    delays = Acoustics.dist2delay(distances)
 
     # Make impulse response
-    left_db[left_db<0] = 0
-    ir_result = Acoustics.make_impulse_response(delays, left_db, emission_duration, sample_frequency)
[…]

### 2020-03-15 library/misc.py

@@ -6,8 +6,30 @@ import numpy
 import scipy.interpolate as interpolate
 from matplotlib import pyplot
 from sklearn.preprocessing import Normalizer
+import string
 
 from sklearn.decomposition import PCA
+
+def template2fft(template):
+    template = template - numpy.mean(template)
+    spectrum = numpy.abs(numpy.fft.fft(template))
+    return spectrum
+
+
+def label(x,y , txt, ax=None, fontsize=12):
+    if type(txt)==int: txt = '(' + string.ascii_letters[txt] + ')'
+    if ax is None: ax = pyplot.gca()
+    pyplot.text(x, y, txt, transform=ax.transAxes,
+                horizontalalignment='center',
+                verticalalignment = 'center',
+                fontsize=fontsize)
+
+def average_correlation_matrix(array):
+    size =array.shape[0]
+    mask = ~numpy.eye(size, dtype=bool)
+    values = array[mask]
+    mn = numpy.mean(values)
+    return mn
 
 
 def scale_synthetic_templates(empirical_templates, synthetic_templates):
@@ -18,6 +40,7 @@ def scale_synthetic_templates(empirical_templates, synthetic_templates):
     pyplot.plot(mean_empirical)
     pyplot.plot(mean_synthetic)
     pyplot.legend(['Empirical', 'Synthetic'])
+    pyplot.title('Before scaling')
     pyplot.show()
 
     # Plot correct mean and offset
@@ -34,6 +57,7 @@ def scale_synthetic_templates(empirical_templates, synthetic_templates):
     pyplot.plot(mean_empirical)
     pyplot.plot(mean_synthetic)
     pyplot.legend(['Empirical', 'Synthetic'])
+    pyplot.title('After scaling scaling')
     pyplot.show()
 
 
@@ -58,8 +82,7 @@ def project_and_reconstruct(pca_model, data):
     return transformed, reconstructed, correlation
 
 
-
-def run_pca(data, criterion=0.99, save_file = None):
+def run_pca(data, criterion, save_file = None):
     # Run all components
     pca_model = PCA()
     pca_model.fit(data)
@@ -76,7 +99,6 @@ def run_pca(data, criterion=0.99, save_file = None):
     result['correlation'] = correlation
     result['suggestion'] = suggestion
 
-
     if save_file is not None:
         numpy.savez_compressed(save_file,
                                original=data,
[…]

### 2020-03-15 library/settings.py

@@ -18,31 +18,45 @@ noise_floor = 0.18 # should be determined on open space
 stochaistic_noise = 0.02 # stdv, See paper for determination, before pca mapping
 integration_time = 350 / 1000000  # integration constant
 initial_zero_time = 6 / 1000  # initial time to omit
-n_components_templates = 210
-n_components_echoes = 210
+
 ####################################################
 ##Generative modelling settings
 ####################################################
+reflection_parameter = -40
+spreading_parameter = -40
+call_level = 110
+detection_threshold = 20
+
 emission_duration = 0.001
-emission_frequency_low = 100000
-emission_frequency_high = 40000
+emission_frequency_low = 40000
+emission_frequency_high = 100000
 emission_frequency_mean = round((emission_frequency_low + emission_frequency_high) / 2)
+
+####################################################
+##PCA settings
+####################################################
+pca_criterion=0.99
+
+####################################################
+##SAVE FILES
+####################################################
 
 synthetic_impulses_file = os.path.join(result_folder, 'synthetic_impulses.npz')
 synthetic_echoes_file = os.path.join(result_folder, 'synthetic_echoes.npz')
 synthetic_templates_file = os.path.join(result_folder, 'synthetic_templates.npz')
 
-synthetic_impulses_pca_results = os.path.join(result_folder, 'synthetic_impulses_pca.npz')
-synthetic_echoes_pca_results = os.path.join(result_folder, 'synthetic_echoes_pca.npz')
+#synthetic_impulses_pca_results = os.path.join(result_folder, 'synthetic_impulses_pca.npz')
+#synthetic_echoes_pca_results = os.path.join(result_folder, 'synthetic_echoes_pca.npz')
 synthetic_templates_pca_results = os.path.join(result_folder, 'synthetic_templates_pca.npz')
 
-pca_impulses_model_file = os.path.join(result_folder, 'PCA_model_impulses.pca')
-pca_echoes_model_file = os.path.join(result_folder, 'PCA_model_echoes.pca')
+#pca_impulses_model_file = os.path.join(result_folder, 'PCA_model_impulses.pca')
+#pca_echoes_model_file = os.path.join(result_folder, 'PCA_model_echoes.pca')
 pca_templates_model_file = os.path.join(result_folder, 'PCA_model_templates.pca')
 
[…]

### 2020-03-15 library/smoothn.py

@@ -9,7 +9,7 @@ import numpy.ma as ma
 def smoothn(y,nS0=10,axis=None,smoothOrder=2.0,sd=None,verbose=False,\
 	s0=None,z0=None,isrobust=False,W=None,s=None,MaxIter=100,TolZ=1e-3,weightstr='bisquare'):
   '''
-   function [z,s,exitflag,Wtot] = smoothn(varargin)
+   function [z,template,exitflag,Wtot] = smoothn(varargin)
 
    SMOOTHN Robust spline smoothing for 1-D to N-D data.
    SMOOTHN provides a fast, automatized and robust discretized smoothing
@@ -82,21 +82,21 @@ def smoothn(y,nS0=10,axis=None,smoothOrder=2.0,sd=None,verbose=False,\
    Examples:
    --------
    # 1-D example
-   x = linspace(0,100,2**8);
-   y = cos(x/10)+(x/50)**2 + randn(size(x))/10;
+   average_spectrum = linspace(0,100,2**8);
+   y = cos(average_spectrum/10)+(average_spectrum/50)**2 + randn(size(average_spectrum))/10;
    y[[70, 75, 80]] = [5.5, 5, 6];
    z = smoothn(y); # Regular smoothing
    zr = smoothn(y,'robust'); # Robust smoothing
-   subplot(121), plot(x,y,'seed_points.',x,z,'k','LineWidth',2)
+   subplot(121), plot(average_spectrum,y,'seed_points.',average_spectrum,z,'k','LineWidth',2)
    axis square, title('Regular smoothing')
-   subplot(122), plot(x,y,'seed_points.',x,zr,'k','LineWidth',2)
+   subplot(122), plot(average_spectrum,y,'seed_points.',average_spectrum,zr,'k','LineWidth',2)
    axis square, title('Robust smoothing')
 
    # 2-D example
    xp = 0:.02:1;
-   [x,y] = meshgrid(xp);
-   f = exp(x+y) + sin((x-2*y)*3);
-   fn = f + randn(size(f))*0.5;
+   [average_spectrum,y] = meshgrid(xp);
+   frequency_axis = exp(average_spectrum+y) + sin((average_spectrum-2*y)*3);
+   fn = frequency_axis + randn(size(frequency_axis))*0.5;
    sample_frequency = smoothn(fn);
    subplot(121), surf(xp,xp,fn), zlim([0 8]), axis square
    subplot(122), surf(xp,xp,sample_frequency), zlim([0 8]), axis square
[…]

### 2020-03-15 old_scraps/weights.py

@@ -1,22 +1,0 @@
-from library import misc
-
-# royal_components are in the rows
-# components_array, shape (n_components, n_features)
-
-
-
-data_set = 'royal'
-selected_dimension = 'lcs'
-file_names = misc.folder_names(data_set, selected_dimension)
-
-pca = misc.pickle_load(file_names['pca_file'])
-
-components = pca.components_
-
-#new_model = keras.models.load_model(file_names['model_file'])
-#weights = new_model.get_weights()
-#%%
-#x = weights[4]
-#x = numpy.abs(x)
-#pyplot.imshow(x)
-#pyplot.show()

### 2020-03-15 plot_training_history.py

@@ -3,11 +3,9 @@ import os
 from matplotlib import pyplot
 from library import smoothn, misc, settings
 
-training_history_plot_file = settings.training_history_plot
-
 pyplot.style.use(settings.style)
 
-pyplot.figure(figsize=(5,2.5))
+pyplot.figure(figsize=(7,3))
 
 for data_set in ['israel', 'royal']:
     if data_set == 'israel': pyplot.subplot(1, 2, 1)
@@ -36,5 +34,5 @@ for data_set in ['israel', 'royal']:
 
 pyplot.tight_layout()
 pyplot.legend(['Azimuth', 'Elevation', 'Location'])
-pyplot.savefig(training_history_plot_file)
+pyplot.savefig(settings.training_history_plot)
 pyplot.show()

### 2020-03-15 plot_training_results.py

@@ -17,7 +17,6 @@ results_els = misc.pickle_load(files_els['results_file'])
 results_lcs = misc.pickle_load(files_lcs['results_file'])
 perfect_memory = misc.pickle_load(files_lcs['perfect_memory_file'])
 
-output_file = os.path.join(settings.figure_folder, data_set + '_performance.pdf')
 
 # %% Plot errors
 
@@ -100,5 +99,6 @@ pyplot.xlabel('Location (meters)')
 pyplot.ylim([0, 5])
 
 pyplot.tight_layout()
-pyplot.savefig(output_file)
+if data_set == 'israel': pyplot.savefig(settings.israel_results_plot)
+if data_set == 'royal': pyplot.savefig(settings.royal_results_plot)
 pyplot.show()

### 2020-03-15 script_plot_components.py

@@ -8,9 +8,8 @@ from library import misc, settings
 from PyAstronomy.pyaC import zerocross1d
 
 pyplot.style.use(settings.style)
-n = settings.n_components_templates
 
-pca_model = misc.pickle_load(settings.pca_templates_file)
+pca_model = misc.pickle_load(settings.pca_templates_model_file)
 components = pca_model.components_
 
 mx = numpy.max(numpy.abs(components))
@@ -19,13 +18,13 @@ x = numpy.array(range(n))
 
 pyplot.figure(figsize=(10, 7))
 
-for i in range(20):
+for i in range(25):
     component = components[i, :]
     crossings = zerocross1d(x, component)
     n_crossings = crossings.size
     title = '$PC_{%i}$' % (i + 1)
 
-    pyplot.subplot(4, 5, i + 1)
+    pyplot.subplot(5, 5, i + 1)
     pyplot.plot(component)
     pyplot.ylim([-mx, mx])
     pyplot.hlines(0, 0, n, alpha=0.25)

### 2020-03-15 script_plot_processing_synthetic.py

@@ -1,0 +1,110 @@
+import math
+
+import numpy
+from matplotlib import pyplot
+
+
+from library import misc
+from library import generate_functions
+from library import settings
+from pyBat import Wiegrebe
+
+sample_frequency = settings.sample_frequency
+number_of_samples = settings.raw_collected_samples
+initial_zero_time = settings.initial_zero_time
+emission_frequency_mean = settings.emission_frequency_mean
+initial_zero_samples = math.ceil(initial_zero_time * sample_frequency)
+
+emission = generate_functions.create_emission()
+
+emission_padding = number_of_samples - emission.shape[0]
+padded_emission = numpy.pad(emission, (0, emission_padding), 'constant')
+
+wiegrebe = Wiegrebe.ModelWiegrebe(sample_frequency, emission_frequency_mean, 4, emission=padded_emission)
+
+generative_model_parameters = {}
+generative_model_parameters['n_seed_points'] = [100]
+generative_model_parameters['n_cloud_points'] = [100]
+distances = generate_functions.generative_model(generative_model_parameters)
+echo_sequence, impulse_response = generate_functions.distances2echo_sequence(distances, emission)
+echo_sequence[0:initial_zero_samples] = 0
+template = generate_functions.echo_sequence2template(echo_sequence, wiegrebe)
+
+# normalize stuff for easier plotting
+echo_sequence = echo_sequence/numpy.max(numpy.abs(echo_sequence))
+
+bm = wiegrebe.filtered
+bm = bm - numpy.min(bm)
+bm = bm / numpy.max(bm)
+
+
+template = template - numpy.min(template)
+template = template / numpy.max(template)
+
+correlations = misc.pickle_load(settings.channel_correlations_file)
+correlations = numpy.array(correlations)
+
+synthetic_template_pca = misc.pickle_load(settings.pca_templates_model_file)
+cumulative = numpy.cumsum(synthetic_template_pca.explained_variance_ratio_)
+
+# %%
+pyplot.style.use(settings.style)
+
+fig = pyplot.figure(figsize=(20, 10))
+
+gs = fig.add_gridspec(8, 3)
+ax = fig.add_subplot(gs[:, 0])
[…]

### 2020-03-15 script_plot_template_examples.py

@@ -1,15 +1,11 @@
 import numpy
-import os
 from matplotlib import pyplot
 from library import misc, settings
+
 
 #
 # Synthetic echoes, templates and reconstruction
 #
-
-mx_echoes = 0.1
-mx_templates = 0.25
-mn_templates = 0.17
 
 index0 = 10
 index1 = 100
@@ -21,20 +17,11 @@ print('-----> LOADING DATA')
 loaded = numpy.load(settings.synthetic_echoes_file)
 synthetic_echoes = loaded['synthetic_echoes']
 
-loaded = numpy.load(settings.scaled_synthetic_templates_file)
-synthetic_templates = loaded['synthetic_templates']
+loaded = numpy.load(settings.synthetic_templates_pca_results)
+synthetic_templates = loaded['original']
+reconstructed_synthetic_templates = loaded['reconstructed']
 
-loaded = numpy.load(settings.reconstructed_synthetic_templates_file)
-reconstructed_synthetic_templates = loaded['reconstructed_synthetic_templates']
-
-loaded = numpy.load(settings.reconstructed_synthetic_echoes_file)
-reconstructed_synthetic_echoes = loaded['reconstructed_synthetic_echoes']
-
-
-pca_model = misc.pickle_load(settings.pca_templates_file)
-cummulative_variance = numpy.cumsum(pca_model.explained_variance_ratio_)
-
-#x = synthetic_templates.flatten()
+#average_spectrum = synthetic_templates.flatten()
 #y = reconstructed_synthetic_templates.flatten()
 
 print('-----> GETTING DATA FROM ARRAYS')
@@ -47,122 +34,118 @@ template_0 = synthetic_templates[index0, :]
 template_1 = synthetic_templates[index1, :]
 template_2 = synthetic_templates[index2, :]
 
-reconstructed_echo_0 = reconstructed_synthetic_echoes[index0, :]
-reconstructed_echo_1 = reconstructed_synthetic_echoes[index1, :]
-reconstructed_echo_2 = reconstructed_synthetic_echoes[index2, :]
-
 reconstructed_template_0 = reconstructed_synthetic_templates[index0, :]
 reconstructed_template_1 = reconstructed_synthetic_templates[index1, :]
[…]

### 2020-03-15 script_run_identify_mechanism.py

@@ -1,0 +1,131 @@
+import numpy
+from matplotlib import pyplot
+from scipy.signal import hilbert
+
+from library import misc
+from library import settings
+from library import generate_functions
+from pyBat import Wiegrebe
+
+number_of_samples = settings.raw_collected_samples
+emission = generate_functions.create_emission()
+emission_padding = number_of_samples - emission.shape[0]
+padded_emission = numpy.pad(emission, (0, emission_padding), 'constant')
+sample_frequency = settings.sample_frequency
+emission_frequency_mean = settings.emission_frequency_mean
+
+wiegrebe = Wiegrebe.ModelWiegrebe(sample_frequency, emission_frequency_mean, 4, emission=padded_emission)
+
+do_calc = True
+
+if do_calc:
+
+    print('---> LOADING SYNTHETIC DATA')
+    loaded = numpy.load(settings.synthetic_echoes_file)
+    synthetic_echoes = loaded['synthetic_echoes']
+    n = synthetic_echoes.shape[0]
+
+    shuffles = [False, False, True, True]
+    filters = [True, False, True, False]
+    results = {}
+    pcs = []
+    labels = []
+
+    for filter, shuffle in zip(filters, shuffles):
+        spectra = []
+        templates = []
+        for i in range(n):
+            print(i, filter, shuffle)
+            signal = synthetic_echoes[i, :]
+            if shuffle: numpy.random.shuffle(signal)
+            template = wiegrebe.run_model(signal, dechirp=True, apply_attenuation=True)
+            if not filter: template = numpy.mean(wiegrebe.compressed, axis=0)
+            spectrum = misc.template2fft(template)
+            spectra.append(spectrum)
+            templates.append(template)
+
+        spectra = numpy.array(spectra)
+        templates = numpy.array(templates)
+        label = '%s_%s' % (filter, shuffle)
+        results[label] = [spectra, templates]
+
+        model, stats = misc.run_pca(templates, criterion=settings.pca_criterion)
+        pcs.append(stats["suggestion"])
+        labels.append(label)
+
+    results['pcs'] = pcs
+    results['labels'] = labels
+    misc.pickle_save(settings.mechanism_data_file, results)
+# %%
+from library import settings
[…]

### 2020-03-15 script_run_training.py

@@ -6,99 +6,76 @@ from tensorflow import keras
 
 from library import misc, process_functions, settings
 
-data_set = 'royal'
+for data_set in ['royal', 'israel']:
+    layers = [50, 75, 75, 50]
+    nr_epochs = 5000
+
+    for selected_dimension in ['lcs', 'azs','els']:
+        file_names = misc.folder_names(data_set, selected_dimension)
+
+        # Read prepared data
+        data = numpy.load(file_names['npz_file'])
+        pca_model = misc.pickle_load(settings.pca_templates_model_file)
+
+        # Select inputs
+        if selected_dimension == 'lcs': unencoded_data = data['long_lcs']
+        if selected_dimension == 'azs': unencoded_data = data['long_azs']
+        if selected_dimension == 'els': unencoded_data = data['long_els']
+        if selected_dimension == 'lcs' and data_set in ['israel', 'royal']: unencoded_data = misc.map_lcs_to_distances(data)
+
+        # Encode targets
+        encoder, targets = process_functions.get_encoding(unencoded_data)
+        target_n = targets.shape[1]
+
+        # Get PCA-ed template inputs
+        templates = data['long_data']
+        inputs = pca_model.transform(templates)
+        n_components = inputs.shape[1]
 
 
-layers = [50, 75, 75, 50]
-nr_epochs = 5000
+        # Scale inputs to a minimum of zero
+        inputs = inputs - numpy.min(inputs)
 
-for selected_dimension in ['lcs', 'azs','els']:
-    file_names = misc.folder_names(data_set, selected_dimension)
+        # Make model
+        model = keras.Sequential()
+        noise_layer = keras.layers.GaussianNoise(input_shape=(n_components,), stddev=settings.stochaistic_noise * 1)
+        output_layer = keras.layers.Dense(target_n, activation='softmax')
+        model.add(noise_layer)
+        for nodes in layers: model.add(keras.layers.Dense(nodes, activation='relu'))
+        model.add(output_layer)
 
-    # Read prepared data
-    data = numpy.load(file_names['npz_file'])
-    pca_model = misc.pickle_load(settings.pca_templates_file)
+        # Train Model
+        loss = keras.losses.CategoricalCrossentropy()
+        model.compile('adam', loss=loss)
+        training_history = model.fit(inputs, targets, epochs=nr_epochs)
+        model.save(file_names['model_file'])
 
-    # Select inputs
[…]

### 2020-03-15 script_synthetic_echoes_to_PCA.py

@@ -3,17 +3,15 @@ import numpy
 from library import settings
 from library import misc
 from library import generate_functions
-from pyBat import Call
 from pyBat import Wiegrebe
 
 repeats = 10
 n_seed_levels = [5, 10, 20, 50, 100, 200]
 n_cloud_levels = [5, 10, 20, 50, 100, 200]
 
-#repeats = 3
-#n_seed_levels = [5, 10]
-#n_cloud_levels = [5, 10]
-
+#repeats = 10
+#n_seed_levels = [100, 200]
+#n_cloud_levels = [100, 200]
 
 sample_frequency = settings.sample_frequency
 number_of_samples = settings.raw_collected_samples
@@ -21,7 +19,6 @@ initial_zero_time = settings.initial_zero_time
 emission_frequency_mean = settings.emission_frequency_mean
 initial_zero_samples = math.ceil(initial_zero_time * sample_frequency)
 
-caller = Call.Call(None, [emission_frequency_mean])
 emission = generate_functions.create_emission()
 
 emission_padding = number_of_samples - emission.shape[0]
@@ -32,18 +29,30 @@ wiegrebe = Wiegrebe.ModelWiegrebe(sample_frequency, emission_frequency_mean, 4, 
 synthetic_impulses = []
 synthetic_echoes = []
 synthetic_templates = []
+correlations = []
 
 for seed_level_i in range(len(n_seed_levels)):
     for cloud_level_i in range(len(n_cloud_levels)):
         print(seed_level_i, cloud_level_i)
         for i in range(repeats):
+            print(i)
             generative_model_parameters = {}
             generative_model_parameters['n_seed_points'] = n_seed_levels[seed_level_i]
             generative_model_parameters['n_cloud_points'] = n_cloud_levels[cloud_level_i]
             distances = generate_functions.generative_model(generative_model_parameters)
-            echo_sequence, impulse_response = generate_functions.distances2echo_sequence(distances, caller, emission)
-            echo_sequence[0:initial_zero_samples] = 0
+            echo_sequence, impulse_response = generate_functions.distances2echo_sequence(distances, emission)
             template = generate_functions.echo_sequence2template(echo_sequence, wiegrebe)
+
+            impulse_response[0:initial_zero_samples] = 0
+            echo_sequence[0:initial_zero_samples] = 0
+            template[0:initial_zero_samples] = 0
[…]

### 2020-03-15 script_tabulate_memory_use.py

@@ -10,13 +10,18 @@ ws_nb = []
 rs = []
 ts = []
 
+loaded = numpy.load(settings.synthetic_templates_pca_results)
+transformed = loaded['transformed']
+n_components_templates = transformed.shape[1]
+
 for data_set in ['israel', 'royal']:
     selected_dimension = 'lcs'
     file_names = misc.folder_names(data_set, selected_dimension)
     data = numpy.load(file_names['npz_file'])
     templates = data['long_data']
 
-    total_n_numbers = templates.shape[0] * settings.n_components_templates
+
+    total_n_numbers = templates.shape[0] * n_components_templates
 
     new_model = keras.models.load_model(file_names['model_file'])
     weights = new_model.get_weights()

### 2020-03-15 todo_list.py

@@ -1,0 +1,2 @@
+#todo: Implement the dechirping the wiegrebe model properly. Now, results are complex numbers. That can't be correct.
+

### 2020-03-15 library/generate_functions.py
Solved the dechirping problem, I think

@@ -82,7 +82,7 @@ def distances2echo_sequence(distances, emission, noise_sd=10):
 
 
 def echo_sequence2template(echo_sequence, wiegrebe):
-    wiegrebe_result = wiegrebe.run_model(echo_sequence, dechirp=True, apply_attenuation=True)
+    wiegrebe_result = wiegrebe.run_model(echo_sequence, apply_attenuation=True)
     wiegrebe_result = wiegrebe_result.reshape(1, -1)
     wiegrebe_result = wiegrebe_result[0]
     return wiegrebe_result

### 2020-03-15 script_run_identify_mechanism.py

@@ -14,13 +14,14 @@ padded_emission = numpy.pad(emission, (0, emission_padding), 'constant')
 sample_frequency = settings.sample_frequency
 emission_frequency_mean = settings.emission_frequency_mean
 
-wiegrebe = Wiegrebe.ModelWiegrebe(sample_frequency, emission_frequency_mean, 4, emission=padded_emission)
+wiegrebe = Wiegrebe.ModelWiegrebe(sample_frequency, emission_frequency_mean, 4)
 
 do_calc = True
 
 if do_calc:
 
     print('---> LOADING SYNTHETIC DATA')
+    # THESE HAVE ALREADY BEEN DECHIRPED .....
     loaded = numpy.load(settings.synthetic_echoes_file)
     synthetic_echoes = loaded['synthetic_echoes']
     n = synthetic_echoes.shape[0]
@@ -38,7 +39,7 @@ if do_calc:
             print(i, filter, shuffle)
             signal = synthetic_echoes[i, :]
             if shuffle: numpy.random.shuffle(signal)
-            template = wiegrebe.run_model(signal, dechirp=True, apply_attenuation=True)
+            template = wiegrebe.run_model(signal, apply_attenuation=True)
             if not filter: template = numpy.mean(wiegrebe.compressed, axis=0)
             spectrum = misc.template2fft(template)
             spectra.append(spectrum)

### 2020-03-15 script_synthetic_echoes_to_PCA.py

@@ -3,7 +3,7 @@ import numpy
 from library import settings
 from library import misc
 from library import generate_functions
-from pyBat import Wiegrebe
+from pyBat import Wiegrebe, Signal
 
 repeats = 10
 n_seed_levels = [5, 10, 20, 50, 100, 200]
@@ -21,10 +21,8 @@ initial_zero_samples = math.ceil(initial_zero_time * sample_frequency)
 
 emission = generate_functions.create_emission()
 
-emission_padding = number_of_samples - emission.shape[0]
-padded_emission = numpy.pad(emission, (0, emission_padding), 'constant')
 
-wiegrebe = Wiegrebe.ModelWiegrebe(sample_frequency, emission_frequency_mean, 4, emission=padded_emission)
+wiegrebe = Wiegrebe.ModelWiegrebe(sample_frequency, emission_frequency_mean, 4)
 
 synthetic_impulses = []
 synthetic_echoes = []
@@ -41,6 +39,7 @@ for seed_level_i in range(len(n_seed_levels)):
             generative_model_parameters['n_cloud_points'] = n_cloud_levels[cloud_level_i]
             distances = generate_functions.generative_model(generative_model_parameters)
             echo_sequence, impulse_response = generate_functions.distances2echo_sequence(distances, emission)
+            echo_sequence = Signal.dechirp(emission, echo_sequence)
             template = generate_functions.echo_sequence2template(echo_sequence, wiegrebe)
 
             impulse_response[0:initial_zero_samples] = 0

### 2020-03-23 Paper/main.tex
Herberts edits

@@
 %number of pca components
-\newcommand{\pca}{25}
-%length of templates before pcs
-\newcommand{\prepca}{92}
+\newcommand{\pca}{xx}
 %number of neurons in models
-\newcommand{\nneurons}{200}
+\newcommand{\nneurons}{xx}
 %number of connections
@@
 
-One mechanism for place recognition would be for bats to parse the stream of echoes returning from the environment into localized and identified objects \citep{Lee2017,Barchi2013,Moss2001,Schnitzler2003,Simmons2012,Ulanovsky2008,Clare2015,Surlykke2016}.
+One mechanism for place recognition would be for bats to parse the stream of echoes returning from the environment into localized and identified objects \citep{Lee2017,Barchi2013,Moss2001,Schnitzler2003,Simmons2012,Ulanovsky2008,Clare2015,Surlykke2016,Geipel2013}.
 However, as we argued before, it is highly questionable whether bats can interpret echoes in terms of a 3D environment model or acoustic image \citep[e.g.,][]{Vanderelst2015,Vanderelst2016,Steckel2013}.
@@
 While the time of arrival of echoes encodes object distance, azimuth and elevation directions are not directly available from the acoustic signal.
-In principle, azimuth and elevation can be estimated using spectral cues imposed by the head-related transfer function [REF].
-However, complex environments, and even single plants [REF], return many overlapping and interfering echoes, making reconstructing the direction of origin of echoes much harder, and perhaps impossible.
+In principle, azimuth and elevation can be estimated using spectral cues imposed by the head-related transfer function \citep[e.g.,][]{Reijniers2010,Aytekin2004}.
+However, complex environments, and even single plants \citep{Yovel2008,Muller2000}, return many overlapping and interfering echoes, making reconstructing the direction of origin of echoes much harder, and perhaps impossible.
 Extracting spatial information is further hampered by the limited temporal resolution of the sonar system \citep{Simmons1989,Wiegrebe1996,Surlykke1996}.
@@
 Also, neurophysiological data suggest that closely spaced objects are not perceived seperatly by echolocating bats.
-In contrast, neural responses to echo cascades from densely spaced objects seem to represent a single extended stimulus \citep{Warnecke2018}.
-\todo[inline]{Look at recent Simmons paper where bats refuse to fly through hoop}
+In contrast, neural responses to echo cascades from densely spaced objects seem to represent a single extended stimulus \citep{Warnecke2018,Knowles2015}.
 
@@
 
-In this paper, we directly evaluate the hypothesis that sonar provides bats inherently with a low dimensional, yet sufficiently detailed, description of the world.
+In this paper, we directly evaluate the hypothesis that sonar provides bats inherently with a low dimensional, yet functionally sufficient, description of the world.
 We test this hypothesis in two steps.
 In a first step, we generate an extensive collection of synthetic echo sequences.
-We assess the compressibility of these echo signals using \PCA.
-Next, we filter the echo sequences using a model of the bat's auditory periphery \citep{Wiegrebe2008}.
+Next, we process the echo sequences using a model of the bat's auditory periphery \citep{Wiegrebe2008}.
+We assess the compressibility of the output of the cochlear model using \PCA.
 The synthetic echo sequences differ greatly in complexity.
@@
 In a second step, we show that echoes collected in real bat habitats can be mapped into the same low-dimensional space, without loss of information.
-We do this by demonstrating that the low dimensional feature space allows recognizing locations in real habitats.
+We verify this by demonstrating that the low dimensional feature space allows recognizing locations in real habitats.
 
@@
 In converting the point to an impulse response, we modeled attenuation of the echoes due to spherical spreading (both ways).
-We varied the amplitude of the impulses according to a normal distribution with a standard deviation of 10 dB.
+We sampled the amplitude of the impulses from a normal distribution with a standard deviation of 10 dB.
 
-The impulse responses were converted to echo sequences by convolving them with a 1 millisecond long linear frequency-swept cosine starting at 100 kHz and ending at 40 Khz.
+The impulse responses were converted to echo sequences by convolving them with a bat-like emission.
+We used a 1 millisecond long linear frequency-swept cosine starting at 100 kHz and ending at 40 Khz.
 Again, these parameters correspond to those used in the acoustic data collection described below.
+The echo sequences were dechirped as follows,
+%
+%ifft(complex conj(ftt emission) .* fft(echo))
+%
+\begin{equation}\label{key}
+E_d \mathcal{F}^{-1}(\mathcal{F}(S)^* . \mathcal{F}(E))
+\end{equation}
+with $\mathcal{F}$ denoting the Fourier transform and $()^*$ the complex conjugate.
+The symbols $S$, $E$, $E_d$ refer to the emission, the echo sequence and the dechirped echo sequence, respectively.
 Examples of the resulting echo sequences are shown in figures \ref{fig:processing} and \ref{fig:synthetictemplates}.
@@
 This model simulates the temporal and spectral resolution of the bat's cochlea.
-The model consists of a gammatone filterbank followed by half-wave rectification and non-linear compression.
-Finally, the model assumes a 1 kHz low-pass filter modeling the response characteristics of the inner hair cells.
+The model consists of a gammatone filterbank followed by half-wave rectification and non-linear compression with an exponent of 0.4.
+Finally, the model includes a 1 kHz low-pass filter modeling the response characteristics of the inner hair cells.
 In the current implementation, we used a gammatone filterbank with xx filters, centered around xx kHz.
 The filters were spaced to correspond equivalent rectangular bandwidths \citep{Glasberg1990}.
-The ouput of the channels were filtered with a time (distance) weighted factor corresponding to the atmospheric attenuation at the channel's center frequency.
+The output of each channel was filtered with a time (distance) weighted factor corresponding to the atmospheric attenuation at the channel's center frequency.
 This filtering approximated frequency-dependent atmospheric attenuation.
@@
 It also illustrates the resulting response for each of the xx channels of the \citet{Wiegrebe2008} model.
-From these responses, it can be inferred that these are highly correlated.
+From these responses, it can be seen that these are highly correlated across channels.
 This is confirmed when analyzing the average cross-channel correlation for all xx synthetic echo sequences (Fig.
@@
 From these high cross-channel correlations, we conclude that the output of the cochlear model can be summarized by averaging across channels without extensive loss of information.
+On average, xx\% of variation in the output of the frequency channels was explained by the average across channels.
 An example of such a cross-channel sum is plotted in figure \ref{fig:processing}xx.
@@
 	\caption{Example of synthetic echo sequence and its processing using the \citet{Wiegrebe2008} model.
-(xx) Example of a single synthetic echo sequence.
+(xx) Example of a single synthetic echo sequence (red).
+The dechirped version is plotted in black.
 (xx-xx) Response of the xx frequency channels in the \citet{Wiegrebe2008} model.
-From these plots, it can be seen that channels highly correlate.
+From these plots, it can be seen that channels correlate highly.
 This is confirmed in panel xx, depicting the average correlation across channels for all xx synthetic echo sequences analyzed.
+Note that the x-scale does not start at zero.
 (xx) The output of the \citet{Wiegrebe2008} model summed across the frequency channels, is referred to as a template in this paper.
@@
 We used \PCA\ to assess the dimensionality of the templates generated from the synthetic echo sequences.
-We found that almost all of the variance in the templates could be captured by \pca\ components (see figure \ref{fig:processing}xx).
-Some examples of templates reconstructed from the principal components are shown in \ref{fig:synthetictemplates}xx).
+We found that almost all of the variance ($> 99\%$) in the templates could be captured by \pca\ components (see figure \ref{fig:processing}xx).
+Some examples of templates reconstructed from the principal components are shown in figure \ref{fig:synthetictemplates}xx).
 
@@
 	\includegraphics[width=1\linewidth]{../results/synthetic_templates}
-	\caption{Example of synthetic echo sequences, their corresponding templates and the templates reconstructed from the \pca\ \pcs.
+	\caption{Example of synthetic dechirped echo sequences, their corresponding templates and the templates reconstructed from the \pca\ \pcs.
 (xx-xx) Three examples of synthetic echo sequence waveforms.
@@
 
-As we did for the synthetic echoes sequences, the echoes trains collected in bat habitats were filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}.
+As we did for the synthetic echoes sequences, the echoes trains collected in bat habitats were dechirped and filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}.
 We concluded from analyzing the synthetic echoes that the response in different frequency channels is highly correlated.
@@
 
-We wished to assess whether the templates projected into the \pca-dimensional space still contained sufficient information to be recognized.
+We assessed whether the templates projected into the \pca-dimensional space still contained sufficient information to be recognized.
 Previously, we assessed whether templates can be used to recognize locations using, what \citet{Baddeley2012} called, a perfect memory approach.
@@
 
-In our previous work \cite{Vanderelst2016}, we showed that the templates could be used to recognize locations in bat habitats.
+In our previous work \citep{Vanderelst2016}, we showed that the templates could be used to recognize locations in bat habitats.
 As we have shown that the templates can be projected into a lower \pca D feature space without loss, it can be expected that these \pca\ features can be used to locations as well.
@@
 
+\todo[inline]{update this to reflect latest results}
 Interestingly, for azimuth and elevation, classification errors were not uniform across directions.
@@
 In that study, the authors adapted the Meddis inner hair cell model \citep{Meddis2006} to assess the amount of information in the acoustic signal preserved by the cochlea.
-These authors concluded that spectral notches introduced by overlapping echoes are faithfully encoded by auditory nerve
-spike patterns.
-However, quantifying the information transfer rate also found that about half of the information (bits) are lost in the transduction from acoustic energy to spike trains.
+These authors concluded that spectral notches introduced by overlapping echoes are faithfully encoded by auditory nerve spike patterns.
+However, quantifying the information transfer rate, they also found that about half of the information (in terms of bits) are lost in the transduction from acoustic energy to spike trains.
 In the current paper, we expand on this work and quantify the dimension reduction due to cochlear encoding.
@@
 We also generated templates using echo sequences of which the samples were randomly shuffled.
-This allows inferring whether the low-dimensionality of the templates is a property of the echo signals (before being filtered through the cochlear model).
+This allows inferring whether the low-dimensionality of the templates is a property of the echo signals (before being filtered by the cochlear model).
 
 Figure \ref{fig:mechanism} shows the number of \pcs\ needed to capture 99\% of the templates variance.
-From this plot, it can be seen that the number of \pcs\ increases when shuffling the echo sequences increases from xx to xx.
+From this plot, it can be seen that the number of \pcs\ increases when shuffling the echo sequences from xx to xx.
 Omitting the 1 kHz filter from the  \citet{Wiegrebe2008} model increases the number of \pcs\ to xx.
@@
 \ref{fig:mechanism}, inset) shows that most energy in the templates is contained in frequencies below 1 kHz.
-Therefore, it ought to be possible to completely represent templates by sampling them at 2 kHz,  or about 70 samples per template, as templates consist of 7499 samples at 219 kHz sample rate.
-This corresponds to the number of \pcs\ required to represent templates derived from echo sequences that were shuffled  (fig.
+Therefore, if we take 1 kHz as the highest relevant frequency, it ought to be possible to completely represent templates by sampling them at 2 kHz.
+The corresponds to about 70 samples per template, as templates consist of 7499 samples at 219 kHz sample rate.
+This corresponds roughly to the number of \pcs\ required to represent templates derived from echo sequences that were shuffled  (fig.
 \ref{fig:mechanism}).
@@
 Therefore, we conclude that the 1 kHz filter is the main, but not exclusive, locus of the dimensionality reduction.
-The structure of the echo signals also increases the dimensionality of the templates.
+The structure of the echo signals also decreases the dimensionality of the templates.
+Therefore, we conclude that the sonar system of bats, through a number of filtering stages, compresses the 3D complexity of the world onto a lower dimensional representation.
 
@@
 Here, we directly demonstrated that the \pca\ principal components are sufficient to discriminate their associated locations using neural networks.
+The neural networks' performance was somewhat that the perfect memory approach we used previously (figs.
+\ref{fig:israelperformance}, \ref{fig:royalperformance}).
 
-Here, we show that a low dimensional signal is sufficient for recognizing environments.
+We showed that a low dimensional signal is sufficient for recognizing environments.
 This is a critical component of navigation.
 However, bats do more than navigating.
-They also use sonar to control flight, for example, avoid obstacles or following guiding structures.
-The description the bat has access to for flight control has the same low-level character than that used for place recognition.
-Therefore, our work suggests that this low-level description is sufficient for flight control as well.
-We have shown before that a low-level description could be sufficient for obstacle avoidance and corridor following \citep{Vanderelst2015a,Mansour2019}.
-However, in that work, we converted both the echoes at the left and at the right ear to template-like representation.
+They also rely on sonar for flight control.
+This includes, avoiding obstacles or following guiding structures.
+The environmental description the bat has access to for flight control has the same low-level character than that used for place recognition.
+Therefore, our current work suggests that this low-level description should be sufficient for flight control as well.
+Indeed, we have previously demonstrated that a low-dimensional description could be sufficient for obstacle avoidance and corridor following \citep{Vanderelst2015a,Mansour2019}.
+In that work, we converted both the echoes at the left and at the right ear to template-like representation.
 Subsequent work will have to demonstrate that the parameters of the \pca D space presented here can be used directly for controlling flight.
@@
 
-Despite their small size, the networks in this paper were able to encode (memorize) the templates as least as well as the perfect memory approach.
+Despite their small size, the networks in this paper were able to encode (memorize) the templates at least as well as the perfect memory approach.
 The neural networks presented here are tiny compared to the density of cortical matter.
@@
 Table \ref{tab:sizes} lists the number of weights for the artificial neural networks and compares this number with the size of the \pca-dimensional templates.
-The number of floats making up the templates is larger than the number of weights by a factor of 2-3.
+The number of floating point numbers making up the templates is larger than the number of weights by a factor of 2-3.
 Therefore, the networks have a smaller memory footprint than the data they encode.
 It should be noted that we have not made an attempt to try and to optimize (minimize) the size of the networks.
+From these numbers, we conclude that the low-dimensional signals provided by the bat's sonar system result in a low memory footprint and that they would use little neural resources to be stored and processed.
 
@@
 
-From our current work, in combination with previous results \citep{Vanderelst2016,Vanderelst2015a,Mansour2019}, we conclude that the auditory periphery of bats maps complex echoes to a limited number of features.
+It has been suggested before that bats' sonar systems provides detailed information about the world.
+A number of authors have referred to sonar as providing the bat with an acoustic image \citep{Lee2017,Barchi2013,Moss2001,Schnitzler2003,Simmons2012,Ulanovsky2008,Clare2015,Surlykke2016,Geipel2013}.
+As mentioned above, this is at odds with some recent behavioral and neurophysiological findings in bats faced with complex echo sequences \citep{Knowles2015,Geberl2013,Warnecke2018}.
+However, it is also at odds with the various filtering stages involved in the emission and the reception of bio-sonar.
+In particular, we argued before that the neurophysiology of the  bats auditory system represents a bottleneck in the amount of information accessible to the bat.
+Indeed, \citet{Reijniers2010a} already noticed that significant amounts of information get lost during cochlear transduction.
+Here, we followed up on that work to quantify the dimensional reduction of the input signals as a result of auditory processing.
+
+From our current work, in combination with previous results \citep{Vanderelst2016,Vanderelst2015a,Mansour2019,Reijniers2010a}, we conclude that the sonar system of bats maps complex echoes to a limited number of features.
 These features can be efficiently stored to recognize scenes.
-Moreover, we conjecture that a (binaraural) version of the features should allow for sonar-based flight control.
-As such, a view of echolocating bats living in a low dimensional sensory space emerges.
+Moreover, we conjecture that a (binaural) version of the features should allow for sonar-based flight control.
+As such, a view of echolocating bats living in a low dimensional sensory space emerges that contradicts existing claims of sonar providing acoustic images.
 

### 2020-03-23 Paper/main_hp.tex
Herberts edits

--- 2020-03-23 Paper/main.tex
@@
 However, as we argued before, it is highly questionable whether bats can interpret echoes in terms of a 3D environment model or acoustic image \citep[e.g.,][]{Vanderelst2015,Vanderelst2016,Steckel2013}.
-Sonar maps the three-dimensional structure of the environment on two one-dimensional waveforms arriving at the ears.
+Sonar projects the three-dimensional structure of the environment on two one-dimensional waveforms arriving at the ears.
 While the time of arrival of echoes encodes object distance, azimuth and elevation directions are not directly available from the acoustic signal.
@@
 
-Low dimensional descriptions potentially contain a lot of information \citep{Kuc1997b,Kuc1997}.
+However, it should be noted that such low dimensional descriptions have been shown to potentially contain a lot of useful information \citep{Kuc1997b,Kuc1997}.
 We previously demonstrated that bats might use sonar template-based scene recognition \citep{Vanderelst2016,Vanderelst2017}.
@@
 The physics of the sonar system and the auditory periphery map the complexity of the 3D environment onto a small number of independent features, irrespective of whether we use artificial or real echo sequences.
-Nevertheless, this limited set of features is sufficiently rich for bats to recognize locations, and as we discuss, support flight control.
+Nevertheless, this limited set of features is sufficiently rich for bats to recognize locations, and as we discuss, could also support flight control.
 
@@
 
-Each set of points was converted to an impulse response with a sampling frequency $f_s$ of 219 kHz.
-The impulse response length was set to 7499 samples.
+Each set of points was converted into an environmental impulse response with a sampling frequency $f_s$ of 219 kHz.
+The environmental impulse response length was set to 7499 samples.
 These parameters correspond to the recording length and the sampling frequency used in the acoustic data collection (see below).
-In converting the point to an impulse response, we modeled attenuation of the echoes due to spherical spreading (both ways).
-We sampled the amplitude of the impulses from a normal distribution with a standard deviation of 10 dB.
+In representing each point reflector by a Dirac impulse, we modeled attenuation of the echoes by introducing an amplitude calculated based on spherical spreading (both ways).
+The final amplitude of each Dirac impulse was determined by sampling from a normal distribution centered on the calculated amplitude and with a standard deviation of 10 dB.
+This final operation represents a very simple model of varying reflector strength due e.g.
+to shape and orientation of the reflecting feature.
 
@@
 \begin{equation}\label{key}
-E_d \mathcal{F}^{-1}(\mathcal{F}(S)^* . \mathcal{F}(E))
+E_d = \mathcal{F}^{-1}((\mathcal{F}(S))^* . \ \mathcal{F}(E))
 \end{equation}
@@
 In the current implementation, we used a gammatone filterbank with xx filters, centered around xx kHz.
-The filters were spaced to correspond equivalent rectangular bandwidths \citep{Glasberg1990}.
-The output of each channel was filtered with a time (distance) weighted factor corresponding to the atmospheric attenuation at the channel's center frequency.
-This filtering approximated frequency-dependent atmospheric attenuation.
+The filters were spaced to correspond to equivalent rectangular bandwidths \citep{Glasberg1990}.
+The output of each channel was attenuated with a time (distance) weighted factor corresponding to the atmospheric attenuation at the channel's center frequency.
+This operation approximated frequency-dependent atmospheric attenuation.
 The first 6 ms of data were removed.
-This was done to ensure the data corresponded with the acoustic data.
-As indicated below, we removed the first 6 ms of those data to avoid including the pickup of the emitted signal.
+This was done to ensure the data corresponded with the measurement data.
+As indicated below, we removed the first 6 ms of the measurement  data to avoid including the pickup of the emitted signal.
 
@@
 It also illustrates the resulting response for each of the xx channels of the \citet{Wiegrebe2008} model.
-From these responses, it can be seen that these are highly correlated across channels.
+From these responses, it can be seen that they are highly correlated across channels.
 This is confirmed when analyzing the average cross-channel correlation for all xx synthetic echo sequences (Fig.
@@
 
-We used \PCA\ to assess the dimensionality of the templates generated from the synthetic echo sequences.
+We used principal components analysis (\PCA\ ) to assess the dimensionality of the templates generated from the synthetic echo sequences.
 We found that almost all of the variance ($> 99\%$) in the templates could be captured by \pca\ components (see figure \ref{fig:processing}xx).
@@
 Panels xx-xx depict the resulting templates (i.e., summation across the output channels of the \citet{Wiegrebe2008} model).
-These panels also show the result of projecting the templates into a \pca D space and transforming the result back to the original coordinates.}
+These panels also show the result of projecting the templates into a \pca D \pc\ space and transforming the result back to the original coordinates.}
 	\label{fig:synthetictemplates}
@@
 Using this model, we have derived a low dimensional space ($D = \pca$), which can be used to describe the templates resulting from the artificial echo sequences.
-This implies that, after filtering through (a model of) the bat's auditory system, these artificial echoes can be converted to low dimensional descriptions.
+This implies that, after filtering through (a model of) the bat's auditory system, these artificial echoes can be converted into low dimensional descriptions.
 The original echo sequences, generated by xx to xx point reflectors effectively map to \pca\ dimensional vector.
@@
 The environment consisted of a walkway lined with rocks and vegetation.
-Unpublished bat monitoring data at the site confirmed that bat used this corridor to commute.
+Unpublished bat monitoring data at the site confirmed that bats used this corridor to commute.
 The ensonification device was placed at 50 positions along a straight line spanning a total of 10 meters Locations were spaced 20 cm apart.
@@
 
-As we did for the synthetic echoes sequences, the echoes trains collected in bat habitats were dechirped and filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}.
+As we did for the synthetic echo sequences, the echo trains collected in bat habitats were dechirped and filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}.
 We concluded from analyzing the synthetic echoes that the response in different frequency channels is highly correlated.
-Therefore, the output of the cochlear model was averaged across frequencies to obtain a time-intensity profile, or template.
-The beam of the emitter used was more narrow than the typically combined hearing and emission directionality of bats \citep{Vanderelst2010a,Reijniers2010,Jakobsen2012}.
-Therefore, templates were averaged across three neighboring directions.
-The first 6 ms of data were removed to avoid using emitted signals for analysis.
+Therefore, the output of the cochlear model was averaged across frequency channels to obtain a time-intensity profile, or template.
+As the beam of the emitter used was more narrow than the typically combined hearing and emission directionality of bats \citep{Vanderelst2010a,Reijniers2010,Jakobsen2012} templates were averaged across three neighboring directions.
+The first 6 ms of data were removed to avoid using the emissions themselves for analysis.
 The processing steps outlined above resulted in 3 $\times$ 217 templates, for each location at each site, corresponding to 217 directions and three repeats.
 
-Among the sets of templates, some had very little energy in them, notably those collected for directions without reflectors.
+Among the set of templates, some had very little energy in them, notably those collected for directions without reflectors.
 For example, aiming the ensonification device upwards often resulted in no echoes returning.
@@
 
-Internal (and, potentially, external) noise results in template values greater than zero, even in the absences of reflectors returning echoes.
+Internal (and, potentially, external) noise results in template values greater than zero, even in the absence of reflectors returning echoes.
 To avoid training the neural network on this noise, we estimated the noise floor in our data by collecting templates while pointing the device to empty space.
@@
 Any template values below this value were set to this value.
-As we argued before \citep{Vanderelst2016}, using the empirical measurement noise to estimate $\noisefloor$ is biologically warranted as bats have a better signal to noise ratio in their auditory systems compared to our measuring device.
+As we argued before \citep{Vanderelst2016}, using the empirical measurement noise to estimate $\noisefloor$ is biologically warranted as bats have a better signal to noise ratio in their auditory system compared to our measuring device.
 Indeed, their dynamic range is at least 30 \db\ larger than our measurement device.
@@
 The Euclidean distance between two echo templates was the metric used to assess how well bats should be able to distinguish between templates.
-This table lookup and distance calculation require a comprehensive search of memory.
-Hence, in our previous work, we established that templates could be classified, but this required a significant amount of memory and computational resources.
-
-Here, we used a more biologically plausible memory system: an artificial neural network.
-These are a neurally inspired way of distributed information storage in artificial neural networks \citep{Mcleod1998}.
-However, we acknowledge that current training methods most likely have no biological counterpart.
+This table lookup and distance calculation requires a comprehensive search of memory that scales with the number of templates memorized.
+Hence, in our previous work, we established the theoretical possibility of classifying templates based on real echoes from environments bats were known to frequent, irrespective of the significant amount of memory and computational resources required to do so.
+Here, we use a more biologically plausible memory system: an artificial neural network, i.e.
+a neurally inspired way of distributed information storage \citep{Mcleod1998}.
+While we acknowledge that current neural network training methods most likely have no biological counterpart, we propose that this approach provides us with a more biologically plausible appraisal of the memory and computational effort required to implement a similar scheme in a bat brain.
 
@@
 The first layer of the neural network added Gaussian noise to each input sample ($\sigma(\stochasticnoise) = 0.01$).
-The following layers were completely connected and used the rectified linear activation function.
-The output layer used a softmax activation function (or the normalized exponential function), normalizing the output into a probability distribution consisting of probabilities proportional to the exponentials of the output layer's activation.
+The following layers were fully connected and used the rectified linear activation function.
+The output layer uses a softmax activation function, i.e.
+the normalized exponential function.
 The loss was calculated as the cross-entropy between the target output and the actual output of the neural network.
@@
 Training a different neural network for each of these features ensured that the training algorithm paid equal attention to each feature.
-When training a single network for all features, the optimizer could preferably reduce errors in the location predictions as this feature has the largest variance.
+When training a single network for all features, the optimization procedure would preferably reduce errors in the location predictions as this feature has the largest variance.
 The full architecture can be considered as a single network with different streams for each feature (fig.
@@
 Figure \ref{fig:empiricaltemplates} shows examples of templates and their reconstructed counterparts.
-Therefore, although the model used to generate echo sequences was very simple, the resulting underlying \pca D feature space could adequately represent templates resulting from real acoustic data.
+Therefore, although the model used to generate echo sequences was very simple, these results indicate that the \pca D feature space derived from the generative echo model could adequately represent templates resulting from real acoustic data.
 
@@
 In our previous work \citep{Vanderelst2016}, we showed that the templates could be used to recognize locations in bat habitats.
-As we have shown that the templates can be projected into a lower \pca D feature space without loss, it can be expected that these \pca\ features can be used to locations as well.
+As we have shown that the templates can be projected into a lower \pca D feature space without loss of information, it can be expected that these \pca\ features can be used to recognize locations as well.
 We found this to be indeed the case.
@@
 	\includegraphics[width=1\linewidth]{../results/royal_results.pdf}
-	\caption{Simular as fig \ref{fig:israelperformance}, but for the Royal Fort Park site.}
+	\caption{Classification results for the Royal Fort Park site.
+Top row: cumulative error distributions for the azimuth, elevation and location classification.
+Middle row: Confusion matrices for the classifications.
+Bottom row: remaining classification entropy for each target (derived from the confusion matrices).}
 	\label{fig:royalperformance}
@@
 In that study, the authors adapted the Meddis inner hair cell model \citep{Meddis2006} to assess the amount of information in the acoustic signal preserved by the cochlea.
-These authors concluded that spectral notches introduced by overlapping echoes are faithfully encoded by auditory nerve spike patterns.
+The authors concluded that spectral notches introduced by two overlapping echoes are faithfully encoded by auditory nerve spike patterns.
 However, quantifying the information transfer rate, they also found that about half of the information (in terms of bits) are lost in the transduction from acoustic energy to spike trains.
@@
 We found that the same \pca D space could be used to describe templates originating from echoes collected in real bat habitats (figure \ref{fig:empiricaltemplates}).
-This suggests the filtering of the auditory periphery provides the bat with a greatly reduced acoustic description of its environment.
+This suggests that the filtering of the auditory periphery provides the bat with a significantly compressed acoustic description of its environment.
 To identify the locus of the compression, we performed a number of additional analyses.
-In particular, we evaluated whether the low-dimensional description is a property of the echo sequences, the gammatone filters, or the 1 kHz low-pass filter in the \citet{Wiegrebe2008} model.
+In particular, we evaluated whether the low-dimensional description is a property of the echo sequences themselves, or whether this property results from the finite bandwidths of the gammatone filters and/or the low-pass filter in the \citet{Wiegrebe2008} model.
 
@@
 \ref{fig:mechanism}, inset) shows that most energy in the templates is contained in frequencies below 1 kHz.
-Therefore, if we take 1 kHz as the highest relevant frequency, it ought to be possible to completely represent templates by sampling them at 2 kHz.
-The corresponds to about 70 samples per template, as templates consist of 7499 samples at 219 kHz sample rate.
+Therefore, if we take 1 kHz as the highest relevant frequency, it ought to be possible to completely represent templates by sampling them at 2 kSamples/sec.
+This corresponds to about 70 samples per template, as templates consist of 7499 samples at 219 kHz sample rate.
 This corresponds roughly to the number of \pcs\ required to represent templates derived from echo sequences that were shuffled  (fig.
@@
 Despite the low dimensional character of the templates, they are sufficient to recognize environments.
-Indeed, this was demonstrated in our previous work \citep{Vanderelst2016}.
-Here, we directly demonstrated that the \pca\ principal components are sufficient to discriminate their associated locations using neural networks.
+Indeed, this was already demonstrated in our previous work \citep{Vanderelst2016}.
+Here, we directly demonstrated that the \pca\ principal components are sufficient to discriminate the associated locations and viewing directions using neural networks.
 The neural networks' performance was somewhat that the perfect memory approach we used previously (figs.
@@
 
-We showed that a low dimensional signal is sufficient for recognizing environments.
-This is a critical component of navigation.
-However, bats do more than navigating.
+While recognizing location in the environment is a critical component of navigation, bats use their sonar not only for navigating.
 They also rely on sonar for flight control.
-This includes, avoiding obstacles or following guiding structures.
-The environmental description the bat has access to for flight control has the same low-level character than that used for place recognition.
-Therefore, our current work suggests that this low-level description should be sufficient for flight control as well.
-Indeed, we have previously demonstrated that a low-dimensional description could be sufficient for obstacle avoidance and corridor following \citep{Vanderelst2015a,Mansour2019}.
-In that work, we converted both the echoes at the left and at the right ear to template-like representation.
-Subsequent work will have to demonstrate that the parameters of the \pca D space presented here can be used directly for controlling flight.
+This includes, avoiding obstacles and/or following guiding structures.
+The environmental description the bat has access to for flight control starts from the same cochlear output  used for place recognition.
+Therefore, our current work suggests that the same low-dimensional description should be sufficient for flight control as well.
+Indeed, we have previously demonstrated that a low-dimensional description is sufficient for obstacle avoidance and corridor following \citep{Vanderelst2015a,Mansour2019}.
+In that work, we converted both the echoes at the left and at the right ear into a template-like representation used as input to the motion controller.
+Subsequent work will have to demonstrate whether the parameters of the \pca D space presented here can be used directly for controlling flight.
 In other words, we conjecture it should be possible to extract \pca\ features from an echo signal and convert this vector into a motor command.
-One can presume a neural network that converts the \pca\ features to flight control parameters.
+One could presume a neural network that converts the \pca\ features to flight control parameters.
 
@@
 
-We were unable to find numbers on the number of neurons in the cortex of bat species.
+We were unable to find values for the number of neurons in the cortex of bat species.
 However, the mouse cortex has been estimated to contain about $9 \cdot 10^4$ neurons or $7 \cdot 10^8$ synapses per cubic millimeter \citep{Braitenberg2013}.
 The neural networks presented in this paper consisted of about \nneurons\ neurons and \nweights\ connections.
-Therefore, in as much as these neural networks can be considered as biologically plausible models of memory, they are orders of magnitude smaller than a single cubic millimeter of mouse cortex.
+Therefore, to the extent that the proposed neural networks can be considered as biologically plausible models of memory, they are orders of magnitude smaller than a single cubic millimeter of mouse cortex.
 Let alone the complete cortex, which has been estimated to contain 75 million neurons \citep{Williams2000}.
@@
 However, it is also at odds with the various filtering stages involved in the emission and the reception of bio-sonar.
-In particular, we argued before that the neurophysiology of the  bats auditory system represents a bottleneck in the amount of information accessible to the bat.
-Indeed, \citet{Reijniers2010a} already noticed that significant amounts of information get lost during cochlear transduction.
-Here, we followed up on that work to quantify the dimensional reduction of the input signals as a result of auditory processing.
+In particular, we argued before that the neurophysiology of the  bats auditory system represents a bottleneck in the amount of raw echo signal information accessible to the bat.
+Indeed, \citet{Reijniers2010a} already noticed that significant compression occurs during cochlear transduction.
+Here, we followed up on that work to quantify the dimensional reduction of the input signals as a result of peripheral auditory processing.
 
-From our current work, in combination with previous results \citep{Vanderelst2016,Vanderelst2015a,Mansour2019,Reijniers2010a}, we conclude that the sonar system of bats maps complex echoes to a limited number of features.
+From our current work, in combination with previous results \citep{Vanderelst2016,Vanderelst2015a,Mansour2019,Reijniers2010a}, we conclude that the sonar system of bats maps complex echo signals onto a limited number of features.
 These features can be efficiently stored to recognize scenes.
 Moreover, we conjecture that a (binaural) version of the features should allow for sonar-based flight control.
-As such, a view of echolocating bats living in a low dimensional sensory space emerges that contradicts existing claims of sonar providing acoustic images.
+As such, a view of echolocating bats living in a low dimensional sensory space emerges that is at odds with the widespread notion of sonar as a sensing modality that provides detailed acoustic images.
 

### 2020-03-24 RUN_ALL.py
The interpolation thing works

@@ -1,11 +1,23 @@
 import script_preprocess_empirical_data
 
+# Get space based on synthetic echoes
 import script_synthetic_echoes_to_PCA
 import script_plot_processing_synthetic
 
+# Visualize reconstruction of syn and emp templates
 import script_plot_template_examples
-import script_plot_components
 
+#Train the neural networks
 import script_run_training
+import script_plot_training_history
 
+# For both israel and royal do...
 import script_run_perfect_memory
+import script_plot_training_results
+
+# Assess where the reduction happens
+import script_run_identify_mechanism
+
+
+import script_tabulate_memory_use
+import script_plot_components

### 2020-03-24 interpolated.py

@@ -3,23 +3,21 @@ import pandas
 import os
 from matplotlib import pyplot
 from tensorflow import keras
+from scipy.signal import convolve
 
 from library import misc, process_functions, settings
 
-#for data_set in ['israel', 'royal']:
+# for data_set in ['royal', 'israel']:
+#     for selected_dimension in ['lcs', 'azs','els']:
+
 for data_set in ['israel']:
-    layers = [50, 75, 75, 50]
-    nr_epochs = 250
-
-    for selected_dimension in ['lcs', 'azs','els']:
+    for selected_dimension in ['azs']:
         file_names = misc.folder_names(data_set, selected_dimension)
-
         # Read prepared data
         data = numpy.load(file_names['npz_file'])
         pca_model = misc.pickle_load(settings.pca_templates_model_file)
 
         # Select inputs
-        if selected_dimension == 'lcs': unencoded_data = data['long_lcs']
         if selected_dimension == 'azs': unencoded_data = data['long_azs']
         if selected_dimension == 'els': unencoded_data = data['long_els']
         if selected_dimension == 'lcs' and data_set in ['israel', 'royal']: unencoded_data = misc.map_lcs_to_distances(data)
@@ -33,42 +31,21 @@ for data_set in ['israel']:
         inputs = pca_model.transform(templates)
         n_components = inputs.shape[1]
 
-        # Make model
-        model = keras.Sequential()
-        noise_layer = keras.layers.GaussianNoise(input_shape=(n_components,), stddev=settings.stochaistic_noise * 1)
-        output_layer = keras.layers.Dense(target_n, activation='softmax')
-        model.add(noise_layer)
-        for nodes in layers: model.add(keras.layers.Dense(nodes, activation='relu'))
-        model.add(output_layer)
+        # Load neural network
+        neural_network = keras.models.load_model(file_names['model_file'])
 
-        # Train Model
-        loss = keras.losses.CategoricalCrossentropy()
-        model.compile('adam', loss=loss)
-        training_history = model.fit(inputs, targets, epochs=nr_epochs)
-        model.save(file_names['model_file'])
-
-        predictions_matrix = model.predict(inputs)
+        predictions_matrix = neural_network.predict(inputs)
         binary_predictions = misc.binarize_prediction(predictions_matrix)
         interpreted_predictions = encoder.inverse_transform(binary_predictions)
         interpreted_predictions = interpreted_predictions.flatten()
 
[…]

### 2020-03-24 library/misc.py

@@ -161,7 +161,7 @@ def folder_names(data_set, dimension):
     if dimension is None: base_name = data_set
 
     npz_file = os.path.join(result_folder, data_set + '.npz')
-    pca_file = os.path.join(result_folder, data_set + '.pca')
+    npz_file_interpolated = os.path.join(result_folder, data_set + '_interpolated_directions.npz')
 
     model_file = os.path.join(result_folder, base_name + '.5h')
     res_file = os.path.join(result_folder, base_name + '.pd')
@@ -175,7 +175,8 @@ def folder_names(data_set, dimension):
     result['result_folder'] = result_folder
     result['log_folder'] = log_folder
     result['npz_file'] = npz_file
-    # call_result['pca_file'] = pca_file
+    result['npz_file_interpolated'] = npz_file_interpolated
+
     result['model_file'] = model_file
     result['results_file'] = res_file
     result['history_file'] = hist_file

### 2020-03-24 library/process_functions.py

@@ -6,25 +6,20 @@ import numpy
 import scipy.io as io
 from scipy.signal import convolve
 from sklearn.preprocessing import OneHotEncoder
+from scipy.interpolate import RegularGridInterpolator
 
 
 def preprocess(file_name, verbose=True):
     fs = settings.sample_frequency
     noise_floor = settings.noise_floor
-    integration_time = settings.integration_time
     initial_zero_time = settings.initial_zero_time
     raw_collected_samples = settings.raw_collected_samples
-
     initial_omission_samples = math.ceil(initial_zero_time * fs)
-    integration_samples = math.ceil(fs * integration_time)
-    #final_samples = math.ceil(raw_collected_samples / integration_samples)
 
     if verbose:
         print('+-' * 10 + '+')
         print('PREPROCESSING', file_name)
-        #print('integration samples', integration_samples)
         print('initial zero samples', initial_omission_samples)
-        #print('final samples', final_samples)
 
     data = io.loadmat(file_name)
     data = data['Templates']
@@ -32,9 +27,7 @@ def preprocess(file_name, verbose=True):
     azimuth = data[0, 0][1]
     elevation = data[0, 0][2]
 
-    #processed = numpy.zeros((7, 31, final_samples, 3))
     processed = numpy.zeros((7, 31, raw_collected_samples, 3))
-
     for i in range(3):
         repetition = templates[:, :, i]
         # Reshape the data
@@ -60,31 +53,20 @@ def preprocess(file_name, verbose=True):
         mask = mask / numpy.sum(mask)
         mn_box = convolve(mn_box, mask, mode='same')
 
-        # # Average across directions and time
-        # mask = numpy.ones((3, 3, integration_samples))
-        # mask = mask / numpy.sum(mask)
-        # mn_box = convolve(mn_box, mask, mode='same')
-        # # Subsample
-        # mn_box = mn_box[:, :, ::integration_samples]
-
         if verbose: print('final samples', mn_box.shape[2])
-
-
         processed[:, :, :, i] = mn_box
         processed[processed < noise_floor] = noise_floor
     return processed, az_box, el_box
 
 
-def process_data_set(data_set, filter_threshold = 0.1):
-    # %%Preprocess andrews
+def process_data_set(data_set, filter_threshold = 0.1, interpolate_directions=False):
     print('#' * 10)
[…]

### 2020-03-24 script_plot_processing_synthetic.py

@@ -7,7 +7,7 @@ from matplotlib import pyplot
 from library import misc
 from library import generate_functions
 from library import settings
-from pyBat import Wiegrebe
+from pyBat import Wiegrebe, Signal
 
 sample_frequency = settings.sample_frequency
 number_of_samples = settings.raw_collected_samples
@@ -16,17 +16,15 @@ emission_frequency_mean = settings.emission_frequency_mean
 initial_zero_samples = math.ceil(initial_zero_time * sample_frequency)
 
 emission = generate_functions.create_emission()
-
-emission_padding = number_of_samples - emission.shape[0]
-padded_emission = numpy.pad(emission, (0, emission_padding), 'constant')
-
-wiegrebe = Wiegrebe.ModelWiegrebe(sample_frequency, emission_frequency_mean, 4, emission=padded_emission)
+wiegrebe = Wiegrebe.ModelWiegrebe(sample_frequency, emission_frequency_mean, 4)
 
 generative_model_parameters = {}
 generative_model_parameters['n_seed_points'] = [100]
 generative_model_parameters['n_cloud_points'] = [100]
 distances = generate_functions.generative_model(generative_model_parameters)
-echo_sequence, impulse_response = generate_functions.distances2echo_sequence(distances, emission)
+echo_sequence_original, impulse_response = generate_functions.distances2echo_sequence(distances, emission)
+echo_sequence = Signal.dechirp(emission, echo_sequence_original)
+
 echo_sequence[0:initial_zero_samples] = 0
 template = generate_functions.echo_sequence2template(echo_sequence, wiegrebe)
 
@@ -48,16 +46,20 @@ synthetic_template_pca = misc.pickle_load(settings.pca_templates_model_file)
 cumulative = numpy.cumsum(synthetic_template_pca.explained_variance_ratio_)
 
 # %%
+
+r = numpy.max(numpy.abs(echo_sequence_original))
+
 pyplot.style.use(settings.style)
 
 fig = pyplot.figure(figsize=(20, 10))
-
 gs = fig.add_gridspec(8, 3)
+
 ax = fig.add_subplot(gs[:, 0])
+ax.plot(echo_sequence_original, color=settings.default_line_color02)
 ax.plot(echo_sequence, color=settings.default_line_color01)
[…]

### 2020-03-24 script_plot_template_examples.py

@@ -92,6 +92,7 @@ pyplot.xticks([])
 pyplot.yticks([])
 pyplot.legend(['Template', 'Reconstructed'], frameon=False)
 pyplot.ylabel('Template amplitude')
+pyplot.xlabel('Time/Distance')
 misc.label(0.1, 0.9, 3, fontsize=14)
 
 print('-----> PLOTTING 7')
@@ -101,6 +102,7 @@ pyplot.plot(reconstructed_template_1, color=settings.default_line_color02, alpha
 pyplot.ylim([mn_templates, mx_templates])
 pyplot.xticks([])
 pyplot.yticks([])
+pyplot.xlabel('Time/Distance')
 misc.label(0.1, 0.9, 4, fontsize=14)
 
 print('-----> PLOTTING 8')
@@ -110,6 +112,7 @@ pyplot.plot(reconstructed_template_2, color=settings.default_line_color02, alpha
 pyplot.ylim([mn_templates, mx_templates])
 pyplot.xticks([])
 pyplot.yticks([])
+pyplot.xlabel('Time/Distance')
 misc.label(0.1, 0.9, 5, fontsize=14)
 
 pyplot.tight_layout()

### 2020-03-24 script_plot_training_history.py

@@ -26,7 +26,7 @@ for data_set in ['israel', 'royal']:
 
         pyplot.plot(trace, color=color)
         pyplot.ylim([0, 4])
-        pyplot.xscale('symlog')
+        #pyplot.xscale('symlog')
         pyplot.xlabel('Epoch')
         if data_set == 'israel': pyplot.ylabel('Loss')
         if data_set == 'israel': pyplot.title('Israel Data')

### 2020-03-24 script_preprocess_empirical_data.py

@@ -1,4 +1,29 @@
-from library import misc, process_functions
+from library import process_functions, misc
+import numpy
 
-process_functions.process_data_set('royal')
+#process_functions.process_data_set('royal')
 process_functions.process_data_set('israel')
+
+#process_functions.process_data_set('royal', interpolate_directions=True)
+process_functions.process_data_set('israel', interpolate_directions=True)
+
+#%%
+from matplotlib import pyplot
+
+data_set = 'israel'
+file_names = misc.folder_names(data_set, None)
+
+
+data1 = numpy.load(file_names['npz_file'])
+data1 = data1['long_data']
+print(data1.shape)
+
+data2 = numpy.load(file_names['npz_file_interpolated'])
+data2 = data2['long_data']
+print(data2.shape)
+
+i = 250
+
+pyplot.plot(data1[i,:])
+pyplot.plot(data2[i,:])
+pyplot.show()

### 2020-03-24 script_run_identify_mechanism.py

@@ -68,7 +68,7 @@ pyplot.style.use(settings.style)
 
 pcs = results['pcs']
 labels = results['labels']
-colors = [settings.default_line_color01, settings.default_line_color01, settings.default_line_color02, settings.default_line_color02]
+colors = [settings.default_line_color01, settings.default_line_color02, settings.default_line_color02, settings.default_line_color02]
 linestyles = ['-', '--', '-', '--']
 
 text_labels = ['1 kHz filter', 'No 1 kHz filter', '1 kHz filter\nshuffled', 'No 1 kHz filter\nshuffled']
@@ -81,7 +81,7 @@ pyplot.bar([0, 1, 2, 3], pcs, color=colors)
 pyplot.xticks([0, 1, 2, 3], text_labels)
 pyplot.ylabel('Number of components')
 pyplot.grid(False)
-pyplot.ylim(0, 1500)
+pyplot.ylim(0, 500)
 
 for i, v in enumerate(pcs):
     pyplot.text(i, v + 10, str(v), color='black', fontweight='bold', horizontalalignment='center')
@@ -91,7 +91,7 @@ inset_axes = inset_axes(pyplot.gca(),
                     width="30%", # width = 30% of parent_bbox
                     height="20%",
                     loc=2,
-                    borderpad=1)
+                    borderpad=2)
 
 
 frequency_axis = numpy.fft.fftfreq(settings.raw_collected_samples, d=1 / 219000) / 1000
@@ -107,6 +107,7 @@ inset_axes.set_xlim(0,5)
 inset_axes.set_xlabel('Freq. (kHz)')
 inset_axes.set_yticks([])
 inset_axes.set_xticks([0,1,2,3,4,5])
+inset_axes.set_title('Av. Template Spectrum',fontsize=10)
 
 # pyplot.subplot(2, 1, 2)
 # frequency_axis = numpy.fft.fftfreq(settings.raw_collected_samples, d=1 / 219000) / 1000

### 2020-03-24 script_run_perfect_memory.py

@@ -18,15 +18,18 @@ if generate_data: process_functions.process_data_set(data_set)
 data = numpy.load(file_names['npz_file'])
 corridor_distances = misc.map_lcs_to_distances(data)
 
-pca = misc.pickle_load(file_names['pca_file'])
+pca = misc.pickle_load(settings.pca_templates_model_file)
+loaded = numpy.load(settings.synthetic_templates_pca_results)
+transformed = loaded['transformed']
+n_components_templates = transformed.shape[1]
+
 
 correct_ids = data['ids']
 correct_ids[:, 0] = corridor_distances
 
 templates = data['long_data']
 pca_templates = pca.transform(templates)
-n_components = settings.n_components_templates
-inputs = pca_templates[:, :n_components]
+inputs = pca_templates[:, :n_components_templates]
 
 # Scale inputs to a minimum of zero
 inputs = inputs - numpy.min(inputs)

### 2020-03-25 Paper/main.tex
updated paper based on herbert comments

@@
 Furthermore, bats have been reported to use sonar-based recognition of landmarks, both under experimental \citep{Jensen2005,Yu2019}, and natural \citep{Verboom1999} conditions.
+Echolocating bats can navigate to salient places \citep{Schnitzler2003}, relying solely on their sonar system.
+Indeed, bats deprived of sight have been found to successfully return to their roost when displaced by up to 40 miles \citep{Stones1969,Williams1966,Davis1970,Mueller1957}.
 
@@
 Together these results suggest that echolocating bats live in a low dimensional perceptual world.
-The physics of the sonar system and the auditory periphery map the complexity of the 3D environment onto a small number of independent features, irrespective of whether we use artificial or real echo sequences.
-Nevertheless, this limited set of features is sufficiently rich for bats to recognize locations, and as we discuss, support flight control.
+The physics of the sonar system and the auditory periphery map the complexity of the 3D environment onto a small number of uncorrelated features, irrespective of whether we use artificial or real echo sequences.
+Nevertheless, this limited set of features is sufficiently rich for bats to recognize locations, and as we discuss, could also support flight control.
 
@@
 Next, for every seed point $s$, we selected $n_{c}$ cloud points normally distributed around $s$, with standard deviation 0.25 m.
+We added a single reflector at a distance between xx and xx to each collection of points.
 This two step process to modeling natural reflectors was inspired by the results of \citet{Yovel2009}, who found that echoes returning from vegetation could be better modeled using a similar two-step process than simply using uniformly spaced points.
@@
 Synthetic echoes consisting of few reflectors are likely to consist of only weak echoes.
-To ensure that all sequences included at least one strong echo, we added a single reflector at a distance between xx and xx to each collection of points.
 
-Each set of points was converted to an impulse response with a sampling frequency $f_s$ of 219 kHz.
-The impulse response length was set to 7499 samples.
+Each set of points was converted into an environmental impulse response with a sampling frequency $f_s$ of 219 kHz.
+The environmental impulse response length was set to 7499 samples.
 These parameters correspond to the recording length and the sampling frequency used in the acoustic data collection (see below).
-In converting the point to an impulse response, we modeled attenuation of the echoes due to spherical spreading (both ways).
-We sampled the amplitude of the impulses from a normal distribution with a standard deviation of 10 dB.
+In representing each point reflector by a Dirac impulse, we modeled attenuation of the echoes by introducing an amplitude calculated based on spherical spreading (both ways).
+The final amplitude of each Dirac impulse was determined by sampling from a normal distribution centered on the calculated amplitude and with a standard deviation of 10 dB.
+This final operation represents a very simple model of varying reflector strength due e.g., to shape and orientation of the reflecting feature.
 
@@
 It also illustrates the resulting response for each of the xx channels of the \citet{Wiegrebe2008} model.
-From these responses, it can be seen that these are highly correlated across channels.
+From these responses, it can be seen that they are highly correlated across channels.
 This is confirmed when analyzing the average cross-channel correlation for all xx synthetic echo sequences (Fig.
 \ref{fig:processing}xx).
+We calculated the correlation $\bar{\rho}$ between the output $X_i$ of each frequency channel $i$ and the mean across channels $\bar{X}$ for each synthetic echo, as follows,
+%
+\begin{equation}
+\bar{\rho} = \frac{1}{n} \sum_i \frac{\mbox{cov}(X_i,\bar{X})}{\sigma_{X_i}, \sigma_{\bar{X}}} 
+\end{equation}
+%
+The distribution of $\bar{\rho}$ is plotted in \ref{fig:processing}xx.
+From this distribution, it can be seen that the value of $\bar{\rho}$ ranged from xx to xx.
 From these high cross-channel correlations, we conclude that the output of the cochlear model can be summarized by averaging across channels without extensive loss of information.
-On average, xx\% of variation in the output of the frequency channels was explained by the average across channels.
+
 An example of such a cross-channel sum is plotted in figure \ref{fig:processing}xx.
@@
 
-In our previous work, the ensonification device was used to ensonify bat habitats at three different sites.
-However, here we only use data collected along two linear pathways through natural habitats.
+Here we use data collected along two linear pathways through natural habitats \citep{Vanderelst2016}.
 At the first site, a public park in Sde Boker, Israel (Fig.
@@
 Overfitting typically reduces the ability of the model to generalize its predictive power to new data.
-However, in this case, the model is not required to generalize.
+However, in this case, the model is not required to generalize to other environments.
 Instead, we are only interested in whether the neural networks can learn and classify the set of training examples.
+Nevertheless, it would be beneficial if the neural networks would be capable of classifying templates corresponding to directions that are slightly different than those included in the data sets.
+Indeed, this is a feature of the Euclidean distance measure used in implementing the perfect memory.
+In that approach, novel templates will be classified to positions with similar templates.
+We verified whether this properties was also found in the neural networks the case by generating a set of templates by interpolating in azimuth and elevation between the directions ensonified in the data set.
+In our data sets, we ensonified 31 different azimuth directions and 7 elevation directions.
+The interpolated data consisted of 30 azimuths and 7 elevations.
+We used the neural networks trained on the original data to classify these interpolated data.
+It was found that the neural network indeed classified these interpolated data to azimuths and elevations close to the positions from which they were interpolated (Figure \ref{fig:interpolation}).
 
@@
 Omitting the 1 kHz filter from the  \citet{Wiegrebe2008} model increases the number of \pcs\ to xx.
-Inspecting the average spectrum of the templates (fig.
+Including the 1 kHz filter, the average spectrum of the templates (fig.
 \ref{fig:mechanism}, inset) shows that most energy in the templates is contained in frequencies below 1 kHz.
-Therefore, if we take 1 kHz as the highest relevant frequency, it ought to be possible to completely represent templates by sampling them at 2 kHz.
-The corresponds to about 70 samples per template, as templates consist of 7499 samples at 219 kHz sample rate.
+Therefore, if we take 1 kHz as the highest relevant frequency, it ought to be possible to completely represent templates by sampling them at 2 kSamples/sec.
+This corresponds to about 70 samples per template, as templates consist of 7499 samples at 219 kHz sample rate.
 This corresponds roughly to the number of \pcs\ required to represent templates derived from echo sequences that were shuffled  (fig.
@@
 	\includegraphics[width=1\linewidth]{../results/mechanism}
-	\caption{}
+	\caption{Barplot depicting the number of required \pcs\ to capture $>99\%$ of the variance in the templates derived from synthetic echoes.
+The label \textit{1 kHz filter} refers to whether the 1 kHz filter was omitted in the implementation of the \citet{Wiegrebe2008} model.
+The label \textit{shuffled} refers to whether the samples of the echoes were randomly shuffled before generating the templates.
+The inset shows the average spectrum of the templates generated using non-shuffled echoes and a version of the \citet{Wiegrebe2008} model including a 1 kHz filter (i.e., the default model).}
 	\label{fig:mechanism}
@@
 Despite the low dimensional character of the templates, they are sufficient to recognize environments.
-Indeed, this was demonstrated in our previous work \citep{Vanderelst2016}.
-Here, we directly demonstrated that the \pca\ principal components are sufficient to discriminate their associated locations using neural networks.
-The neural networks' performance was somewhat that the perfect memory approach we used previously (figs.
+Indeed, this was already demonstrated in our previous work \citep{Vanderelst2016}.
+Here, we directly demonstrated that the \pca\ principal components are sufficient to discriminate the associated locations and viewing directions using neural networks.
+The neural networks' performance was somewhat better than the perfect memory approach we used previously (figs.
 \ref{fig:israelperformance}, \ref{fig:royalperformance}).
 
-We showed that a low dimensional signal is sufficient for recognizing environments.
-This is a critical component of navigation.
-However, bats do more than navigating.
+While recognizing location in the environment is a critical component of navigation, bats use their sonar not only for navigating.
 They also rely on sonar for flight control.
-This includes, avoiding obstacles or following guiding structures.
-The environmental description the bat has access to for flight control has the same low-level character than that used for place recognition.
-Therefore, our current work suggests that this low-level description should be sufficient for flight control as well.
-Indeed, we have previously demonstrated that a low-dimensional description could be sufficient for obstacle avoidance and corridor following \citep{Vanderelst2015a,Mansour2019}.
-In that work, we converted both the echoes at the left and at the right ear to template-like representation.
-Subsequent work will have to demonstrate that the parameters of the \pca D space presented here can be used directly for controlling flight.
+This includes, avoiding obstacles and/or following guiding structures.
+The environmental description the bat has access to for flight control starts from the same cochlear output  used for place recognition.
+Therefore, our current work suggests that the same low-dimensional description should be sufficient for flight control as well.
+Indeed, we have previously demonstrated that a low-dimensional description is sufficient for obstacle avoidance and corridor following \citep{Vanderelst2015a,Mansour2019}.
+In that work, we converted both the echoes at the left and at the right ear into a template-like representation used as input to the motion controller.
+Subsequent work will have to demonstrate whether the parameters of the \pca D space presented here can be used directly for controlling flight.
 In other words, we conjecture it should be possible to extract \pca\ features from an echo signal and convert this vector into a motor command.
-One can presume a neural network that converts the \pca\ features to flight control parameters.
+One could presume a neural network that converts the \pca\ features to flight control parameters.
 
@@
 
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=1\linewidth]{../results/interpolation}
+	\caption{sadasasasdasd}
+	\label{fig:interpolation}
+\end{figure}
+
 \end{document}

### 2020-03-25 Paper/main.tex
filled all the xxs

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand{\bookmark}{}
+
+\newcommand{\MinDistance}{1}
+\newcommand{\MaxDistance}{7}
+
+\newcommand{\MinDistanceFirst}{1}
+\newcommand{\MaxDistanceFirst}{2}
+
+\newcommand{\ns}{[5, 10, 20, 50, 100, 200]}
+\newcommand{\nc}{[5, 10, 20, 50, 100, 200]}
+\newcommand{\MinDiracs}{25}
+\newcommand{\MaxDiracs}{40,000}
+
+\newcommand{\GenerateRepeats}{10}
+
+\newcommand{\nfilters}{8}
+\newcommand{\FilterCenter}{70 kHz}
+
+\newcommand{\CorrelationLow}{0.97}
+\newcommand{\CorrelationHigh}{$>$ 0.99}
+
+\newcommand{\pcaExplainedVar}{98\%}
+
+\newcommand{\Epochs}{1000}
+
+\newcommand{\FigVIIInumberII}{252}
+\newcommand{\FigVIIInumberIII}{63}
+\newcommand{\FigVIIInumberIIII}{296}
+
+\newcommand{\Compression}{3-5}
+
 %number of pca components
-\newcommand{\pca}{xx}
+\newcommand{\pca}{35}
 %number of neurons in models
-\newcommand{\nneurons}{xx}
+\newcommand{\nneurons}{250 neurons (not including the input and output layers)}
 %number of connections
-\newcommand{\nweights}{37 thousand}
+\newcommand{\nweights}{50 thousand}
 
@@
 They use the landscape intensively for foraging, roosting, and commuting.
-They have excellent spatial memory \citep{Barchi2013,VonHelversen2005} and can rely on sonar to navigate to notable locations like foraging grounds, drinking places \citep[see][for references]{Vanderelst2016,Vanderelst2017}.
+They have excellent spatial memory \citep{Barchi2013,VonHelversen2005} and can rely on sonar to navigate to notable locations like foraging grounds and drinking places \citep[see][for references]{Vanderelst2016,Vanderelst2017}.
+Bats deprived of sight have been found to successfully return to their roost when displaced by up to 40 miles \citep{Stones1969,Williams1966,Davis1970,Mueller1957}.
 Furthermore, bats have been reported to use sonar-based recognition of landmarks, both under experimental \citep{Jensen2005,Yu2019}, and natural \citep{Verboom1999} conditions.
-Echolocating bats can navigate to salient places \citep{Schnitzler2003}, relying solely on their sonar system.
-Indeed, bats deprived of sight have been found to successfully return to their roost when displaced by up to 40 miles \citep{Stones1969,Williams1966,Davis1970,Mueller1957}.
 
@@
 We previously demonstrated that bats might use sonar template-based scene recognition \citep{Vanderelst2016,Vanderelst2017}.
-In this approach, bats are assumed to compare the acoustic signature of the environment to a set of previously stored signatures.
+Under this hypothesis, bats compare the acoustic signature of the environment to a set of previously stored signatures.
 We operationalized these acoustic signatures as time-intensity profiles, which could be derived directly from the cochlear output.
@@
 Nevertheless, the output of the cochlear model can be mapped onto a low-dimensional space.
-In a second step, we show that echoes collected in real bat habitats can be mapped into the same low-dimensional space, without loss of information.
+In a second step, we show that echoes collected in real bat habitats can be mapped into the same low-dimensional space, without significant loss of information.
 We verify this by demonstrating that the low dimensional feature space allows recognizing locations in real habitats.
@@
 The distances of the point reflectors were selected using a two-step process.
-First we selected $n_{s}$ seed points $s$, uniformly distributed over a range from 0.5 to 7 meters.
+First we selected $n_{s}$ seed points $s$, uniformly distributed over a range from \MinDistance\ to \MaxDistance\ meters.
 Next, for every seed point $s$, we selected $n_{c}$ cloud points normally distributed around $s$, with standard deviation 0.25 m.
-We added a single reflector at a distance between xx and xx to each collection of points.
+We added a single reflector at a distance between \MinDistanceFirst\ and \MaxDistanceFirst\ to each collection of points.
 This two step process to modeling natural reflectors was inspired by the results of \citet{Yovel2009}, who found that echoes returning from vegetation could be better modeled using a similar two-step process than simply using uniformly spaced points.
-We generated xx sets of points for each combination of $n_s \in [xx,xx,xx, xx]$ and $n_c \in [xx, xx, xx, xx]$.
+We generated \GenerateRepeats\ sets of points for each combination of $n_s \in \ns$ and $n_c \in \nc$.
 Synthetic echoes consisting of few reflectors are likely to consist of only weak echoes.
@@
 Finally, the model includes a 1 kHz low-pass filter modeling the response characteristics of the inner hair cells.
-In the current implementation, we used a gammatone filterbank with xx filters, centered around xx kHz.
+In the current implementation, we used a gammatone filterbank with \nfilters\ filters, centered around \FilterCenter.
 The filters were spaced to correspond to equivalent rectangular bandwidths \citep{Glasberg1990}.
@@
 
-Figure \ref{fig:processing} depicts a single example of a synthetic echo sequence.
-It also illustrates the resulting response for each of the xx channels of the \citet{Wiegrebe2008} model.
+Figure \ref{fig:processing}a depicts a single example of a synthetic echo sequence.
+It also illustrates the resulting response for each of the \nfilters\ channels of the \citet{Wiegrebe2008} model (panels b-i).
 From these responses, it can be seen that they are highly correlated across channels.
-This is confirmed when analyzing the average cross-channel correlation for all xx synthetic echo sequences (Fig.
-\ref{fig:processing}xx).
+This is confirmed when analyzing the average cross-channel correlation for all synthetic echo sequences (Fig.
+\ref{fig:processing}i).
 We calculated the correlation $\bar{\rho}$ between the output $X_i$ of each frequency channel $i$ and the mean across channels $\bar{X}$ for each synthetic echo, as follows,
@@
 %
-The distribution of $\bar{\rho}$ is plotted in \ref{fig:processing}xx.
-From this distribution, it can be seen that the value of $\bar{\rho}$ ranged from xx to xx.
+The distribution of $\bar{\rho}$ is plotted in \ref{fig:processing}j.
+From this distribution, it can be seen that the value of $\bar{\rho}$ ranged from \CorrelationLow\ to \CorrelationHigh.
 From these high cross-channel correlations, we conclude that the output of the cochlear model can be summarized by averaging across channels without extensive loss of information.
 
-An example of such a cross-channel sum is plotted in figure \ref{fig:processing}xx.
+An example of such a cross-channel sum is plotted in figure \ref{fig:processing}k.
 In correspondence to our previous work \citep{Vanderelst2016}, we refer to this average across frequency channels as a template.
@@
 	\caption{Example of synthetic echo sequence and its processing using the \citet{Wiegrebe2008} model.
-(xx) Example of a single synthetic echo sequence (red).
+(a) Example of a single synthetic echo sequence (red).
 The dechirped version is plotted in black.
-(xx-xx) Response of the xx frequency channels in the \citet{Wiegrebe2008} model.
+(b-i) Response of the \nfilters\ frequency channels in the \citet{Wiegrebe2008} model.
 From these plots, it can be seen that channels correlate highly.
-This is confirmed in panel xx, depicting the average correlation across channels for all xx synthetic echo sequences analyzed.
+This is confirmed in panel j, depicting the average correlation across channels for all synthetic echo sequences analyzed.
 Note that the x-scale does not start at zero.
-(xx) The output of the \citet{Wiegrebe2008} model summed across the frequency channels, is referred to as a template in this paper.
-The template resulting from the echo sequence in panel xx is shown in panel xx.
-Panel xx shows the cumulative proportion of explained variance as a function of number of \pc\ across all templates for all synthetic echo sequences.}
+(k) The output of the \citet{Wiegrebe2008} model summed across the frequency channels, is referred to as a template in this paper.
+Panel (l) shows the cumulative proportion of explained variance as a function of number of \pc\ across all templates for all synthetic echo sequences.}
 	\label{fig:processing}
@@
 We used principal components analysis (\PCA\ ) to assess the dimensionality of the templates generated from the synthetic echo sequences.
-We found that almost all of the variance ($> 99\%$) in the templates could be captured by \pca\ components (see figure \ref{fig:processing}xx).
-Some examples of templates reconstructed from the principal components are shown in figure \ref{fig:synthetictemplates}xx).
+We found that almost all of the variance ($> 99\%$) in the templates could be captured by \pca\ components (see figure \ref{fig:processing}l).
+Some examples of templates reconstructed from the principal components are shown in figure \ref{fig:synthetictemplates}).
 
@@
 	\caption{Example of synthetic dechirped echo sequences, their corresponding templates and the templates reconstructed from the \pca\ \pcs.
-(xx-xx) Three examples of synthetic echo sequence waveforms.
-Panels xx-xx depict the resulting templates (i.e., summation across the output channels of the \citet{Wiegrebe2008} model).
+(a-c) Three examples of synthetic echo sequence waveforms.
+Panels (d-f) depict the resulting templates (i.e., summation across the output channels of the \citet{Wiegrebe2008} model).
 These panels also show the result of projecting the templates into a \pca D \pc\ space and transforming the result back to the original coordinates.}
@@
 This implies that, after filtering through (a model of) the bat's auditory system, these artificial echoes can be converted into low dimensional descriptions.
-The original echo sequences, generated by xx to xx point reflectors effectively map to \pca\ dimensional vector.
+The original echo sequences, generated by \MinDiracs\ to \MaxDiracs\ point reflectors effectively map to a \pca\ dimensional vector.
 In the remainder of the paper, we will assess whether this low dimensional space allows describing the echoes from real bat habitats as collected using an ensonification device.
@@
 We used the neural networks trained on the original data to classify these interpolated data.
-It was found that the neural network indeed classified these interpolated data to azimuths and elevations close to the positions from which they were interpolated (Figure \ref{fig:interpolation}).
 
@@
 We found that the templates derived from acoustic data collected in real habitats could be well represented by the \pca\ components derived from synthetic data.
-This \pca-dimensional space could represent xx\% of the variance in the templates.
+This \pca-dimensional space could represent \pcaExplainedVar\ of the variance in the templates.
 Figure \ref{fig:empiricaltemplates} shows examples of templates and their reconstructed counterparts.
@@
 This figure shows that, in all cases, training almost reached an asymptote around or before 250 epochs.
-However, some additional training can be seen up to xx epochs.
+However, some additional training can be seen up to \Epochs\ epochs (when learning was arbitrarily halted).
+
+%%
+%xx THIS PARAGRAPH TAKES NUMBERS FROM FIG 6
+%%
 
@@
 The performance was obtained by assuming that the most active output neuron represented the classification of the network.
-Inspecting the cumulative distribution of error reveals that the neural network encoding outperformed the perfect memory approach.
+Inspecting the cumulative distribution of error reveals that the neural network encoding performed as least as well as the perfect memory approach.
 The number of small errors was larger for the neural network than for the perfect memory.
-The neural network was able to classify about 70\% of templates with an error of fewer than 30 degrees in azimuth.
+The neural network was able to classify about 80\% of templates with an error of fewer than 30 degrees in azimuth.
 In elevation, almost all templates were classified with an error below 15 degrees.
-Most (70\%) errors in location attribution were smaller than 0.5 meters.
+Most (78\%) errors in location attribution were smaller than 0.5 meters.
 The confusion matrices confirm that most classification errors were small.
 
-\todo[inline]{update this to reflect latest results}
 Interestingly, for azimuth and elevation, classification errors were not uniform across directions.
 This can be seen by the classification entropy plotted in figure \ref{fig:israelperformance}.
-In particular, these curves show interesting peaks around zero azimuth and medium elevations.
+In particular, these curves show  peaks around zero azimuth and medium elevations.
 The geometry of the environment can explain these peaks.
@@
 The same is true for elevations above 0.
-Elevations around 25 return very few echoes and this serves as a reliable indicator for classifying these templates, explaining the drop in entropy for elevation = 25 degrees.
 
@@
 
+%%
+%xx THIS PARAGRAPH TAKES NUMBERS FROM FIG 7
+%%
+
 Figure \ref{fig:royalperformance} shows the results for the Royal Fort Park data.
 These results mirror those in figure \ref{fig:israelperformance}.
-Again the neural network outperformed the perfect memory approach by having more smaller errors.
-Overall, the performance for this data set was somewhat better than for the Israel data.
+Again the neural network performed at least as good as the perfect memory approach.
+For azimuth, the perfect memory approach slightly outperform the neural network (notice that y-scale in the top panels of this figure differ from figure \ref{fig:israelperformance}).
+Overall, the performance for this data set was better than for the Israel data.
+This is probably due to the fact that echoes were collected closer to reflectors (vegetation) than in the Israel data set, which leads to strong echoes.
 
@@
 	\caption{Classification results for the Royal Fort Park site.
-Top row: cumulative error distributions for the azimuth, elevation and location classification.
+Top row: cumulative error distributions for the azimuth, elevation and location classification (Notice the different y-scale for these plots than in figure \ref{fig:israelperformance}).
 Middle row: Confusion matrices for the classifications.
@@
 Figure \ref{fig:mechanism} shows the number of \pcs\ needed to capture 99\% of the templates variance.
-From this plot, it can be seen that the number of \pcs\ increases when shuffling the echo sequences from xx to xx.
-Omitting the 1 kHz filter from the  \citet{Wiegrebe2008} model increases the number of \pcs\ to xx.
+From this plot, it can be seen that the number of \pcs\ increases when shuffling the echo sequences from \pca\ to \FigVIIInumberIII\.
+Omitting the 1 kHz filter from the  \citet{Wiegrebe2008} model increases the number of \pcs\ to \FigVIIInumberII.
 Including the 1 kHz filter, the average spectrum of the templates (fig.
@@
 The structure of the echo signals also decreases the dimensionality of the templates.
-Therefore, we conclude that the sonar system of bats, through a number of filtering stages, compresses the 3D complexity of the world onto a lower dimensional representation.
+The sonar system of bats, through a number of filtering stages, compresses the 3D complexity of the world onto a lower dimensional representation.
 
@@
 However, the mouse cortex has been estimated to contain about $9 \cdot 10^4$ neurons or $7 \cdot 10^8$ synapses per cubic millimeter \citep{Braitenberg2013}.
-The neural networks presented in this paper consisted of about \nneurons\ neurons and \nweights\ connections.
+The neural networks presented in this paper consisted of \nneurons\ and \nweights\ connections.
 Therefore, to the extent that the proposed neural networks can be considered as biologically plausible models of memory, they are orders of magnitude smaller than a single cubic millimeter of mouse cortex.
@@
 Table \ref{tab:sizes} lists the number of weights for the artificial neural networks and compares this number with the size of the \pca-dimensional templates.
-The number of floating point numbers making up the templates is larger than the number of weights by a factor of 2-3.
+The number of floating point numbers making up the templates is larger than the number of weights by a factor of \Compression.
 Therefore, the networks have a smaller memory footprint than the data they encode.
@@
 	\toprule
-	Site &       T &     n &      W &  W (nb\} &      R \\
+	Site &       T &     n &      W &  W (nb) &      R \\
 	\midrule
-	Israel &  206661 &  5299 &  53322 &   52425 &  25.8\% \\
-	Royal &  302991 &  7769 &  51945 &   51075 &  17.1\% \\
+	Israel &  183430 &  5395 &  52875 &   51975 &  28.8\% \\
+	Royal &  264146 &  7769 &  51345 &   50475 &  19.4\% \\
 	\bottomrule

### 2020-03-25 Paper/main.tex
edits to captions etc

@@
 	\includegraphics[width=1\linewidth]{figures/networks}
-	\caption{Depiction of the neural network architecture trained for each site.}
+	\caption{Depiction of the neural network architecture trained for each site.
+The network had 4 hidden layers.
+The Gaussian Noise Layer adds noise with $\sigma(\stochasticnoise) = 0.01$, as described in the main text.}
 	\label{fig:networks}
@@
 	\includegraphics[width=1\linewidth]{../results/components}
-	\caption{}
+	\caption{Plot of the first 25 \pcs\ (out of \pca).
+The number in each panel indicates the number of zero crossings in the template.
+This number increases as a function of template number, indicating that most information is contained in lower template frequencies.}
 	\label{fig:components}
@@
 	\includegraphics[width=1\linewidth]{../results/interpolation}
-	\caption{sadasasasdasd}
+	\caption{Neural network categorization results for the templates obtained by interpolating between the 7 elevation and 31 azimuth directions ensonified for the Israel and Royal Fort site.
+(a,b) results for the azimuth and elevation for the Israel data.
+(c,d) Same for the Royal Fort data.
+Because the templates classified in these results were interpolated, the network can be considered to have correctly classified the template if it is assigned to direction $n$ or direction $n+1$.
+For example, a template obtained by interpolating between azimuth directions 140 and 130 degrees could be assigned to either these directions.
+The number in each panel indicates the percentage of data point on the diagonal (direction $n$) or the diagonal + 1 ($n+1$).}
 	\label{fig:interpolation}

### 2020-03-25 script_evaluate_interpolation.py

@@ -1,0 +1,79 @@
+import numpy
+import pandas
+from matplotlib import pyplot
+from tensorflow import keras
+
+from library import misc, process_functions, settings
+
+pyplot.figure(figsize=(5, 5))
+panel_index = 1
+
+for data_set in ['israel', 'royal']:
+    pca_model = misc.pickle_load(settings.pca_templates_model_file)
+    for selected_dimension in ['azs', 'els']:
+        file_names = misc.folder_names(data_set, selected_dimension)
+        # Read prepared data
+        data = numpy.load(file_names['npz_file_interpolated'])
+        # Select inputs
+        if selected_dimension == 'azs': unencoded_data = data['long_azs']
+        if selected_dimension == 'els': unencoded_data = data['long_els']
+        if selected_dimension == 'lcs' and data_set in ['israel', 'royal']: unencoded_data = misc.map_lcs_to_distances(data)
+
+        # Encode targets
+        encoder, targets = process_functions.get_encoding(unencoded_data)
+        target_n = targets.shape[1]
+
+        # Get PCA-ed template inputs
+        templates = data['long_data']
+        inputs = pca_model.transform(templates)
+        n_components = inputs.shape[1]
+
+        # Load neural network
+        neural_network = keras.models.load_model(file_names['model_file'])
+        predictions_matrix = neural_network.predict(inputs)
+        binary_predictions = misc.binarize_prediction(predictions_matrix)
+        interpreted_predictions = encoder.inverse_transform(binary_predictions)
+        interpreted_predictions = interpreted_predictions.flatten()
+
+        results = {'target': unencoded_data, 'prediction': interpreted_predictions}
+        results = pandas.DataFrame(results)
+        results['dummy'] = 1
+
+        # Get confusion table and plot it
+        colormap = settings.cmap
+        colormap.set_bad(color='gray')
+
+        pyplot.subplot(2, 2, panel_index)
+
+        ######################################
+        table, _ = misc.make_confusion_matrix(results, normalize=False)
+        ss = numpy.sum(table)
+        sd = numpy.sum(numpy.diag(table, k=0)) + numpy.sum(numpy.diag(table, k=1))
+        pct = '%.2f' % ((sd / ss) * 100) + '%'
+        #######################################
+
+        table, labels = misc.make_confusion_matrix(results, normalize=True)
[…]

### 2020-03-25 Paper/main.tex
after grammarly

@@
 First, \citet{Geberl2019} confirmed that bio-sonar has a low spatial resolution, both for distance and angular dimensions.
-Also, neurophysiological data suggest that closely spaced objects are not perceived seperatly by echolocating bats.
+Also, neurophysiological data suggest that closely spaced objects are not perceived separately by echolocating bats.
 In contrast, neural responses to echo cascades from densely spaced objects seem to represent a single extended stimulus \citep{Warnecke2018,Knowles2015}.
 
-However, it should be noted that such low dimensional descriptions have been shown to potentially contain a lot of useful information \citep{Kuc1997b,Kuc1997}.
+However, it should be noted that such low dimensional descriptions have been shown to contain a lot of useful information \citep{Kuc1997b,Kuc1997} potentially.
 We previously demonstrated that bats might use sonar template-based scene recognition \citep{Vanderelst2016,Vanderelst2017}.
@@
 The physics of the sonar system and the auditory periphery map the complexity of the 3D environment onto a small number of uncorrelated features, irrespective of whether we use artificial or real echo sequences.
-Nevertheless, this limited set of features is sufficiently rich for bats to recognize locations, and as we discuss, could also support flight control.
+Nevertheless, this limited set of features is sufficiently rich for bats to recognize locations, and as we discuss, they could also support flight control.
 
@@
 We added a single reflector at a distance between \MinDistanceFirst\ and \MaxDistanceFirst\ to each collection of points.
-This two step process to modeling natural reflectors was inspired by the results of \citet{Yovel2009}, who found that echoes returning from vegetation could be better modeled using a similar two-step process than simply using uniformly spaced points.
+This two-step process of modeling natural reflectors was inspired by the results of \citet{Yovel2009}, who found that echoes returning from vegetation could be better modeled using a similar two-step process than simply using uniformly spaced points.
 We generated \GenerateRepeats\ sets of points for each combination of $n_s \in \ns$ and $n_c \in \nc$.
@@
 The final amplitude of each Dirac impulse was determined by sampling from a normal distribution centered on the calculated amplitude and with a standard deviation of 10 dB.
-This final operation represents a very simple model of varying reflector strength due e.g., to shape and orientation of the reflecting feature.
+This last operation represents a straightforward model of varying reflector strength due, e.g., to shape and orientation of the reflecting feature.
 
@@
 Note that the x-scale does not start at zero.
-(k) The output of the \citet{Wiegrebe2008} model summed across the frequency channels, is referred to as a template in this paper.
-Panel (l) shows the cumulative proportion of explained variance as a function of number of \pc\ across all templates for all synthetic echo sequences.}
+(k) The output of the \citet{Wiegrebe2008} model summed across the frequency channels.
+This is referred to as a template in this paper.
+Panel (l) shows the cumulative proportion of explained variance as a function of the number of \pc\ across all templates for all synthetic echo sequences.}
 	\label{fig:processing}
@@
 In this section, we used a generative model for naturalistic echoes.
-We assumed that echoes sequences received by bats can be considered as generated by anisotropically distributed point reflectors.
+We assumed that echoes sequences received by bats could be considered as generated by anisotropically distributed point reflectors.
 Using this model, we have derived a low dimensional space ($D = \pca$), which can be used to describe the templates resulting from the artificial echo sequences.
@@
 
-We reused the large library of echoes gathered in real bat habitats by \citet{Vanderelst2016}.
+We reused the extensive library of echoes gathered in real bat habitats by \citet{Vanderelst2016}.
 Below, we briefly discuss the construction of templates based on these echoes, and we refer to our previous work for more details.
@@
 We concluded from analyzing the synthetic echoes that the response in different frequency channels is highly correlated.
-Therefore, the output of the cochlear model was averaged across frequency channels to obtain a time-intensity profile, or template.
+Therefore, the output of the cochlear model was averaged across frequency channels to obtain a time-intensity profile or template.
 As the beam of the emitter used was more narrow than the typically combined hearing and emission directionality of bats \citep{Vanderelst2010a,Reijniers2010,Jakobsen2012} templates were averaged across three neighboring directions.
@@
 The templates from both sites were mapped into the \pca-dimensional space derived from the synthetic echoes, described above.
-We assessed how well the templates collected in real habitats can be represented by the \pca\ components by mapping the templates into this space and transforming the results back to the original space.
+We assessed how well templates collected in real habitats could be represented by the \pca\ components by mapping the templates into this space and transforming the results back to the original space.
 Next, we calculated the correlation (and explained variance) between the original templates and the reconstructed templates.
@@
 We assessed whether the templates projected into the \pca-dimensional space still contained sufficient information to be recognized.
-Previously, we assessed whether templates can be used to recognize locations using, what \citet{Baddeley2012} called, a perfect memory approach.
+Previously, we assessed whether templates could be used to recognize locations using, what \citet{Baddeley2012} called, a perfect memory approach.
 Templates were stored in a lookup table.
 The Euclidean distance between two echo templates was the metric used to assess how well bats should be able to distinguish between templates.
-This table lookup and distance calculation requires a comprehensive search of memory that scales with the number of templates memorized.
+This table lookup and distance calculation require a comprehensive search of memory that scales with the number of templates memorized.
 Hence, in our previous work, we established the theoretical possibility of classifying templates based on real echoes from environments bats were known to frequent, irrespective of the significant amount of memory and computational resources required to do so.
-Here, we use a more biologically plausible memory system: an artificial neural network, i.e.
-a neurally inspired way of distributed information storage \citep{Mcleod1998}.
+Here, we use a more biologically plausible memory system: an artificial neural network, i.e., a neurally inspired way of distributed information storage \citep{Mcleod1998}.
 While we acknowledge that current neural network training methods most likely have no biological counterpart, we propose that this approach provides us with a more biologically plausible appraisal of the memory and computational effort required to implement a similar scheme in a bat brain.
@@
 The following layers were fully connected and used the rectified linear activation function.
-The output layer uses a softmax activation function, i.e.
-the normalized exponential function.
+The output layer uses a softmax activation function, i.e., the normalized exponential function.
 The loss was calculated as the cross-entropy between the target output and the actual output of the neural network.
@@
 Instead, we are only interested in whether the neural networks can learn and classify the set of training examples.
-Nevertheless, it would be beneficial if the neural networks would be capable of classifying templates corresponding to directions that are slightly different than those included in the data sets.
+Nevertheless, it would be beneficial if the neural networks would be capable of classifying templates corresponding to directions that are slightly different from those included in the data sets.
 Indeed, this is a feature of the Euclidean distance measure used in implementing the perfect memory.
-In that approach, novel templates will be classified to positions with similar templates.
-We verified whether this properties was also found in the neural networks the case by generating a set of templates by interpolating in azimuth and elevation between the directions ensonified in the data set.
-In our data sets, we ensonified 31 different azimuth directions and 7 elevation directions.
-The interpolated data consisted of 30 azimuths and 7 elevations.
+In that approach, novel templates will be classified as positions associated with similar templates.
+We verified whether this property was also found in the neural networks the case by generating a set of templates by interpolating in azimuth and elevation between the directions ensonified in the data set.
+In our data sets, we ensonified 31 different azimuth directions and seven elevation directions.
+The interpolated data consisted of 30 azimuths and six elevations.
 We used the neural networks trained on the original data to classify these interpolated data.
@@
 The corridor (fig.
-\ref{fig:datacollection}) does not return a lot of echoes from azimuths around zero.
+\ref{fig:datacollection}) does not return many echoes from azimuths around zero.
 Indeed, this direction is mostly open space.
@@
 Again the neural network performed at least as good as the perfect memory approach.
-For azimuth, the perfect memory approach slightly outperform the neural network (notice that y-scale in the top panels of this figure differ from figure \ref{fig:israelperformance}).
+For azimuth, the perfect memory approach outperforms the neural network (notice that the y-scale in the top panels of this figure differ from figure \ref{fig:israelperformance}).
 Overall, the performance for this data set was better than for the Israel data.
-This is probably due to the fact that echoes were collected closer to reflectors (vegetation) than in the Israel data set, which leads to strong echoes.
+Probably, this is because echoes were collected closer to reflectors (vegetation) than in the Israel data set, which leads to strong echoes.
 
@@
 
-We assumed echoes received by bats can be modeled as originating from a large collection of point reflectors.
+We assumed echoes received by bats could be modeled as originating from a large collection of point reflectors.
 Passing these synthetic echoes through a model of the auditory periphery, we obtained a time-intensity profile (template) for each echo sequence.
@@
 This suggests that the filtering of the auditory periphery provides the bat with a significantly compressed acoustic description of its environment.
-To identify the locus of the compression, we performed a number of additional analyses.
-In particular, we evaluated whether the low-dimensional description is a property of the echo sequences themselves, or whether this property results from the finite bandwidths of the gammatone filters and/or the low-pass filter in the \citet{Wiegrebe2008} model.
+To identify the locus of the compression, we performed additional analyses.
+In particular, we evaluated whether the low-dimensional description is a property of the echo sequences themselves, or whether this property results from the finite bandwidths of the gammatone filters or the low-pass filter in the \citet{Wiegrebe2008} model.
 
-We generated templates from the synthetic echo sequences using two versions of the \citet{Wiegrebe2008} model; with and without the 1 Khz low-pass filter.
+We generated templates from the synthetic echo sequences using two versions of the \citet{Wiegrebe2008} model, with and without the 1 kHz low-pass filter.
 Per default, the model includes this filter to model the response characteristics of the inner hair cells \citep{Meddis2006,Reijniers2010a}.
-By removing this lowpass filter, we can assess to what extent it is responsible for the observed compressibility of the templates.
+By removing this low-pass filter, we can assess to what extent it is responsible for the observed compressibility of the templates.
 We also generated templates using echo sequences of which the samples were randomly shuffled.
-This allows inferring whether the low-dimensionality of the templates is a property of the echo signals (before being filtered by the cochlear model).
+This allows testing whether the templates' low-dimensionality is a property of the echo signals (before being filtered by the cochlear model).
 
@@
 The structure of the echo signals also decreases the dimensionality of the templates.
-The sonar system of bats, through a number of filtering stages, compresses the 3D complexity of the world onto a lower dimensional representation.
+The sonar system of bats, through several filtering stages, compresses the 3D complexity of the world onto a lower-dimensional representation.
 
@@
 They also rely on sonar for flight control.
-This includes, avoiding obstacles and/or following guiding structures.
+This includes avoiding obstacles and following guiding structures.
 The environmental description the bat has access to for flight control starts from the same cochlear output used for place recognition.
@@
 Indeed, we have previously demonstrated that a low-dimensional description is sufficient for obstacle avoidance and corridor following \citep{Vanderelst2015a,Mansour2019}.
-In that work, we converted both the echoes at the left and at the right ear into a template-like representation used as input to the motion controller.
+In that work, we converted the echoes at the left and right ear into a template-like representation.
+We used this representation as input to the robot controller.
 Subsequent work will have to demonstrate whether the parameters of the \pca D space presented here can be used directly for controlling flight.
@@
 
-Despite their small size, the networks in this paper were able to encode (memorize) the templates at least as well as the perfect memory approach.
+Despite their small size, the neural networks were able to encode (memorize) the templates at least as well as the perfect memory approach.
 The neural networks presented here are tiny compared to the density of cortical matter.
@@
 Table \ref{tab:sizes} lists the number of weights for the artificial neural networks and compares this number with the size of the \pca-dimensional templates.
-The number of floating point numbers making up the templates is larger than the number of weights by a factor of \Compression.
+The number of floating-point numbers making up the templates is larger than the number of weights by a factor of \Compression.
 Therefore, the networks have a smaller memory footprint than the data they encode.
-It should be noted that we have not made an attempt to try and to optimize (minimize) the size of the networks.
-From these numbers, we conclude that the low-dimensional signals provided by the bat's sonar system result in a low memory footprint and that they would use little neural resources to be stored and processed.
+It should be noted that we have not attempted to try and to optimize (minimize) the size of the networks.
+From these numbers, we conclude that the low-dimensional signals provided by the bat's sonar system result in a low memory footprint and that they would use limited neural resources to be stored and processed.
 
@@
 	\label{tab:sizes}
-	\caption{Table indicating the size of both template data sets and number of weights in the networks.
+	\caption{Table indicating the size of both template data sets and the number of weights in the networks.
 The template size $T$ is obtained as $\pca \times n$, with $n$ the number of templates in the set.
-The number of weights $W$ includes all connections between neurons as well as the bias weights for each of the three networks trained.
+The number of weights $W$ includes all connections between neurons and the bias weights.
 $W (nb)$ omits the bias weights.
@@
 
-It has been suggested before that bats' sonar systems provides detailed information about the world.
-A number of authors have referred to sonar as providing the bat with an acoustic image \citep{Lee2017,Barchi2013,Moss2001,Schnitzler2003,Simmons2012,Ulanovsky2008,Clare2015,Surlykke2016,Geipel2013}.
+It has been suggested that bats' sonar systems provide detailed information about the world.
+Several authors have referred to sonar as providing the bat with an acoustic image \citep{Lee2017,Barchi2013,Moss2001,Schnitzler2003,Simmons2012,Ulanovsky2008,Clare2015,Surlykke2016,Geipel2013}.
 As mentioned above, this is at odds with some recent behavioral and neurophysiological findings in bats faced with complex echo sequences \citep{Knowles2015,Geberl2013,Warnecke2018}.
 However, it is also at odds with the various filtering stages involved in the emission and the reception of bio-sonar.
-In particular, we argued before that the neurophysiology of the  bats auditory system represents a bottleneck in the amount of raw echo signal information accessible to the bat.
+In particular, we argued before that the neurophysiology of the bats' auditory system represents a bottleneck in the amount of raw echo signal information accessible to the bat.
 Indeed, \citet{Reijniers2010a} already noticed that significant compression occurs during cochlear transduction.
@@
 	\includegraphics[width=1\linewidth]{../results/interpolation}
-	\caption{Neural network categorization results for the templates obtained by interpolating between the 7 elevation and 31 azimuth directions ensonified for the Israel and Royal Fort site.
+	\caption{Neural network categorization results for the templates obtained by interpolating between the seven elevation and 31 azimuth directions ensonified for the Israel and Royal Fort site.
 (a,b) results for the azimuth and elevation for the Israel data.

### 2020-03-25 RUN_ALL.py
Final version of code?

@@ -1,23 +1,37 @@
+def run(file_name):
+    file_name = file_name + '.py'
+    with open(file_name) as f:
+        code = compile(f.read(), file_name, 'exec')
+        exec(code)
+
+#%%
 import script_preprocess_empirical_data
 
 # Get space based on synthetic echoes
-import script_synthetic_echoes_to_PCA
-import script_plot_processing_synthetic
-
+run('script_synthetic_echoes_to_PCA')
+run('script_plot_processing_synthetic')
+#%%
 # Visualize reconstruction of syn and emp templates
-import script_plot_template_examples
+run('script_plot_template_examples')
+#%%Train the neural networks
+run('script_run_training')
+run('script_plot_training_history')
 
-#Train the neural networks
-import script_run_training
-import script_plot_training_history
+#%%
 
-# For both israel and royal do...
-import script_run_perfect_memory
+run('script_run_perfect_memory')
+#%%
 import script_plot_training_results
-
+run('script_plot_training_results')
+#%%
 # Assess where the reduction happens
-import script_run_identify_mechanism
+run('script_run_identify_mechanism')
 
+#%%
+run('script_evaluate_interpolation')
 
-import script_tabulate_memory_use
-import script_plot_components
+#%%
+
+run('script_tabulate_memory_use')
+run('script_plot_components')
+

### 2020-03-25 interpolated.py

@@ -1,67 +1,0 @@
-import numpy
-import pandas
-import os
-from matplotlib import pyplot
-from tensorflow import keras
-from scipy.signal import convolve
-
-from library import misc, process_functions, settings
-
-# for data_set in ['royal', 'israel']:
-#     for selected_dimension in ['lcs', 'azs','els']:
-
-for data_set in ['israel']:
-    for selected_dimension in ['azs']:
-        file_names = misc.folder_names(data_set, selected_dimension)
-        # Read prepared data
-        data = numpy.load(file_names['npz_file'])
-        pca_model = misc.pickle_load(settings.pca_templates_model_file)
-
-        # Select inputs
-        if selected_dimension == 'azs': unencoded_data = data['long_azs']
-        if selected_dimension == 'els': unencoded_data = data['long_els']
-        if selected_dimension == 'lcs' and data_set in ['israel', 'royal']: unencoded_data = misc.map_lcs_to_distances(data)
-
-        # Encode targets
-        encoder, targets = process_functions.get_encoding(unencoded_data)
-        target_n = targets.shape[1]
-
-        # Get PCA-ed template inputs
-        templates = data['long_data']
-        inputs = pca_model.transform(templates)
-        n_components = inputs.shape[1]
-
-        # Load neural network
-        neural_network = keras.models.load_model(file_names['model_file'])
-
-        predictions_matrix = neural_network.predict(inputs)
-        binary_predictions = misc.binarize_prediction(predictions_matrix)
-        interpreted_predictions = encoder.inverse_transform(binary_predictions)
-        interpreted_predictions = interpreted_predictions.flatten()
-
-        #pyplot.scatter(unencoded_data, interpreted_predictions)
-        #pyplot.show()
-        neural_network.evaluate(inputs, targets)
-
-        results = {'target': unencoded_data, 'prediction': interpreted_predictions}
-        results = pandas.DataFrame(results)
-        results['dummy'] = 1
-
-        # Get confusion table and plot it
-        table, labels = misc.make_confusion_matrix(results)
-        pyplot.matshow(table)
-        pyplot.colorbar()
-        pyplot.title(data_set + ' ' + selected_dimension)
-        pyplot.show()
-
-        m = numpy.mean(templates, axis=0)
-        pyplot.plot(m)
-        pyplot.show()
-
-        m = numpy.mean(inputs, axis=0)
-        pyplot.plot(m)
-        pyplot.show()
-
-
-
-

### 2020-03-25 library/misc.py

@@ -7,8 +7,22 @@ import scipy.interpolate as interpolate
 from matplotlib import pyplot
 from sklearn.preprocessing import Normalizer
 import string
-
+from scipy.stats import linregress
 from sklearn.decomposition import PCA
+from sklearn import preprocessing
+
+def cochleogram2r2(cochleogram):
+    cochleogram  = preprocessing.normalize(cochleogram)
+    c = cochleogram.shape[0]
+    n = cochleogram.shape[1]
+    flat = numpy.reshape(cochleogram, (c * n))
+    mean = numpy.mean(cochleogram, axis=0)
+    mean = numpy.tile(mean, (1, 8))
+    mean = mean.flatten()
+    slope, intercept, r_value, p_value, std_err = linregress(flat, mean)
+    r2 = r_value ** 2
+    return r2
+
 
 def template2fft(template):
     template = template - numpy.mean(template)
@@ -16,13 +30,14 @@ def template2fft(template):
     return spectrum
 
 
-def label(x,y , txt, ax=None, fontsize=12):
+def label(x,y , txt, ax=None, fontsize=12, color='black'):
     if type(txt)==int: txt = '(' + string.ascii_letters[txt] + ')'
     if ax is None: ax = pyplot.gca()
     pyplot.text(x, y, txt, transform=ax.transAxes,
                 horizontalalignment='center',
                 verticalalignment = 'center',
-                fontsize=fontsize)
+                fontsize=fontsize,
+                color=color)
 
 def average_correlation_matrix(array):
     size =array.shape[0]
@@ -219,20 +234,24 @@ def make_confusion_matrix(results, normalize=True):
     return table, labels
 
 
-def label_confusion_matrix(labels):
+def label_confusion_matrix(labels, shift_xaxis=0):
     ticks = list(range(0, len(labels)))
     if len(labels) == 7: ticks = list(range(0, 7, 2))
-    if len(labels) == 31: ticks = list(range(6, 31, 6))
+    if len(labels) == 31: ticks = list(range(1, 31, 10)) + [30]
     if len(labels) == 50: ticks = list(range(10, 50, 10))
     if len(labels) == 40: ticks = list(range(10, 40, 10))
     ticks_locs = numpy.array(ticks)
 
-    new_labels = []
-    for x in labels: new_labels.append('%.1f' % x)
-    new_labels = numpy.array(new_labels)
[…]

### 2020-03-25 library/process_functions.py

@@ -1,12 +1,14 @@
 import math
-from library import misc, settings
 import os
+
 import natsort
 import numpy
 import scipy.io as io
+from scipy.interpolate import RegularGridInterpolator
 from scipy.signal import convolve
 from sklearn.preprocessing import OneHotEncoder
-from scipy.interpolate import RegularGridInterpolator
+
+from library import misc, settings
 
 
 def preprocess(file_name, verbose=True):
@@ -59,7 +61,7 @@ def preprocess(file_name, verbose=True):
     return processed, az_box, el_box
 
 
-def process_data_set(data_set, filter_threshold = 0.1, interpolate_directions=False):
+def process_data_set(data_set, filter_threshold=0.1, interpolate_directions=False):
     print('#' * 10)
     print(data_set)
     print('#' * 10)
@@ -91,11 +93,8 @@ def process_data_set(data_set, filter_threshold = 0.1, interpolate_directions=Fa
 
     # Select only those templates above threshold
     summed = numpy.sum(all_mns, axis=3)
-    summed = numpy.array(summed)
     threshold = numpy.min(summed) + filter_threshold
     include = summed > threshold
-
-    print(threshold)
 
     if interpolate_directions:
         print('---> INTERPOLATE')
@@ -105,24 +104,23 @@ def process_data_set(data_set, filter_threshold = 0.1, interpolate_directions=Fa
         d2 = numpy.arange(0, 31)
         d3 = numpy.arange(0, n_samples)
 
-        d1i = d1 + 0 #elevations
-        d2i = d2 + 0.5 #azimuths
+        d1i = d1 + 0.5  # elevations
+        d2i = d2 + 0.5  # azimuths
 
-        a,b,c,d = numpy.meshgrid(d0, d1i, d2i, d3, indexing='ij')
+        a, b, c, d = numpy.meshgrid(d0, d1i, d2i, d3, indexing='ij')
         print(a.shape)
         print(b.shape)
         print(c.shape)
         print(d.shape)
 
-        R = RegularGridInterpolator((d0, d1, d2, d3), all_mns, bounds_error=False,fill_value=1)
-        all_mns = R((a,b,c,d))
-        print(all_mns.shape)
+        R = RegularGridInterpolator((d0, d1, d2, d3), all_mns, bounds_error=False, fill_value=None)
+        all_mns = R((a, b, c, d))
 
-        # Average across azimuth elevation to get interpolated version
[…]

### 2020-03-25 library/settings.py

@@ -1,7 +1,7 @@
 import os
 import numpy
 import palettable.colorbrewer.qualitative as qualitative
-
+import matplotlib
 ####################################################
 ##General settings
 ####################################################
@@ -73,6 +73,8 @@ default_line_color01 = '#46433A'
 default_line_color02 = '#CE534D'
 
 colormap = 'hot'
+cmap = matplotlib.cm.hot
+
 azs_color = '#e41a1c'
 els_color = '#377eb8'
 lcs_color = '#4daf4a'
@@ -88,5 +90,5 @@ israel_results_plot = os.path.join(result_folder, 'israel_results.pdf')
 royal_results_plot = os.path.join(result_folder, 'royal_results.pdf')
 
 pca_components_plot = os.path.join(result_folder, 'components.pdf')
-
 mechanism_plot = os.path.join(result_folder, 'mechanism.pdf')
+interpolation_results_plot = os.path.join(result_folder, 'interpolation.pdf')

### 2020-03-25 script_evaluate_interpolation.py

@@ -47,26 +47,32 @@ for data_set in ['israel', 'royal']:
 
         ######################################
         table, _ = misc.make_confusion_matrix(results, normalize=False)
-        ss = numpy.sum(table)
-        sd = numpy.sum(numpy.diag(table, k=0)) + numpy.sum(numpy.diag(table, k=1))
+        table[-1, :] = numpy.nan
+        table[:, -1] = numpy.nan
+
+        ss = numpy.nansum(table)
+        sd = numpy.nansum(numpy.diag(table, k=0)) + numpy.nansum(numpy.diag(table, k=1))
         pct = '%.2f' % ((sd / ss) * 100) + '%'
         #######################################
 
         table, labels = misc.make_confusion_matrix(results, normalize=True)
-        labels = labels + 5
 
+        # Mask data from positions that were extrapolated
         table[-1, :] = None
         table[:, -1] = None
-        pyplot.imshow(table, cmap=colormap, vmin=0, vmax=1)
 
-        misc.label_confusion_matrix(labels)
+        pyplot.imshow(table, cmap=colormap, vmin=0, vmax=1)
+        misc.label_confusion_matrix(labels, shift_xaxis=5)
+
+        # Hide the gray data
         pyplot.xlim([-0.5, len(labels) - 1.5])
         pyplot.ylim([-0.5, len(labels) - 1.5])
 
         misc.label(0.9, 0.1, panel_index - 1, color='white')
         misc.label(0.25, 0.9, pct, color='white')
-        if panel_index in [1, 4]: pyplot.ylabel('Target')
-        if panel_index in [4, 5, 6]: pyplot.xlabel('Output')
+        if panel_index in [1, 3]: pyplot.ylabel('Target (interpolated positions)')
+        if panel_index in [3, 4]: pyplot.xlabel('Output (original directions)')
+
         panel_index = panel_index + 1
 
 

### 2020-03-25 script_plot_processing_synthetic.py

@@ -74,7 +74,6 @@ for x in range(8):
     ax.text(5000, 0.5, title)
     misc.label(0.95, 0.8, x+1, fontsize=20)
 
-
 # ax = fig.add_subplot(gs[0:4, 2])
 # ax.plot(cumulative, color=settings.default_line_color)
 # ax.set_ylabel('Cumm. expl. variance')
@@ -82,39 +81,28 @@ for x in range(8):
 
 ax = fig.add_subplot(gs[0:2, 2])
 r = ax.hist(correlations, color=settings.default_line_color01, alpha=0.5)
-r = ax.hist(correlations**2, color=settings.default_line_color02, alpha=0.5)
 ax.set_ylabel('Echo Seq. Count')
-ax.set_xlabel('Cross-channel correlation/variance')
-ax.set_xlim([0.75, 1])
-ax.legend(['Correlation', 'Covariance'], loc=3)
-
+ax.set_xlabel('Correlation, $\\bar{\\rho}$')
+ax.set_xlim([0.95, 1])
+#ax.legend(['Correlation', 'Covariance'], loc=3)
 
 mn = numpy.mean(correlations)
 h = numpy.max(r[0])
 ax.vlines(mn,0,h,colors=settings.default_line_color01)
-misc.label(0.05, 0.8, x+1, fontsize=20)
-
-mn = numpy.mean(correlations**2)
-h = numpy.max(r[0])
-ax.vlines(mn,0,h,colors=settings.default_line_color02)
-misc.label(0.05, 0.8, x+1, fontsize=20)
+misc.label(0.05, 0.8, x+2, fontsize=20)
 
 ax = fig.add_subplot(gs[2:6, 2])
 ax.plot(template, color=settings.default_line_color01)
 ax.set_xticks([])
 ax.set_yticks([])
 ax.set_ylim([-0.1,1.1])
-misc.label(0.95, 0.95, x+2, fontsize=20)
-
+misc.label(0.95, 0.95, x+3, fontsize=20)
 
 ax = fig.add_subplot(gs[6:8, 2])
 ax.plot(cumulative, color=settings.default_line_color01)
 ax.set_ylabel('Cumm. expl. variance')
 ax.set_xlabel('Nr of components')
-misc.label(0.95, 0.8, x+3, fontsize=20)
-
-
-
+misc.label(0.95, 0.8, x+4, fontsize=20)
 
 pyplot.tight_layout()
 pyplot.savefig(settings.synthetic_processing_plot)

### 2020-03-25 script_plot_training_results.py

@@ -4,101 +4,103 @@ import os
 from library import misc, settings
 
 pyplot.style.use(settings.style)
-
-data_set = 'royal'
-
 loss_function = 'Categorical Cross Entropy'
 
-files_lcs = misc.folder_names(data_set, 'lcs')
-files_azs = misc.folder_names(data_set, 'azs')
-files_els = misc.folder_names(data_set, 'els')
-results_azs = misc.pickle_load(files_azs['results_file'])
-results_els = misc.pickle_load(files_els['results_file'])
-results_lcs = misc.pickle_load(files_lcs['results_file'])
-perfect_memory = misc.pickle_load(files_lcs['perfect_memory_file'])
+for data_set in ['israel','royal']:
+
+    files_lcs = misc.folder_names(data_set, 'lcs')
+    files_azs = misc.folder_names(data_set, 'azs')
+    files_els = misc.folder_names(data_set, 'els')
+    results_azs = misc.pickle_load(files_azs['results_file'])
+    results_els = misc.pickle_load(files_els['results_file'])
+    results_lcs = misc.pickle_load(files_lcs['results_file'])
+    perfect_memory = misc.pickle_load(files_lcs['perfect_memory_file'])
 
 
-# %% Plot errors
+    # %% Plot errors
 
-fig, axes = pyplot.subplots(nrows=3, ncols=3)
-fig.set_figheight(10)
-fig.set_figwidth(12)
+    fig, axes = pyplot.subplots(nrows=3, ncols=3)
+    fig.set_figheight(10)
+    fig.set_figwidth(12)
 
-pyplot.sca(axes[0, 0])
-counts = misc.get_error_histogram(results_azs, normalize=True, cummmulative=True)
-counts_pm = misc.get_error_histogram(perfect_memory['results_azs'], normalize=True, cummmulative=True)
-pyplot.plot(counts.error, counts.number, color=settings.azs_color)
-pyplot.plot(counts_pm.error, counts_pm.number, color='k', alpha=0.5)
-_, entries = misc.plot_inference_lines(counts.error, counts.number, [30, 45, 90])
-pyplot.ylim([0, 1.1])
-pyplot.xlabel('Error (degrees)')
-pyplot.ylabel('Proportion of templates')
-pyplot.legend(['Neural network', 'Perfect memory'] + entries)
-pyplot.title('Azimuth')
[…]

### 2020-03-25 script_preprocess_empirical_data.py

@@ -1,29 +1,7 @@
-from library import process_functions, misc
-import numpy
+from library import process_functions
 
-#process_functions.process_data_set('royal')
+process_functions.process_data_set('royal')
 process_functions.process_data_set('israel')
 
-#process_functions.process_data_set('royal', interpolate_directions=True)
+process_functions.process_data_set('royal', interpolate_directions=True)
 process_functions.process_data_set('israel', interpolate_directions=True)
-
-#%%
-from matplotlib import pyplot
-
-data_set = 'israel'
-file_names = misc.folder_names(data_set, None)
-
-
-data1 = numpy.load(file_names['npz_file'])
-data1 = data1['long_data']
-print(data1.shape)
-
-data2 = numpy.load(file_names['npz_file_interpolated'])
-data2 = data2['long_data']
-print(data2.shape)
-
-i = 250
-
-pyplot.plot(data1[i,:])
-pyplot.plot(data2[i,:])
-pyplot.show()

### 2020-03-25 script_run_perfect_memory.py

@@ -4,68 +4,67 @@ import scipy.spatial.distance as distance
 
 from library import misc, process_functions, settings
 
-data_set = 'israel'
 generate_data = False
 iterations = 10
 
-print('Running perfect memory for', data_set)
 
-file_names = misc.folder_names(data_set, None)
-output_file = file_names['perfect_memory_file']
-if generate_data: process_functions.process_data_set(data_set)
+for data_set in ['royal', 'israel']:
 
-# Read prepared data
-data = numpy.load(file_names['npz_file'])
-corridor_distances = misc.map_lcs_to_distances(data)
+    print('Running perfect memory for', data_set)
 
-pca = misc.pickle_load(settings.pca_templates_model_file)
-loaded = numpy.load(settings.synthetic_templates_pca_results)
-transformed = loaded['transformed']
-n_components_templates = transformed.shape[1]
+    file_names = misc.folder_names(data_set, None)
+    output_file = file_names['perfect_memory_file']
+    if generate_data: process_functions.process_data_set(data_set)
+
+    # Read prepared data
+    data = numpy.load(file_names['npz_file'])
+    corridor_distances = misc.map_lcs_to_distances(data)
+
+    pca = misc.pickle_load(settings.pca_templates_model_file)
+    loaded = numpy.load(settings.synthetic_templates_pca_results)
+    transformed = loaded['transformed']
+    n_components_templates = transformed.shape[1]
+
+    correct_ids = data['ids']
+    correct_ids[:, 0] = corridor_distances
+
+    templates = data['long_data']
+    pca_templates = pca.transform(templates)
+    inputs = pca_templates[:, :n_components_templates]
 
 
-correct_ids = data['ids']
-correct_ids[:, 0] = corridor_distances
+    summed_ids = numpy.zeros(correct_ids.shape)
+    for repeat in range(iterations):
+        print('iteration', repeat, '/', iterations)
+        # Make noisey templates
+        noise = numpy.random.normal(0, settings.stochaistic_noise, inputs.shape)
+        noisy_templates = inputs + noise
+        # rows: inputs, cols: noisy_templates
+        distances = distance.cdist(inputs, noisy_templates)
+        # for each input, what noisy template is closest?
+        closest = numpy.argmin(distances, axis=1)
+        summed_ids = summed_ids + correct_ids[closest, :]
[…]

### 2020-03-25 script_run_training.py

@@ -6,10 +6,10 @@ from tensorflow import keras
 
 from library import misc, process_functions, settings
 
-#for data_set in ['israel', 'royal']:
-for data_set in ['israel']:
+for data_set in ['israel', 'royal']:
+#for data_set in ['royal']:
     layers = [50, 75, 75, 50]
-    nr_epochs = 250
+    nr_epochs = 1000
 
     for selected_dimension in ['lcs', 'azs','els']:
         file_names = misc.folder_names(data_set, selected_dimension)

### 2020-03-25 script_synthetic_echoes_to_PCA.py

@@ -9,9 +9,9 @@ repeats = 10
 n_seed_levels = [5, 10, 20, 50, 100, 200]
 n_cloud_levels = [5, 10, 20, 50, 100, 200]
 
-#repeats = 10
-#n_seed_levels = [100, 200]
-#n_cloud_levels = [100, 200]
+#repeats = 3
+#n_seed_levels = [5]
+#n_cloud_levels = [5]
 
 sample_frequency = settings.sample_frequency
 number_of_samples = settings.raw_collected_samples
@@ -51,7 +51,9 @@ for seed_level_i in range(len(n_seed_levels)):
 
             c = numpy.corrcoef(cochleogram)
             r = misc.average_correlation_matrix(c)
+            #r = misc.cochleogram2r2(cochleogram)
             correlations.append(r)
+
 
             synthetic_impulses.append(impulse_response)
             synthetic_echoes.append(echo_sequence)
@@ -99,3 +101,7 @@ transformed, reconstructed, correlation = misc.project_and_reconstruct(pca_model
 
 print(results_templates)
 print(correlation)
+
+
+
+

### 2020-03-25 dechirp_test.py
cleaned up directory a bit

@@ -1,32 +1,0 @@
-from numpy.fft import fft, ifft
-from numpy import conj, pad, abs
-from library import generate_functions
-from matplotlib import pyplot
-
-pyplot.figure(figsize=(10,10))
-
-emission = generate_functions.create_emission()
-signal = generate_functions.create_emission()
-
-emission = pad(emission, (0, 100), 'constant')
-signal = pad(signal, (50, 50), 'constant')
-
-a = conj(fft(emission))
-b = fft(signal)
-c = ifft(a * b)
-dechriped = abs(c)
-
-
-pyplot.subplot(2,2,1)
-pyplot.plot(emission)
-pyplot.title('Emission')
-
-pyplot.subplot(2,2,2)
-pyplot.plot(signal)
-pyplot.title('Signal == Emission')
-
-pyplot.subplot(2,2,3)
-pyplot.plot(dechriped)
-pyplot.title('Dechirped signal (magnitude)')
-
-pyplot.show()

### 2020-03-25 todo_list.py

@@ -1,2 +1,0 @@
-#todo: Implement the dechirping the wiegrebe model properly. Now, results are complex numbers. That can't be correct.
-
